### 2017-01-09 introduction.tex
Initial Commit

\section{Introduction}

This file contains comments made to papers. The comments are structured as five
paragraph essays about those papers.

Each paper is identified by the code \texttt{IND\_XXX}, where \texttt{XXX} is a
number with 3 leading zeros. Each paper is also identified by a reference in
the references section. As some papers reference each other, these references
are kept in the comments.

### 2017-01-09 papers.tex
Initial Commit

\section{Papers}

% =============================================================================
\subsection{\texttt{IND\_007} - Data Mining Algorithms to Classify Students}

Paper is \TODO{cite}.

This paper presents the results of a few experiments made to Moodle data. It
also presents a data mining tool for Moodle called Moodle Data Mining Tool.
Finally, the papers shows some results on the accuracy of some algorithms.

The first section, after the introduction, talks about the background of data
mining in \gls{lms}. It goes on describing a few different tasks and citing
other works in the area. The second section describes the Moodle Data Mining
Tool. The third section presents a few experiments and their results. The
experiments intend on evaluating the performance of some algorithms. The
algorithms are executed with data from 7 courses of the Moodle of Cordoba
University.

This papers is very useful for it gives a few resources on data mining in
\gls{lms}. The resources state different techniques and models used in Moodle
Data, and from here it is possible to explore similar works.

The experiments are not very detailed. A few points are made on resampling, for
the data has a clear imbalance. Then the accuracy of some algorithms executed
with the University's data is shown. It is visible that the accuracy of every
algorithm is not better then 65\%, which is very strange.

A lot of data mining methods and algorithms are listed. The experiments don't
really go into detail, and the Moodle Data Mining Tool doesn't play a big role
in the experiments or in the conclusion. Nevertheless, it is a great paper that
points to useful resources.

### 2017-01-11 introduction.tex
Better introduction

@@
 
-This file contains comments made to papers. The comments are structured as five
-paragraph essays about those papers.
+This document contains comments made to several papers which relate Data Mining
+done to Moodle or other~\gls{lms} systems. The document is done in the scope of
+a project where data from the Moodle of the University of Évora is mined. The
+state of the art of that project is done from research described in this
+document.
+
+Each paper in this document has a general comment about it, a comment about
+what are the more relevant points from it, and a comment regarding further work
+from this paper. The general comments about the paper are structured as a five
+paragraph essay.
 
@@
 the references section. As some papers reference each other, these references
-are kept in the comments.
+are kept in the comments. The reason to identify a particular paper with a code
+is because it makes it easy to store them in disk due to their long names.

### 2017-01-11 papers.tex
Adds better build process
Adds bibliography

@@
 
-Paper is \TODO{cite}.
+Paper is \cite{ind_007}. Lists a lot of resources for other DM in \gls{lms}.
+
+\subsubsection{Comment}
 
@@
 points to useful resources.
+
+\subsubsection{Relevant Points}
+
+What is taken from this paper is the argument that ``(\dots) one of the most
+useful DM tasks in e-learning is classification''.
+
+\begin{itemize}
+    \item Discover potential student groups with similar characteristics and
+        reactions to a particular pedagogical strategy; \TODO{Yet to be
+        explore}
+    \item Detect students' misuse or game-playing; \TODO{Yet to be explore}
+    \item Group students who are hint-driven or failure-driven and find common
+        misconceptions that students possess; \TODO{Yet to be explore}
+    \item Identify learners with low motivation and find remedial actions to
+        lower drop-out rates; \TODO{Yet to be explore}
+    \item Predict/classify students when using intelligent tutoring systems;
+        \TODO{Yet to be explore}
+\end{itemize}
+
+In detail, there are methods that have been applied to detect students marks.
+
+\begin{itemize}
+    \item Predicting students' grades from test scores using neural networks;
+        \TODO{Yet to be explore}
+    \item Predicting student academic success (classes that are successful or
+        not) using discriminant function analysis; \TODO{Yet to be explore}
+    \item Classifying students using genetic algorithms to predict their final
+        grade; \TODO{Yet to be explore}
+    \item Predicting a student's academic success (to classify as low, medium
+        and high risk classes) using different data mining methods; \TODO{Yet
+        to be explore}
+    \item Predicting a student’s marks (pass and fail classes) using regression
+        techniques in Hellenic Open University data; \TODO{Yet to be explore}
+    \item Using neural network models from Moodle logs; \TODO{Yet to be
+        explore}
+\end{itemize}
+
+\subsubsection{Work From This Paper}
+
+From this paper we will analyse the listed techniques and methodologies.
+
+% =============================================================================
+\subsection{\texttt{IND\_008} - Discovering decision knowledge from web log
+portfolio for managing classroom processes by applying decision tree and data
+cube technology}
+
+\WORKON{Need to finish this paper's comment}
+
+Paper is~\cite{ind_008}. An experiment in which web logs of a \gls{lms} get
+analysed in order to find good student behaviour. This paper was found
+from~\cite{ind_007}.
+
+\subsubsection{Comment}
+
+\gls{lms} systems, like Moodle, generate logs which describe that activities of
+it's users. Those logs contain entries for when students login, when they ask
+questions, when they offer proposals on so on. Analysing these logs in order to
+understand if student performance is good or bad is not trivial. So
+some~\gls{dm} techniques are used to gain an understanding of student
+performance.
+
+The paper states that the data from the logs is initially transformed to be
+inserted into a relational database. SQL is then used to pre-process that data.
+Data cube technology is used for multidimensional data viewing. Decision trees
+are also constructed from the data.
+
+\subsubsection{Relevant Points}
+
+\TODO{Write this section}
+
+\subsubsection{Work From This Paper}
+
+\TODO{Write this section}

### 2017-01-11 papers.tex
Made article into a report

@@
-\section{Papers}
+\chapter{Papers}
 
 % =============================================================================
-\subsection{\texttt{IND\_007} - Data Mining Algorithms to Classify Students}
+\section{\texttt{IND\_007} - Data Mining Algorithms to Classify Students}
 
@@
 
-\subsubsection{Comment}
+\subsection{Comment}
 
@@
 
-\subsubsection{Relevant Points}
+\subsection{Relevant Points}
 
@@
 
-\subsubsection{Work From This Paper}
+\subsection{Work From This Paper}
 
@@
 % =============================================================================
-\subsection{\texttt{IND\_008} - Discovering decision knowledge from web log
+\section{\texttt{IND\_008} - Discovering decision knowledge from web log
 portfolio for managing classroom processes by applying decision tree and data

### 2017-01-11 papers.tex
Adds comment to ind_009

@@
 % =============================================================================
-\section{\texttt{IND\_008} - Discovering decision knowledge from web log
-portfolio for managing classroom processes by applying decision tree and data
-cube technology}
+\section{\texttt{IND\_008} - Discovering Decision Knowledge From Web Log
+Portfolio For Managing Classroom Processes By Applying Decision Tree And Data
+Cube Technology}
 
@@
 \TODO{Write this section}
+
+% =============================================================================
+\section{\texttt{IND\_009} - Predicting Performance From Test Scores Using
+Backpropagation and Counterpropagation}
+
+Paper is~\cite{ind_009}. Neural Networks are used to predict students grades
+from placement test responses. This paper was found from~\cite{ind_007}.
+
+\subsection{Comment}
+
+Backpropagation and Counterpropagation are two types of neural networks that
+are trained in different ways. This paper uses both types of networks to
+predict students grades. The data set used in this paper is small, but more
+data is generated from it. The data refers to placement tests for a course on
+Calculus I at West University Boulevard in Florida.
+
+The initial data set is small. It contains 116 entries. Each entry has 25
+features, which although not specified in the paper, are stated to be results
+from placement tests. Each entry also contains the final grade of the student,
+from \textit{A} to \textit{D} stating a different level of approval, and
+\textit{F} when a student failed. Because the data is small, some data has been
+generated.
+
+The networks contains 25 inputs, for the features, and 5 outputs, for each
+grade. An output of \textit{A} will be a vector of the form $ [1, -1, -1, -1,
+-1] $, for \textit{B} it will be $ [-1, 1, -1, -1, -1] $, and so on. The
+networks contain a single hidden layer which number of units varies for
+different trainings. The number of units are 15, 20, 25, and 30 for different
+tests.
+
+It is shown how the networks are trained and are able to achieve correctly
+classify students, given their input data. It seems that an optimal number of
+units in the hidden layer is 25. Some networks achieve 80\% of correctly
+classified instances for a pure dataset. Training networks with generated
+datasets is yields greater correctly classified percentages between 90\% and
+100\%.
+
+Although the results shown are very good, the paper doesn't go into detail
+about the inputs of the networks. It is important to know to format of the data
+of the students placement tests, how the data was pre-processed, and into what
+was it pre-processed. Like this, the paper leaves to much to speculation. It
+would also be nice to see how these methods are applied in other courses.
+
+The paper presents the hypothesis that neural networks are good for predicting
+student's performance based on existing placement tests for a given course on
+Calculus I. It then shows how data is taken and used to generate more data. The
+networks are trained and are able to achieve great results.
+
+\subsection{Relevant Points}
+
+This paper is important because it is prof that neural networks can be used to
+classify student performance. It is also important because it shows a useful
+and successful tasks in the world of~\gls{edm}.
+
+Ideas from this paper may be used for later study.
+
+\subsection{Work From This Paper}
+
+This paper only has four references, and all of them are related with Neural
+Networks. Those references might came in handy if work is to be done in Neural
+Networks, but not for data mining in~\gls{lms}.

### 2017-01-11 introduction.tex
Made article into a report
Adds details to introduction

@@
-\section{Introduction}
+\chapter{Introduction}
 
@@
 
+The five paragraph essay, despite it's name, doesn't necessarily has five
+literal paragraphs. It does however follow a structure of the form:
+
+\begin{itemize}
+    \item Introduction
+    \item Exposure
+    \item Arguments for
+    \item Arguments against
+    \item Conclusions
+\end{itemize}
+
 Each paper is identified by the code \texttt{IND\_XXX}, where \texttt{XXX} is a
@@
 is because it makes it easy to store them in disk due to their long names.
+
+Note that the number in \texttt{IND\_XXX} doesn't imply an order in the papers
+listed in the document. Also, just because there is a paper in
+\texttt{IND\_904} and a paper \texttt{IND\_906}, that doesn't imply that a
+paper \texttt{IND\_905} needs to exists in the document.

### 2017-01-16 papers.tex
Adds another comment

@@
 
-Paper is \cite{ind_007}. Lists a lot of resources for other DM in \gls{lms}.
+Paper is~\cite{ind_007}. Lists a lot of resources for \gls{dm} in \gls{lms}.
 
@@
 What is taken from this paper is the argument that ``(\dots) one of the most
-useful DM tasks in e-learning is classification''.
+useful~\gls{dm} tasks in e-learning is classification''. The paper then lists a
+few sources where~\gls{edm} has been used. They are the following:
 
@@
     \item Discover potential student groups with similar characteristics and
-        reactions to a particular pedagogical strategy; \TODO{Yet to be
-        explore}
+        reactions to a particular pedagogical strategy;~\cite{ind_008}
     \item Detect students' misuse or game-playing; \TODO{Yet to be explore}
@@
 \begin{itemize}
-    \item Predicting students' grades from test scores using neural networks;
-        \TODO{Yet to be explore}
+    \item Predicting students' grades from test scores using neural
+        networks;~\cite{ind_009}
     \item Predicting student academic success (classes that are successful or
@@
 
-\WORKON{Need to finish this paper's comment}
-
 Paper is~\cite{ind_008}. An experiment in which web logs of a \gls{lms} get
@@
 questions, when they offer proposals on so on. Analysing these logs in order to
-understand if student performance is good or bad is not trivial. So
-some~\gls{dm} techniques are used to gain an understanding of student
-performance.
+understand if student performance is good or bad is not trivial because they
+were not made for pedagogical purposes. So some~\gls{dm} techniques are used to
+gain an understanding of student performance.
 
@@
 Data cube technology is used for multidimensional data viewing. Decision trees
-are also constructed from the data.
+are also constructed from the data. These decision trees constructed from the
+web logs are the final results which shed light on student behaviour.
+
+The logs in questions are made in a system that implements IBIS. The logs are
+processed and inserted into data cubes for later analysis. The aim of the
+analysis is to check if certain pedagogical strategies work. The analysis later
+use decision trees.
+
+The paper is too long and repeats a few sections. It is not that clear how the
+logs are processed or how SQL and data cubes are used for analysus. It is also
+not clear how decision trees are built or what they represent.
+
+This is a useful paper with some different methods worth reading about, even if
+data cube isn't that common in the rest of the literature explored until this
+point.
 
@@
 
-\TODO{Write this section}
+This paper is useful because it is an example of mining web logs for~\gls{edm}.
+Also useful for it shows the usage of data cubes, which isn't that common.
 
@@
 
-\TODO{Write this section}
+No further work is done for this paper.
 
@@
 
-Ideas from this paper may be used for later study.
+Ideas from this paper may be used for later study, namely how data is divided
+into fold for training and testing, how resampling is done to generate more
+data, and how it also influences the correctness of the trained models.
 
@@
 Networks, but not for data mining in~\gls{lms}.
+
+% =============================================================================
+\section{\texttt{IND\_003} - Web Usage Mining for Predicting Final Marks of
+Students That Use Moodle Courses}
+
+Paper is~\cite{ind_003}. Paper~\cite{ind_007} is related to this one, for the
+authors and experiments in this paper are the same as in~\cite{ind_007}, but
+are presented with much more detail.
+
+\subsection{Comment}
+
+This paper mines web logs of~\gls{lms} in order to take meaning from them. Like
+in~\cite{ind_007} and~\cite{ind_008}, it is asserted that extracting meaningful
+information from logs is not trivial, so~\gls{dm} methodologies are used. The
+paper makes it so, while presenting the various algorithms testing with
+different dataset, having some of which being the results of resampling.
+
+The paper shows how data from seven courses is taken and mined. That data
+consists of web logs detailing the activities of students in Moodle. Those
+activities are, for example, when a student logged in, when it answered a
+question, when it saw a forum post, etc. Several algorithms are executed over
+this data in order to classify students. Further pre-processing tasks are
+executed over the same data and differences in accuracy are observed.
+
+We are able to see the results of various algorithms executed with different
+datasets. It is possible to observe how different pre-processing methodologies
+influence the final results. The paper presents several ways to pre-process
+data, and many algorithms.
+
+Looking at all of the results, we see that no algorithm got a better ``Global
+Percentage of Correctly Classified'' percentage than 66\%. This level of
+accuracy leaves much to be desired. The geometric mean is used to evaluate the
+accuracy of algorithms with imbalanced data. We see that many algorithms get
+0\% in this indicator, while other got percentages below 50\% and close to 0\%.
+This is very bad.
+
+Decision trees are concluded to be the best model for they perform better in
+term of ``Global Percentage of Correctly Classified'' and in geometric mean.
+These models are best executed with the presented methods for pre-processing.
+The final values are still considered to be low, but the results make sense.
+
+\subsection{Relevant Points}
+
+Again, we see how using data mining on~\gls{lms} logs is useful. We also see
+how classification is useful.
+
+This paper presents a few tools which will be extremely important to
+pre-process data in order to prepare it for the algorithms. It is intend that
+decision trees are a good model to classify students into different groups
+given their activity in Moodle.
+
+\subsection{Work From This Paper}
+
+Find more about Moodle Data Mining Tool and find more work done by these
+authors.
+
+Look further into using decision trees to mine Moodle data.

### 2017-01-16 papers.tex
Adds details to ind_003

@@
 
+\subsection{Reading Notes}
+
+\Gls{lms} are being widely used. Moodle is very famous.
+
+\Gls{lms} store a lot of data regarding student profiles, their grades, when
+they login, etc. Analysing this data is hard to do by hand. Data Mining
+techniques may be used.
+
+Data Mining has been applied to a variety of problems in e-learning. Like:
+
+\begin{itemize}
+    \item Dealing with assessment of students' learning performance.
+    \item Providing course adaptations and learning recommendations.
+    \item etc. See more at 17
+\end{itemize}
+
+The objective of this paper is to classify students with similar final marks
+into different groups depending on the activities carried out in a web-based
+course.
+
+Data from Cordoba University is used in 7 Moodle engineering courses. The data
+comes from three online activities:
+
+\begin{itemize}
+    \item Quizzes
+    \item Assignments
+    \item Forums
+\end{itemize}
+
+The data was divided into 10 pairs of training and test data files for
+stratified 10-fold cross-validation (reference 81).
+
+The first experiment consists on executing every algorithm with three datasets
+and getting the geometric mean:
+
+\begin{itemize}
+    \item First dataset. All 438 instances/students with all features.
+    \item Second dataset. Filtered instances where students didn't complete
+        every Moodle activity. Got 135 instances.
+    \item Third dataset. All instance but only with the 4 features which were
+        selected using a few methods.
+\end{itemize}
+
+The second experiment is basically the same as the first, but the data ha two
+pre-processing tasks applied to it: discretization and rebalancing.
+
+Paper also gives some background on data mining. Also talks about Moodle Data
+Mining Tool.
+
 \subsection{Comment}

### 2017-01-16 papers.tex
Adds new paper

@@
 Look further into using decision trees to mine Moodle data.
+
+% =============================================================================
+\section{\texttt{IND\_010} - Using Educational Data Mining Methods to Study the
+Impact of Virtual Classroom in E-Learning}
+
+Paper is~\cite{ind_010}. I haven't seen this paper has a reference of the
+previous ones. But found it on Google Scholar while searching for Moodle Data
+Mining. The techniques used in this paper are similar to~\cite{ind_008,
+ind_003}.
+
+\subsection{Reading Notes}
+
+The paper starts by talking about difficulties in e-learning environments. Such
+as students feeling lonely and unsupported. They also talk about how web logs
+need to be mined so meaningful information is taken from them.
+
+This study is done to web logs. They have 824 students in 11 courses, but claim
+that the number is superior. The study is done only to students who are remote.
+They say that other studies such as 8, 13, and 14 e-learning platforms was used
+to facilitate teacher-student interactions.
+
+They start by preprocessing a group of features of students. Then they do
+feature ranking using a few algorithms and build a decision tree based on the
+Gain Ratio metric.
+
+The decision tree predicts students grades based on their features.
+
+\subsection{Comment}
+
+This paper is another~\gls{dm} analysis made to Moodle. Like in previous works,
+such as~\cite{ind_008, ind_003}, this paper goes through the process of
+pre-processing Moodle data from e-learning courses, applying~\gls{dm}
+techniques, and interpreting the results. The objectives of this paper are to
+predict student grades given their usage of Moodle in an e-learning
+environment.
+
+The origins of the data are shown to come from 824 students in 11 courses. The
+features of the data come from logs which say when a student logged in, when it
+asked a question, and so on. The data mining process consists on applying a few
+feature selection algorithms, such as gain ratio, $ \chi^2 $, SVN, and others,
+to rank the features. In the end, a decision tree is built from the ranking
+using gain ratio.
+
+The built tree is able to predict the grades of a student given their features.
+The first two levels of a built tree are shown and some interpretation is
+given. For example, assertions are made about students that participate less
+then 11 times in virtual classrooms which will probably fail the exams.
+
+Although the results are presented and seem logical, the probabilities given
+for each assertion of the decision tree are made so with small probabilities.
+The previous example, for instance, only has 55\% probability. Some ways to
+evaluate the accuracy of the built tree should be taken.
+
+The methods listed in this paper are well explained and the results are clear.
+Building decisions trees to predict students grades is a sound objective and a
+common one in~\gls{edm}. The only negative points are the fact that the built
+tree is not evaluated.
+
+\subsection{Relevant Points}
+
+This paper is very relevant because it displays similar techniques
+to~\cite{ind_008, ind_003}. The techniques for pre-processing and mining, the
+datasets, and the conclusions are all very similar and will doubtless be very
+important as a reference in future work.
+
+\subsection{Work From This Paper}
+
+References 18, 16, 5, 8, are 9 worth looking into.

### 2017-01-16 papers.tex
Adds another papers

@@
         reactions to a particular pedagogical strategy;~\cite{ind_008}
-    \item Detect students' misuse or game-playing; \TODO{Yet to be explore}
+    \item Detect students' misuse or game-playing; (2)
     \item Group students who are hint-driven or failure-driven and find common
-        misconceptions that students possess; \TODO{Yet to be explore}
+        misconceptions that students possess; (34)
     \item Identify learners with low motivation and find remedial actions to
-        lower drop-out rates; \TODO{Yet to be explore}
+        lower drop-out rates; (9)
     \item Predict/classify students when using intelligent tutoring systems;
-        \TODO{Yet to be explore}
+        (16)
 \end{itemize}
@@
     \item Predicting student academic success (classes that are successful or
-        not) using discriminant function analysis; \TODO{Yet to be explore}
+        not) using discriminant function analysis; (21)
     \item Classifying students using genetic algorithms to predict their final
-        grade; \TODO{Yet to be explore}
+        grade; (30)
     \item Predicting a student's academic success (to classify as low, medium
-        and high risk classes) using different data mining methods; \TODO{Yet
-        to be explore}
+        and high risk classes) using different data mining methods; (30)
     \item Predicting a student’s marks (pass and fail classes) using regression
-        techniques in Hellenic Open University data; \TODO{Yet to be explore}
-    \item Using neural network models from Moodle logs; \TODO{Yet to be
-        explore}
+        techniques in Hellenic Open University data; (18)
+    \item Using neural network models from Moodle logs; (11)
 \end{itemize}
@@
 References 18, 16, 5, 8, are 9 worth looking into.
+
+% =============================================================================
+\section{\texttt{IND\_011} - Predicting students' marks from Moodle logs using
+neural network models}
+
+Paper is~\cite{ind_011}. But found it on Google Scholar while searching for
+Moodle Data Mining. This is another paper that uses neural networks
+like~\cite{ind_009} did.
+
+\subsection{Comment}
+
+Another paper that attempts to predict student's success given their usage of
+Moodle. In this case we see the usage of neural networks, as we've seen
+in~\cite{ind_009}. Like in~\cite{ind_007, ind_008, ind_003, ind_010}, Moodle
+logs are processed alongside final grades in order to build features for
+students.  Neural networks are then trained using that data.
+
+The usual information is taken from logs. For example, date and time of access
+of Moodle, if a user saw an item, if it wrote somethings in a forum etc. This
+data is pre-processed in order to build a dataset. The built dataset for these
+experiments is made from entries of 240 students. The trained neural networks
+are Radial Basis Functions. They are trained in a few different ways.
+
+It is shown how training a neural network with an incremental hidden layer
+algorithm yields a model capable of achieving 80\% of correctly classified
+instances. Other ways to train the networks are shown to be very successful
+achieving percentages above 70\%.
+
+Not much is known about how the success metric is calculated, or how the
+alternative ways to train the network actually work. But they are listed
+nonetheless. The paper leaves a few questions opened, namely the usefulness of
+the actual study.
+
+Neural networks are presented as a good model to predict students success, and
+are shown to be, in fact, a great model for it. Different ways to train a
+network are shown and some hints as to how data is pre-processed into useful
+datasets are also shown. Even if the paper is short, doesn't go into detail
+about certain aspects of the project, and leaves things opened, it is still a
+great resource.
+
+\subsection{Relevant Points}
+
+Networks networks are once again shown to work in predicting student's success
+given their Moodle usage habits.
+
+\subsection{Work From This Paper}
+
+See reference 7 in this paper about Radial Basis Functions.

### 2017-01-19 papers.tex
Changed paper's index

@@
 % =============================================================================
-\section{\texttt{IND\_007} - Data Mining Algorithms to Classify Students}
+\section{\texttt{IND\_001} - Data Mining Algorithms to Classify Students}
 
-Paper is~\cite{ind_007}. Lists a lot of resources for \gls{dm} in \gls{lms}.
+Paper is~\cite{ind_001}. Lists a lot of resources for \gls{dm} in \gls{lms}.
 
@@
     \item Discover potential student groups with similar characteristics and
-        reactions to a particular pedagogical strategy;~\cite{ind_008}
+        reactions to a particular pedagogical strategy;~\cite{ind_002}
     \item Detect students' misuse or game-playing; (2)
@@
     \item Predicting students' grades from test scores using neural
-        networks;~\cite{ind_009}
+        networks;~\cite{ind_003}
     \item Predicting student academic success (classes that are successful or
@@
 % =============================================================================
-\section{\texttt{IND\_008} - Discovering Decision Knowledge From Web Log
+\section{\texttt{IND\_002} - Discovering Decision Knowledge From Web Log
 Portfolio For Managing Classroom Processes By Applying Decision Tree And Data
@@
 
-Paper is~\cite{ind_008}. An experiment in which web logs of a \gls{lms} get
+Paper is~\cite{ind_002}. An experiment in which web logs of a \gls{lms} get
 analysed in order to find good student behaviour. This paper was found
-from~\cite{ind_007}.
+from~\cite{ind_001}.
 
@@
 % =============================================================================
-\section{\texttt{IND\_009} - Predicting Performance From Test Scores Using
+\section{\texttt{IND\_003} - Predicting Performance From Test Scores Using
 Backpropagation and Counterpropagation}
 
-Paper is~\cite{ind_009}. Neural Networks are used to predict students grades
-from placement test responses. This paper was found from~\cite{ind_007}.
+Paper is~\cite{ind_003}. Neural Networks are used to predict students grades
+from placement test responses. This paper was found from~\cite{ind_001}.
 
@@
 % =============================================================================
-\section{\texttt{IND\_003} - Web Usage Mining for Predicting Final Marks of
+\section{\texttt{IND\_004} - Web Usage Mining for Predicting Final Marks of
 Students That Use Moodle Courses}
 
-Paper is~\cite{ind_003}. Paper~\cite{ind_007} is related to this one, for the
-authors and experiments in this paper are the same as in~\cite{ind_007}, but
+Paper is~\cite{ind_004}. Paper~\cite{ind_001} is related to this one, for the
+authors and experiments in this paper are the same as in~\cite{ind_001}, but
 are presented with much more detail.
@@
 This paper mines web logs of~\gls{lms} in order to take meaning from them. Like
-in~\cite{ind_007} and~\cite{ind_008}, it is asserted that extracting meaningful
+in~\cite{ind_001} and~\cite{ind_002}, it is asserted that extracting meaningful
 information from logs is not trivial, so~\gls{dm} methodologies are used. The
@@
 % =============================================================================
-\section{\texttt{IND\_010} - Using Educational Data Mining Methods to Study the
+\section{\texttt{IND\_005} - Using Educational Data Mining Methods to Study the
 Impact of Virtual Classroom in E-Learning}
 
-Paper is~\cite{ind_010}. I haven't seen this paper has a reference of the
+Paper is~\cite{ind_005}. I haven't seen this paper has a reference of the
 previous ones. But found it on Google Scholar while searching for Moodle Data
-Mining. The techniques used in this paper are similar to~\cite{ind_008,
-ind_003}.
+Mining. The techniques used in this paper are similar to~\cite{ind_002,
+ind_004}.
 
@@
 This paper is another~\gls{dm} analysis made to Moodle. Like in previous works,
-such as~\cite{ind_008, ind_003}, this paper goes through the process of
+such as~\cite{ind_002, ind_004}, this paper goes through the process of
 pre-processing Moodle data from e-learning courses, applying~\gls{dm}
@@
 This paper is very relevant because it displays similar techniques
-to~\cite{ind_008, ind_003}. The techniques for pre-processing and mining, the
+to~\cite{ind_002, ind_004}. The techniques for pre-processing and mining, the
 datasets, and the conclusions are all very similar and will doubtless be very
@@
 % =============================================================================
-\section{\texttt{IND\_011} - Predicting students' marks from Moodle logs using
+\section{\texttt{IND\_006} - Predicting students' marks from Moodle logs using
 neural network models}
 
-Paper is~\cite{ind_011}. But found it on Google Scholar while searching for
+Paper is~\cite{ind_006}. But found it on Google Scholar while searching for
 Moodle Data Mining. This is another paper that uses neural networks
-like~\cite{ind_009} did.
+like~\cite{ind_003} did.
 
@@
 Moodle. In this case we see the usage of neural networks, as we've seen
-in~\cite{ind_009}. Like in~\cite{ind_007, ind_008, ind_003, ind_010}, Moodle
+in~\cite{ind_003}. Like in~\cite{ind_001, ind_002, ind_004, ind_005}, Moodle
 logs are processed alongside final grades in order to build features for

### 2017-01-26 papers.tex
Adds new paper

@@
 See reference 7 in this paper about Radial Basis Functions.
+
+% =============================================================================
+\section{\texttt{IND\_007} - Educational data mining: A survey from 1995 to
+2005}
+
+Paper is~\cite{ind_007}. This is an overview of~\gls{edm}.
+
+Starts by comparing e-learning Data Mining to e-commerce. States that the
+differences are as follows:
+
+\begin{itemize}
+    \item Domain: E-commerce's purpose is to guide clients in purchasing.
+        E-learning's purpose is to guide students in learning.
+    \item Data: E-learning has more data then logs, because we have student's
+        information.
+    \item Objective: The objective of e-commerce is to maximise profit. In
+        e-learning is to improve learning, which is more subjective.
+    \item Technique: Techniques are different.
+\end{itemize}
+
+EDM may be oriented towards different actors who take different points of view
+and interest from the analysis. Those actors are:
+
+\begin{itemize}
+    \item Students, who want to improve their studying method.
+    \item Educators, who want feedback on their education methods, and who want
+        to classify students or group their needs and other features.
+    \item Oriented towards academics responsible and administrators, who want
+        to optimize site performance, among other indicators.
+\end{itemize}
+
+The paper lists types of e-learning systems~(\gls{lms}) and how their logs are
+made. Also states a few limitations on those logs.
+
+The paper lists data pre-processing tasks and techniques that will also be
+useful.
+
+% =============================================================================
+\section{\texttt{IND\_008} - Educational Data Mining: A Review of the State of
+the Art}
+
+Paper is~\cite{ind_008}.
+
+This paper is a very useful overview of the~\gls{edm} field. It goes on to
+explain several points about this field.
+
+The paper asserts that instrumental educational software and databases
+containing student information provide big repositories of data which can be
+studied to obtain meaning from them. Those studies are made using Statistical
+Analysis and Data Mining methods.
+
+EDM has emerged from the need to implement Data Mining methods over these
+repositories of data. \gls{edm} as now been subdivided into three areas:
+
+\begin{itemize}
+    \item Offline education
+    \item E-Learning and~\gls{lms}
+    \item Intelligent tutoring systems (ITS) and Adaptive Educational
+        Hypermedia Systems (AEHS)
+\end{itemize}
+
+An important point is that some pre-processing method and some post-processing
+methods for data visualization are not necessarily~\gls{dm} topics, but are
+closely related, therefore, they are mentioned in this paper. The mentioned
+techniques for pre and post-processing are relevant in the~\gls{edm} field.
+
+The following is a list of Stakeholders of~\gls{edm} projects. It is very
+important that this is listed in the final thesis text because it provides the
+whole motivation for having done a project like this in the first place.
+
+\begin{itemize}
+    \item Learners/Students
+    \item Educators/Teachers
+    \item Course Developers/Educational Researchers
+    \item Learning Providers/Universities
+    \item School Administrators and similar
+\end{itemize}
+
+For Learners. The results of can be used to recommend activities to students
+and suggest interesting learning experiments which are alternative to the
+experiences of a particular student. These results can also be used to
+recommend interesting courses or point to useful discussions.
+
+For Educators. The results may be used to get objective feedback about
+instruction and analyse students. It becomes possible to do the following:
+
+\begin{itemize}
+    \item Find common mistakes in students
+    \item Predict student grades based on student activities
+    \item Determine effective activities
+\end{itemize}
+
+For Course Developers. Evaluating the effectiveness of a particular course, how
+it is structured, what activities are the most useful and effective, are
+important points to take into account when making changes to a course.
+
+For Learning Providers. These are the entities that actually provide the means
+to implement learning. These are, for example, Universities, both public and
+private, and similar learning institutions.~\gls{edm} can give a powerful
+overview of what is the most cost-effective way of assuring good grades,
+retention of student, and addition of students who will do well in university.
+
+For School Administrators.~\gls{edm} can give insight on the best ways to
+organize human and material resources. In the context of this
+project,~\gls{edm} is useful to evaluate the effectiveness of~\gls{edm}.
+
+The paper states that publications in~\gls{edm} have grown a lot and gives some
+number for it. There are also a list of publication that are considered to be
+extremely relevant.

### 2017-02-10 data.tex
Adds data and experiments parts

\chapter{Data}

% =============================================================================
\section{Data Exploration}

To begin with, we explore what was initially provided in terms of data. The
repository is a spreadsheet book with 12 sheets. The sheets contain raw data
and some calculated fields.

In order to explore the data using the bash tools (cut, grep, etc) and tools
like Python with Pandas, they were exported to a CSV format. Some
transformations may be applied to these files in other to clean useless points
or make the data easier to process.

To begin, each sheet is analysed individually. Conclusions on each list are
registered in the fist section of this chapter.

\subsection{Sheet 1 - \texttt{UCs\_SIIUE\_Moodle.csv}}

This sheet lists a few courses (referred to in their native name as
\textit{Unidades Curriculares}). Some courses have a Moodle page while others
do not.

The sheets contain four non blank columns. The first and second column are a
list of courses represented has their Course Id and Moodle Id, which is
different.

The third column is just a calculated copy of the first. It doesn't add
anything and it might have been a loose processing step at some point.

The fourth column contains a list of course with no Moodle and therefore no
Moodle Id.

Only the following columns are relevant:

\begin{itemize}
    \item \texttt{Cod UC} - Course's code
    \item \texttt{No moodle} - Course's code in the Moodle
\end{itemize}

\subsection{Sheet 2 - \texttt{complete\_rate.csv}}

This sheet contain a listing of the completion rate of each course. The list
has tree columns. A column for course id in the Moodle (\texttt{c\_moodleid}),
the number of students in that course (\texttt{n\_students}), and the
completion rate of those courses (\texttt{complete\_rate}).

So a course with 100 students having a completion rate of 75\% means that 75
students actually completed that course.

\subsection{Sheet 3 - \texttt{completed.csv}}

Contains a listing of which student was approved at which course. The sheet has
three columns. The first column is the course id of the Moodle
(\texttt{c\_moodleid}), the second is the user id of the Moodle
(\texttt{u\_moodleid}), and the third is a column (\texttt{completed}) which
has the value 1 if the student completed that course, or 0 if not.

The user id may be assigned to a professor, but in this sheet every user id is
of a student.

\subsection{Sheet 4 - \texttt{Notas UCS SIIUE.csv}}

There is a lot of information in this sheet.

\TODO{Explore}

\subsection{Sheet 5 - \texttt{Profile}}

This sheet contains some analysis made to the data. To begin with, there are
the following columns:

\begin{itemize}
    \item \texttt{name\_c} - Name of the course
    \item \texttt{name\_s} - Name of the student
    \item \texttt{id\_c} - Id of the course
    \item \texttt{n\_resrc\_c} - Number of resources in the course
    \item \texttt{n\_actv\_c} - Number of activities in the course
    \item \texttt{id\_s} - Id of the student
    \item \texttt{n\_cons\_s} - \TODO{}
    \item \texttt{n\_subm\_s} - \TODO{}
    \item \texttt{result\_s} - Results of that student in that course
\end{itemize}

In order for a student to get approved, some creteria must be followed.

\subsection{Sheet 6 - \texttt{ResultadosUCMoodle}}

This sheet contains a lot of information about the courses. To begin with, we
get a relation between the three ways of identifying a course in the data. A
course is identified by the Course's Code (\texttt{Cod Siiue}), Course's Code
in the Moodle (\texttt{Cod Moodle}), or by a course id (\texttt{course}). The
course id is an integer number, while the codes contain information about the
department.

The department (\texttt{Dep}) is calculate from the course's code (\texttt{Cod
Siiue}) by taking the first three characters from that course. So a course with
code \texttt{INF7185\_1} will be from the department \texttt{INF}.

The sheet presents the first and last register of an activity in the Moodle,
and also the interval in days from the first activity to the last. The fields
are \texttt{min}, \texttt{max}, and \texttt{int}.

\TODO{Fields in between}

The sheet finally specifies the number of students who enrolled in a course and
the number of students who were approved on that course.

\begin{itemize}
    \item \texttt{course} - Course's id
    \item \texttt{Cod Moodle} - Course's Code in the Moodle
    \item \texttt{Cod Siiue} - Course's Code
    \item \texttt{min} - First activity of course
    \item \texttt{max} - Last activity of course
    \item \texttt{int} - Interval between first and last activity of course
    \item \texttt{actsub}
    \item \texttt{actcons}
    \item \texttt{nactsubd}
    \item \texttt{consal}
    \item \texttt{consdoc}
    \item \texttt{subal}
    \item \texttt{subdoc}
    \item \texttt{nalu}
    \item \texttt{ndoc}
    \item \texttt{InscSIIUE} - Number of students who enrolled
    \item \texttt{Aprov} - Number of students who got approved
    \item \texttt{Dep} - Department of the course
\end{itemize}

\subsection{Sheet 7 - \texttt{Correlação}}

This is a sheet that implements the calculation of correlation between the
variables of the previous sheet \texttt{ResultadosUCMoodle}. The data it self
is coppied from the previous sheet and a new table to the right of the first
table is added with the correlation itself.

The data in here doesn't need to be preprocessed to a new sheet, but the notion
of correlation, which can be calculated using alternative methods, may be used
later.

\subsection{Sheet 8 - \texttt{Alunos SIIUE E-L}}

There are a few issues with this sheet. First of all, the sheet appears to have
two tables which are not really related, maybe? Anyway. The first table has a
student number, his name, and what appears to be a count of it's enrollments.
It also has a calculated fields stating his/her's first and last name and
another column with the students number which is a calculated copy of the
first.

The second table has Moodle usernames. A first and last name, another count
(which I assume is a count of enrollments but I can't be sure because the field
simply states ``count'').

Oh! The second number in the first table is used for a calculated value in the
second table.

So the last three columns of the second table are all calculated values. The
first appears to fetch a number from the first. It goes like this. If the name
of the student in the second table is equal to the first and last name of the
first, then fetch that number. Else, it is an error.

Next, we get a value, is the username in the Moodle contains a number. If not,
we get an error.

Lastly. We try to get the first calculated number (the fourth column of the
second table) if it exists. If it doesn't, we try to get the second calculated
value (the fifth column of the second table). If that doesn't exist too, we
write 0.

Up to this point I'm not sure what the purpose of this sheet is. I believe it
gives me a relation between a student in SIIUE and in Moodle, but I have
serious doubts about the correctiveness of that relation.

After examining the sheet a second, I now believe that the first table has a
listing o students and their SIIUE internal id or student number. The second
table simply has usernames of Moodle. So the second table has students and
professors, while the first one has students only.

About the counts. The count of the first table is indeed a number of courses
where the students is enrolled in. The count of the second table is the total
number of log entries which is a copy (a copy-paste values only copy) of the
information in the \texttt{AlunosMoodle} sheet.

I don't think this sheet alone is much useful. But it did give insight on how
this data repository is organized.

\subsection{Sheet 9 - \texttt{AlunosMoodle}}

These are the actual logs of the accesses to Moodle. This is a big table that
contains the following fields:

\begin{itemize}
    \item \texttt{c\_id} - Course ID in the Moodle
    \item \texttt{c\_cod} - Course ID
    \item \texttt{c\_role} - Role of the user
    \item \texttt{c\_user} - User's code
    \item \texttt{c\_nome} - User's name
    \item \texttt{c\_uid} - User's ID in the Moodle
    \item \texttt{crud} - \gls{crud} activity
    \item \texttt{semana} - Number of the week in which the events happened
    \item \texttt{n} - Number of time the event happened on that week
    \item \texttt{numAluno} - Number of the student, which is calculated from
        elsewhere, even when the user is not a student
    \item \texttt{SNot} - If the student was approved, or failed, or something
        else
\end{itemize}

Basically, we get a list of entries specifying activities in Moodle. Each entry
contains information on the course, the user, and the activity itself.

For the course information we have course id and course code (Moodle code).

For user, we get the user's role, username, actual name of user, and some
number.

Finally, we get what kind of CRUD activity it was (if create activity, delete,
etc), the week in which said activity occurred, and number of times the
activity occurred in that week. The activities are aggregated by type and week.

This table then contains the fields \texttt{numAluno} and \texttt{SNot}. These
are calculate fields. The first attempts to take a student's number from the
username, if possible. The second calculates if that particular student was
approved in the course for which the entries refers to. Both of these fields
are useless.

There is a second table. This table gives a username, actual name, and total
number of activities of that user in the log. It is useful, but it should be in
it's own sheet. The fields of that table have no name, but in order to be
consistent with the table on the left, we get:

\begin{itemize}
    \item \texttt{c\_user}
    \item \texttt{c\_nome}
    \item \texttt{count}
\end{itemize}

\subsection{Sheet 10 - \texttt{CalculosAlunos}}

I'm not sure what his sheet is. It has a few values copied from other tables,
has ``\textit{Calculos}'', on the name, but it doesn't really calculate
anything. All the values are literal.

It may be discarded.

\subsection{Sheet 11 - \texttt{export\_weekly}}

This table seems to have a line for each pain of course and user ids. It then
has a few columns of log activities. For each intersection of pair of course
and user ids we get a 1 or 0 for one of the columns. Presumably, we get a true,
that is, 1, if a user A made that activity in course B in that week.

This isn't really useful because the same can be easily calculated using other
tools when needed.

\subsection{Sheet 12 - \texttt{export\_activity}}

This table seems to have a few more calculations, but I don't fully understand
them. It has at least a calculation of a standard deviation, and little else.

Again, this isn't really useful because the same can be easily calculated using
other tools when needed.

% =============================================================================
\section{Objects, Entities, and Features}

After a first analysis of the data, the objects in the data are identified.
These objects are the courses, students, results, etc. Data mining tasks are
applied over these objects which are described by their features.

This section specifies objects in this data repository and states their
features.

\subsection{Courses}

In the data, a course is identified by one of three unique ids. They are, the
Course ID, SIIUE's Course's Code, and Moodle Course's Code. The Course id is
simply a unique positive integer number which is probably assigned as an
incremental value when the courses where created in some database.

The two course's code come from their register in the SIIUE system and in the
Moodle system. For the most part they are the same, when they are not the same
the differences are minimal.

Table~\ref{tab:course_code_example} shows an example of the three ids. Notice how \texttt{CourseCodeMoodle} is either the same as \texttt{CourseCodeSiiue} or has just minimal differences (in this case there is an underscore follwed by a number).

\begin{table}[h!]
    \centering

    \begin{tabular}{l l l}
\texttt{CourseId} & \texttt{CourseCodeSiiue} & \texttt{CourseCodeMoodle} \\ \hline
        496       & \texttt{ARC10548}        & \texttt{ARC10548}         \\
        1392      & \texttt{GES7182}         & \texttt{GES7182\_2}       \\
        79        & \texttt{GES7182}         & \texttt{GES7182\_4}       \\
        1535      & \texttt{GES7182}         & \texttt{GES7182\_3}       \\
    \end{tabular}

    \caption
        [Example of course id and codes in the \texttt{CoursesGeneral} Dataset]
        {Example of course id and codes in the \texttt{CoursesGeneral} Dataset.}

    \label{tab:course_code_example}
\end{table}

Each course has the features listed in
table~\ref{tab:courses_general_features}. The table displayed the features and
their data type. Each course has a name, a department to which they belong, a
cycle (if Master's Degree, Phd, etc), and a degree. These features are all
strings.

A course also has a regime, a semester, a season, and a type. These features
are enumerated types because they have a limited domain. For example, the
semester may either be ``\textit{Par}'' or ``\textit{Ímpar}'' (Even or Odd),
the season may be ``\textit{Normal}'' or ``\textit{Special}''.

Finally, the credits are the only field which is a number.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Feature             & Data Type  \\ \hline
        \texttt{Department} & String     \\
        \texttt{Cycle}      & String     \\
        \texttt{CourseName} & String     \\
        \texttt{Regime}     & Enumerated \\
        \texttt{Credits}    & Number     \\
        \texttt{Degree}     & String     \\
        \texttt{Semestre}   & Enumerated \\
        \texttt{Season}     & Enumerated \\
        \texttt{Type}       & Enumerated \\
    \end{tabular}

    \caption
        [Data types of features of Courses]
        {Data types of features of Courses as they appear in the
        \texttt{CoursesGeneral} dataset.}

    \label{tab:courses_general_features}
\end{table}

Some courses have no Moodle. They are therefore useless for our these
experiments.

Some statistics are taken from the courses data as shown in
table~\ref{tab:courses_stats}.

The domain of the enumerated fields are listed in table~\ref{tab:courses_enum}.
The field ``type'' only has the domain ``\textit{Normal}'', so it is a field
which is useless for further analysis. The field ``Semester'' has a missing
values. In the actual data there is only one course which doesn't have a value
for this field.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Number of Courses     & 77 \\ \hline
        Number of Departments & 9  \\ \hline
        Number of Degrees     & 11 \\ \hline
    \end{tabular}

    \caption
        [Statistics on the Courses]
        {Statistics on the Courses.}

    \label{tab:courses_stats}
\end{table}

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l |}
        \hline
        Regime   & \{ A, O, S \}                                    \\ \hline
        Semester & \{ ``\textit{Par}'', ``\textit{Ímpar}'', n/a \}  \\ \hline
        Season   & \{ ``\textit{Normal}'', ``\textit{Especial}'' \} \\ \hline
        Type     & \{ ``\textit{Normal}'' \}                        \\ \hline
    \end{tabular}

    \caption
        [Values for enumerated types on courses]
        {Values for enumerated types on courses.}

    \label{tab:courses_enum}
\end{table}

\subsection{Moodle Users - Students, and Professors}

The logs contain the actions of the Moodle users which role can be of a
professors or a students. The data contains features for Students but no
features for professors. Table~\ref{tab:students_features}, contains all the
features of students.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Feature                   & Data Type  \\ \hline
        \textit{StudentsId}       & Number     \\
        \textit{StudentsNumber}   & Number     \\
        \textit{MoodleUsername}   & String     \\
        \textit{Name}             & String     \\
        \textit{EnrollementCount} & Number     \\
        \textit{StudentType}      & Enumerated \\
    \end{tabular}

    \caption
        [Student features]
        {Student features.}

    \label{tab:students_features}
\end{table}

From table~\ref{tab:students_features} it is observable that a student may be
identified by it's internal Moodle Student ID, or by it's student number as
given by SIIUE. Some students in this data repository do not have a
\textit{StudentsNumber} value, but those students also don't have any activity
in the logs.

The Students Number has a leading character indicating the cycle of that
student. For example, a Master's Degree student has a number of the form
\textit{m12345}, starting with an \textit{m}, and a PhD student has a number
\textit{p12345} starting with a \textit{p}. In the data, the values for
\textit{StudentsNumber} are typed as number and only contain the number part.
But the field \textit{MoodleUsername} has the complete student's number.

The values for the \textit{StudentType} field are:

\begin{itemize}
    \item \textit{Normal}
    \item \textit{Interno}
\end{itemize}

Professors have only the features listed in table~\ref{tab:professor_features}.
The data contains a field called \textit{Role} which states if that Moodle user
is a professor or a student. Supposedly, if the user is a professor the role
has a value of 5. If it is a student, it has a value of 3. However, we see
entries for professors with a role value of 3 and vice versa. It is unknown if
this role actually state the if a user is a professor or student or something
else.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Feature                 & Data Type \\ \hline
        \textit{MoodleUsername} & String    \\
        \textit{Name}           & String    \\
    \end{tabular}

    \caption
        [Professor features]
        {Professor features.}

    \label{tab:professor_features}
\end{table}

Table~\ref{tab:moodle_users_stats} shows some stats regarding Moodle users.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Number of Students   & 145 \\ \hline
        Number of Professors & 87  \\ \hline
        Total                & 232 \\ \hline
    \end{tabular}

    \caption
        [Professor features]
        {Professor features.}

    \label{tab:moodle_users_stats}
\end{table}

\subsection{Results}

The data repository contains the results of students in the courses where they
participated. A listing of grades students had in particular courses is
available. Each object of that dataset, referred to as a Result, has the
features listing in table~\ref{tab:results_fetures}. The values for the
enumerated types are in table~\ref{tab:results_fetures_enum}.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Feature                  & Data Type  \\ \hline
        \textit{CourseCodeSiiue} & String     \\
        \textit{StudentsNumber}  & Number     \\
        \textit{Grade}           & Number     \\
        \textit{Results}         & Enumerated \\
        \textit{FinalResult}     & Enumerated \\
    \end{tabular}

    \caption
        [Results features]
        {Results features.}

    \label{tab:results_fetures}
\end{table}

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l |}
        \hline
        Results & \{ \textit{Anulado},
                     \textit{Aprovado},
                     \textit{Desistiu},
                     \textit{Faltou},
                     \textit{Reprovado}
                  \} \\ \hline
        FinalResult & \{ N, S, n/a \}  \\ \hline
    \end{tabular}

    \caption
        [Values for enumerated types on results]
        {Values for enumerated types on results.}

    \label{tab:courses_enum}
\end{table}

A student has a grade if he was approved (Results field has a value of
\textit{Aprovado}). The value for \textit{FinalResult} is \textit{S} if and
only if the value of \textit{Results} is \textit{Aprovado}. Else, it will be
\textit{N}.

The fields \textit{Results} and \textit{FinalResult} are always empty together
and in that cases there is never any grade.

Table~\ref{tab:results_stats} shows some statistics on the results data.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Number of results           & 716 \\ \hline
        Number of non-empty results & 57  \\ \hline
        Number of approved students & 334 \\ \hline
    \end{tabular}

    \caption
        [Statistics on the Results]
        {Statistics on the Results.}

    \label{tab:results_stats}
\end{table}

\subsection{Moodle Logs}

Moodle logs contain the actual activities of Moodle users. Each log entry
states a user who performed a~\gls{crud} action in a course page. The actions
are aggregated per week. Table~\ref{tab:moodle_logs_features} shows the
features of the Moodle logs, and table~\ref{tab:moodle_logs_enum} shows the
values of the enumerated fields. There are a total of 25207 entries in the
Logs.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Feature                            & Data Type  \\ \hline
        \textit{CourseCodeMoodle}          & String     \\
        \textit{Role}                      & Enumerated \\
        \textit{MoodleUsername}            & String     \\
        \textit{StudentsId}                & Number     \\
        \textit{CRUD}                      & Enumerated \\
        \textit{Week}                      & Number     \\
        \textit{NumberOfActivitiesPerWeek} & Number     \\
        \textit{StudentsNumber}            & Number     \\
    \end{tabular}

    \caption
        [Moode Logs features]
        {Moode Logs features.}

    \label{tab:moodle_logs_features}
\end{table}

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l |}
        \hline
        Role & \{ 3, 5 \} \\ \hline
        CRUD & \{ C, R, U, D \}  \\ \hline
    \end{tabular}

    \caption
        [Values for enumerated types on Moodle Logs]
        {Values for enumerated types on Moodle Logs.}

    \label{tab:moodle_logs_enum}
\end{table}

% =============================================================================
\section{Preprocessed Datasets}

From the original data we make an initial pre processing in order to have a
consistent repository of datasets ready for further analysis. The pre-processed
datasets have consistent names and fields.

% CoursesGeneral.csv
% CoursesEntriesOverview.csv
% MoodleUsers.csv
% Students.csv
% CoursesStudents.csv
% Results.csv
% MoodleLogs.csv
% CoursesNoMoodle.csv

### 2017-02-10 exps.tex
Adds data and experiments parts

\chapter{Experiments}

\section{EXP\_001 - Plotting Activity of Students over the Semester}

\subsection{Objective}

To look for patterns in the usage of Moodle we look into plotting the activity
of students over the courses over a semester. There are many ways of doing so.

One idea is to plot the average number of activities of students on one course.
Over a semester, several students access a specific course where they read and
write forum messages, deliver assignments, etc. For the various courses, we
plot the average of the total number of activities, and the average number of
each individual~\gls{crud} activities (\textbf{C}reate, \textbf{R}ead, etc).

\subsection{Procedure}

\subsection{Conclusions}

Some graphs show distinct peaks. What are they fore? Is it for a new
assignment, or just before an evaluation?

Some courses have a greater number of peaks them others. Do these courses have
more students, more assignments, better grades?

### 2017-02-11 papers.tex
Reorganizes background section

@@
-\chapter{Papers}
+\section{Initial Papers in Detail}
+
+This section contain long comments about the first papers which were explored
+for this project.
 
 % =============================================================================
-\section{\texttt{IND\_001} - Data Mining Algorithms to Classify Students}
+\subsection{\texttt{IND\_001} - Data Mining Algorithms to Classify Students}
 
@@
 
-\subsection{Comment}
+\subsubsection{Comment}
 
@@
 
-\subsection{Relevant Points}
+\subsubsection{Relevant Points}
 
@@
 
-\subsection{Work From This Paper}
+\subsubsection{Work From This Paper}
 
@@
 % =============================================================================
-\section{\texttt{IND\_002} - Discovering Decision Knowledge From Web Log
+\subsection{\texttt{IND\_002} - Discovering Decision Knowledge From Web Log
 Portfolio For Managing Classroom Processes By Applying Decision Tree And Data
@@
 % =============================================================================
-\section{\texttt{IND\_003} - Predicting Performance From Test Scores Using
+\subsection{\texttt{IND\_003} - Predicting Performance From Test Scores Using
 Backpropagation and Counterpropagation}
@@
 % =============================================================================
-\section{\texttt{IND\_004} - Web Usage Mining for Predicting Final Marks of
+\subsection{\texttt{IND\_004} - Web Usage Mining for Predicting Final Marks of
 Students That Use Moodle Courses}
@@
 
-\subsection{Reading Notes}
+\subsubsection{Reading Notes}
 
@@
 % =============================================================================
-\section{\texttt{IND\_005} - Using Educational Data Mining Methods to Study the
+\subsection{\texttt{IND\_005} - Using Educational Data Mining Methods to Study the
 Impact of Virtual Classroom in E-Learning}
@@
 % =============================================================================
-\section{\texttt{IND\_006} - Predicting students' marks from Moodle logs using
+\subsection{\texttt{IND\_006} - Predicting students' marks from Moodle logs using
 neural network models}
@@
 % =============================================================================
-\section{\texttt{IND\_007} - Educational data mining: A survey from 1995 to
+\subsection{\texttt{IND\_007} - Educational data mining: A survey from 1995 to
 2005}
@@
 % =============================================================================
-\section{\texttt{IND\_008} - Educational Data Mining: A Review of the State of
+\subsection{\texttt{IND\_008} - Educational Data Mining: A Review of the State of
 the Art}

### 2017-02-11 data.tex
Adds new paper
Finished some points in the data section

@@
 like Python with Pandas, they were exported to a CSV format. Some
-transformations may be applied to these files in other to clean useless points
-or make the data easier to process.
+transformations may be applied to these files in other to clean useless data
+points or make the data easier to process.
 
 To begin, each sheet is analysed individually. Conclusions on each list are
-registered in the fist section of this chapter.
+registered in the fist section of this chapter. After analysing the original
+data, the second section of the chapter asserts which entities and objects are
+present in this data. Finally, the third section lists the datasets which
+represent those entities and objects.
 
@@
 
-There is a lot of information in this sheet.
+There is a lot of information in this sheet. Basically, this sheet contains a
+big list. That list has a course, a student, and the student's results on that
+course. The results contain the grade, if available, if the student was
+approved or not, and if not, why? Did he failed the evaluations or simply
+missed them?
 
-\TODO{Explore}
+Some statistics are calculated in a table to the right of this sheet. Those
+calculated values are useful, but should be recalculated elsewhere in a later
+time if necessary.
 
 \subsection{Sheet 5 - \texttt{Profile}}
+
+\TODO{Some information is missing for this sheet}
 
@@
 
-In order for a student to get approved, some creteria must be followed.
+In order for a student to get approved, some criteria must be followed. That
+criteria is described in the sheet itself.
 
 \subsection{Sheet 6 - \texttt{ResultadosUCMoodle}}
+
+\TODO{Some information is missing for this sheet}
 
@@
 
-    \label{tab:courses_enum}
+    \label{tab:results_fetures_enum}
 \end{table}

### 2017-02-11 exps.tex
Adds exp 001

@@
-\chapter{Experiments}
+\chapter{Experiments and Statistics}
 
+% =============================================================================
 \section{EXP\_001 - Plotting Activity of Students over the Semester}
 
-\subsection{Objective}
+To look for patterns in the usage of Moodle we look into plotting the activity
+of students over the courses over a semester. The six courses with the greatest
+number of activities are chosen and for each of them two plots are made.
 
-To look for patterns in the usage of Moodle we look into plotting the activity
-of students over the courses over a semester. There are many ways of doing so.
+One plot displays the overall number of~\gls{crud} activities over the semester
+and the number of \textbf{C}reate activities, \textbf{R}ead activities, etc.
+Figure~\ref{fig:exp_001_all} show these plots for two courses. As it can be
+seen, the number of Read activities is far greater then the others, which is to
+be expected since to do any other activity, at least a Read activity has to
+take place. For example, a student needs to open the Moodle page, and therefore
+do a Read, to deliver a project, and therefore do a Create.
 
-One idea is to plot the average number of activities of students on one course.
-Over a semester, several students access a specific course where they read and
-write forum messages, deliver assignments, etc. For the various courses, we
-plot the average of the total number of activities, and the average number of
-each individual~\gls{crud} activities (\textbf{C}reate, \textbf{R}ead, etc).
+It is hypothesis that the number is also greater because students will usually
+just read new messages in the Moodle, or consult new materials, without doing
+any other activity.
 
-\subsection{Procedure}
+The second plot made to each course is similar, but only the Create, Update,
+and Delete activities are plotted. Figure~\ref{fig:exp_001_cud} show two
+examples these plots. Notice that the number of creates is not always greater
+then the number of updates, and vice-versa. But the number of deletes is always
+below creates and updates. This concludes that the delete activity is simply
+not that common.
 
-\subsection{Conclusions}
+\begin{figure}[h!]
+    \centering
 
-Some graphs show distinct peaks. What are they fore? Is it for a new
-assignment, or just before an evaluation?
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{imgs/fig_0_CTE-UMCED-2_all}
+        \label{subfig:exp_001_0_all}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{imgs/fig_3_INF7194_2_all}
+        \label{subfig:exp_001_3_all}
+    \end{subfigure}
 
-Some courses have a greater number of peaks them others. Do these courses have
-more students, more assignments, better grades?
+    \caption
+        [Exp 001 plots for all activities]
+        {Exp 001 plots for all activities.}
+
+    \label{fig:exp_001_all}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{imgs/fig_0_CTE-UMCED-2_cud}
+        \label{subfig:exp_001_0_cud}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{imgs/fig_3_INF7194_2_cud}
+        \label{subfig:exp_001_3_cud}
+    \end{subfigure}
+
+    \caption
+        [Exp 001 plots for C, U, and D activities]
+        {Exp 001 plots for C, U, and D activities.}
+
+    \label{fig:exp_001_cud}
+\end{figure}

### 2017-02-11 ResearchNotes/exps.tex
Moved experiments to this repository

--- 2017-02-11 exps.tex
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_0_CTE-UMCED-2_all}
+        \includegraphics[width=\linewidth]{../Exps/exp_001_results/fig_0_CTE-UMCED-2_all}
         \label{subfig:exp_001_0_all}
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_3_INF7194_2_all}
+        \includegraphics[width=\linewidth]{../Exps/exp_001_results/fig_3_INF7194_2_all}
         \label{subfig:exp_001_3_all}
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_0_CTE-UMCED-2_cud}
+        \includegraphics[width=\linewidth]{../Exps/exp_001_results/fig_0_CTE-UMCED-2_cud}
         \label{subfig:exp_001_0_cud}
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_3_INF7194_2_cud}
+        \includegraphics[width=\linewidth]{../Exps/exp_001_results/fig_3_INF7194_2_cud}
         \label{subfig:exp_001_3_cud}

### 2017-02-11 ResearchNotes/papers.tex
Moved experiments to this repository

--- 2017-02-11 papers.tex
@@
 % =============================================================================
-\subsection{\texttt{IND\_008} - Educational Data Mining: A Review of the State of
-the Art}
+\subsection{\texttt{IND\_008} - Educational Data Mining: A Review of the State
+of the Art}
 
@@
 extremely relevant.
+
+% =============================================================================
+\subsection{\texttt{IND\_009} - Evolutionary algorithms for subgroup discovery
+applied to e-learning data}
+
+Paper is~\cite{ind_009}. Applying evolutionary algorithms to find association
+rules.
+
+Association rules are rules that link events together following the form, ``If
+event $ A $ happened, the event $ B $ is likely to occur''. These are fuzzy
+rules, because an event $ A $ happening doesn't necessarily imply that event $
+B $ will always happen, rather, the rules give a probabilistic notion that
+should be understood has, if $ A $ happened, then $ B $ will probably also
+happen.
+
+The paper talks about subgroups in the context of association rules and genetic
+algorithms to learn rules from data. The classic association rules learning
+algorithms, which are Apriori-SD and CN2-SD, are mentioned.
+
+Like other papers from these authors, the usual Moodle data is shown and
+several algorithms are applied over it. The results for those algorithms is
+presented and discussed. The results seemed to be quite good.

### 2017-02-11 Exps/data.py

@@ -1,0 +1,19 @@
+import os
+
+
+import pandas
+
+
+data_root = "../../Data/Processed"
+
+CoursesEntriesOverview = os.path.join(data_root, "CoursesEntriesOverview.csv")
+CoursesGeneral         = os.path.join(data_root, "CoursesGeneral.csv")
+CoursesNoMoodle        = os.path.join(data_root, "CoursesNoMoodle.csv")
+CoursesStudents        = os.path.join(data_root, "CoursesStudents.csv")
+MoodleLogs             = os.path.join(data_root, "MoodleLogs.csv")
+MoodleUsers            = os.path.join(data_root, "MoodleUsers.csv")
+Results                = os.path.join(data_root, "Results.csv")
+Students               = os.path.join(data_root, "Students.csv")
+
+def load(sheet):
+    return pandas.read_csv(sheet)

### 2017-02-11 Exps/exp_001.py

@@ -1,0 +1,130 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Plot activities of students over the semester.
+#
+# This experiment aims at making different plots in order to visualize the #
+# activity of students over the semester. He pick the six courses with the most
+# activities registered in the logs. We then plot those activities over the
+# whole semester and see the relation between the overall number of activities,
+# activities that are only Create activities, only Update, etc.
+#
+# It is noticeable that Read activities are disproportionally greater than the
+# other activities, so Create, Update, and Delete activities are also plotted
+# in their own graphs
+# =============================================================================
+
+import os
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+import data
+
+
+output_dir = "exp_001_results"
+
+
+def outputFileName(index, courses, plot_type):
+    file_name = "fig_%d_%s_%s" % (index, courses, plot_type)
+    return os.path.join(output_dir, file_name)
+
+
+def takeGroup(df):
+    return df.groupby(["CourseCodeMoodle", "Week"]).agg("sum")
+
+
+def takeCRUD(df, crud):
+    return df.loc[lambda df: df["CRUD"] == crud]
+
+
+def plotCourse(index, course, groups):
+    # Get this course's groups
+    group_all = groups["all"].loc[course]
+    group_C   = groups["C"].loc[course]
+    group_R   = groups["R"].loc[course]
+    group_U   = groups["U"].loc[course]
+    group_D   = groups["D"].loc[course]
+
+    # Get weeks
+    weeks_all = np.array(group_all.index.get_level_values("Week").tolist())
+    weeks_C = np.array(group_C.index.get_level_values("Week").tolist())
+    weeks_R = np.array(group_R.index.get_level_values("Week").tolist())
+    weeks_U = np.array(group_U.index.get_level_values("Week").tolist())
+    weeks_D = np.array(group_D.index.get_level_values("Week").tolist())
+
+    # Get number of activities
+    act_all = np.array(group_all.values.tolist()).flatten()
+    act_C = np.array(group_C.values.tolist()).flatten()
+    act_R = np.array(group_R.values.tolist()).flatten()
+    act_U = np.array(group_U.values.tolist()).flatten()
[…]

### 2017-02-11 Exps/exp_002.py

@@ -1,0 +1,91 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Mean of activities per week. It falls. Write this conclusion
+
+"""
+The mean of the activities per week drops. But the number of courses which end
+drops with it.
+
+Is the number of activities is consistent along the semester? Or does it drop?
+
+Make further experiments.
+"""
+
+# =============================================================================
+
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+
+import data
+
+
+def numberOfActs(alunos_moodle_data):
+    # Get maximum of weeks
+    max_weeks = alunos_moodle_data["Week"].max()
+
+    # Select only useful columns for this experience
+    orig_cols = alunos_moodle_data[[
+        "CourseCodeMoodle",
+        "Week",
+        "NumberOfActivitiesPerWeek"
+    ]]
+
+    # Group by course code and week number, having the total number of
+    # activities for that week and for that course.
+    group = orig_cols.groupby(["CourseCodeMoodle", "Week"], as_index=False).agg("sum")
+    group_sum = group.groupby(by="Week").agg("sum")
+    sum_values = group_sum["NumberOfActivitiesPerWeek"].values.tolist()
+
+    pyplot.plot(range(len(sum_values)), np.array(sum_values))
+    pyplot.show()
+
+
+def weeks(alunos_moodle_data):
+    # Get maximum of weeks
+    max_weeks = alunos_moodle_data["Week"].max()
+
+    # Select only useful columns for this experience
+    orig_cols = alunos_moodle_data[[
+        "CourseCodeMoodle",
+        "Week"
+    ]]
+
+    # Number of courses with activities until this point
+    group = orig_cols.groupby(["CourseCodeMoodle"], as_index=False).agg("max")
+
+    counts = group.groupby(["Week"]).count()
+
+    max_week_per_course = []
+    for i in range(max_weeks+1):
+        if i in counts.index:
+            max_week_per_course.append(counts.loc[i].values[0])
+        else:
+            max_week_per_course.append(0)
+
+    # Reverse acc
+    rev_acc = []
+    acc = 0
+    for i in max_week_per_course[::-1]:
+        acc += i
+        rev_acc.insert(0, acc)
+
+    fig, p1 = pyplot.subplots()
[…]

### 2017-02-12 ResearchNotes/exps.tex
Adds statistics

--- 2017-02-11 exps.tex
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_0_CTE-UMCED-2_all}
+        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_0_CTE-UMCED-2_all}
         \label{subfig:exp_001_0_all}
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_3_INF7194_2_all}
+        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_3_INF7194_2_all}
         \label{subfig:exp_001_3_all}
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_0_CTE-UMCED-2_cud}
+        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_0_CTE-UMCED-2_cud}
         \label{subfig:exp_001_0_cud}
@@
         \centering
-        \includegraphics[width=\linewidth]{imgs/fig_3_INF7194_2_cud}
+        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_3_INF7194_2_cud}
         \label{subfig:exp_001_3_cud}

### 2017-02-12 src/exp_001.py

@@ -21,6 +21,7 @@ import pandas
 from matplotlib import pyplot
 
 import data
+import util
 
 
 output_dir = "exp_001_results"
@@ -120,10 +121,7 @@ if __name__ == "__main__":
     }
 
     # Create the directory for the results if it doesn't exist
-    try:
-        os.mkdir(output_dir)
-    except FileExistsError:
-        pass
+    util.makeDir(output_dir)
 
     # Get week number and activities number of a specific course
     for index, course in enumerate(common_courses.index.tolist()):

### 2017-02-12 src/stat_001.py

@@ -1,0 +1,35 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# This experiment plots the number of courses for each possible number of
+# credits in a bars chart.
+# =============================================================================
+
+import os
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+import data
+import util
+
+
+output_dir = "stat_001_results"
+
+
+if __name__ == "__main__":
+    df = data.load(data.CoursesGeneral)
+
+    groups = df.groupby("Credits").agg("count")
+
+    print(groups["CourseName"])
+
+    courses = groups.index.tolist()
+    counts = groups["CourseName"].values.tolist()
+
+    util.makeDir(output_dir)
+
+    pyplot.bar(courses, counts)
+    pyplot.title("Number of Courses Grouped by Credits")
+    pyplot.savefig(os.path.join(output_dir, "counts"))

### 2017-02-12 src/util.py

@@ -1,0 +1,7 @@
+import os
+
+def makeDir(dir_name):
+    try:
+        os.mkdir(dir_name)
+    except FileExistsError:
+        pass

### 2017-02-12 ResearchNotes/stat_exps.tex
Adds text on statistics

--- 2017-02-12 ResearchNotes/exps.tex
@@
-\chapter{Experiments and Statistics}
+\chapter{Statistics and Experiments}
 
 % =============================================================================
-\section{EXP\_001 - Plotting Activity of Students over the Semester}
+\section{Statistics}
+
+\subsection{\texttt{STAT\_001} - Group Courses by Credits}
+
+This experiment plots the number of courses for each possible number of credits
+in a bars chart. The results of this was listed in
+table~\ref{tab:courses_credits}.
+
+\subsection{\texttt{STAT\_002} - List Domain of Courses Fields}
+
+Gets domain of the fields of courses. Relevant results are listed in some
+tables in section~\ref{sec:data_courses}.
+
+% =============================================================================
+\section{Experiments}
+
+\subsection{\texttt{EXP\_001} - Plotting Activity of Students over the
+Semester}
 

### 2017-02-12 src/stat_002.py

@@ -1,0 +1,35 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Get the domain for a few fields of the courses
+# =============================================================================
+
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+
+import data
+
+def uniq(cg, field):
+    print("%s:" % (field))
+
+    for i in cg[field].unique():
+        print("\t%s" % (i))
+
+    print("")
+
+
+if __name__ == "__main__":
+    # Load data
+    cg = data.load(data.CoursesGeneral)
+
+    uniq(cg, "Department")
+    uniq(cg, "Cycle")
+    uniq(cg, "Regime")
+    uniq(cg, "Credits")
+    uniq(cg, "Degree")
+    uniq(cg, "Semester")
+    uniq(cg, "Season")
+    uniq(cg, "Type")

### 2017-02-12 ResearchNotes/stat_exps.tex
Adds stat 003

@@
 tables in section~\ref{sec:data_courses}.
+
+\subsection{\texttt{STAT\_003} - Number of Courses per Degree}
+
+A list of courses per degree is computed. This gives an idea of how many
+courses there are in a given degree. A list of courses per degree and semester
+is also given to give an idea of how the courses are distributed across the two
+semesters.
+
+Tables~\ref{tab:stat_003_res_1} and~\ref{tab:stat_003_res_2} show these results.
+
+It should be noticeable that some degrees only have courses in one semester.
+The names of the degrees are not available, only their codes. But it is
+noticeable that the degree with the greater number of courses is
+\texttt{B\_M\_REST\_EINF\_E}, which appears to be an Informatics Master's
+Degree.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{l r}
+        Degree                             & CourseCodeMoodle \\ \hline
+        \texttt{B\_M\_EHEA (478)}          & 7                \\
+        \texttt{B\_M\_REST\_EINF\_E (578)} & 28               \\
+        \texttt{B\_M\_REST\_TET (457)}     & 7                \\
+        \texttt{B\_PD\_M\_E (580)}         & 8                \\
+        \texttt{CCDND\_CPM (344)}          & 1                \\
+        \texttt{FC\_CVP\_FDOC (513)}       & 2                \\
+        \texttt{FC\_CVP\_UAM (514)}        & 3                \\
+        \texttt{FC\_EDM (564)}             & 1                \\
+        \texttt{PG\_B\_AE (438)}           & 10               \\
+        \texttt{PG\_B\_EGNEG (554)}        & 10               \\
+    \end{tabular}
+
+    \caption
+        [Number of Courses per Degree]
+        {Number of Courses per Degree.}
+
+    \label{tab:stat_003_res_1}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{l l r}
+        Degree                             & Semester       & CourseCodeMoodle \\ \hline
+        \texttt{B\_M\_EHEA (478)}          & \textit{Par}   & 3                \\
+                                           & \textit{Ímpar} & 4                \\
+        \texttt{B\_M\_REST\_EINF\_E (578)} & \textit{Par}   & 14               \\
+                                           & \textit{Ímpar} & 14               \\
+        \texttt{B\_M\_REST\_TET (457)}     & \textit{Par}   & 2                \\
+                                           & \textit{Ímpar} & 5                \\
+        \texttt{B\_PD\_M\_E (580)}         & \textit{Par}   & 3                \\
+                                           & \textit{Ímpar} & 5                \\
+        \texttt{PG\_B\_EGNEG (554)}        & \textit{Par}   & 4                \\
+                                           & \textit{Ímpar} & 5                \\
+        \texttt{PG\_B\_AE (438)}           & \textit{Par}   & 5                \\
+                                           & \textit{Ímpar} & 5                \\
+        \texttt{FC\_EDM (564)}             & \textit{Par}   & 1                \\
+        \texttt{CCDND\_CPM (344)}          & \textit{Ímpar} & 1                \\
+        \texttt{FC\_CVP\_FDOC (513)}       & \textit{Ímpar} & 2                \\
+        \texttt{FC\_CVP\_UAM (514)}        & \textit{Ímpar} & 3                \\
+    \end{tabular}
+
+    \caption
+        [Number of Courses per Degree and Semester]
+        {Number of Courses per Degree and Semester.}
+
+    \label{tab:stat_003_res_2}
+\end{table}
 

### 2017-02-12 src/stat_003.py

@@ -1,0 +1,31 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Gets the number of courses per degree and the number of courses per degree
+# and semester
+# =============================================================================
+
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+
+import data
+
+
+if __name__ == "__main__":
+    # Load data
+    df = data.load(data.CoursesGeneral)
+
+    # Get by degree
+    by_deg_df = df[["Degree", "CourseCodeMoodle"]]
+    by_deg = by_deg_df.groupby("Degree").agg("count")
+
+    # Get by degree and semester
+    by_deg_sem_df = df[["Degree", "CourseCodeMoodle", "Semester"]]
+    by_deg_sem = by_deg_sem_df.groupby(["Degree", "Semester"]).agg("count")
+
+    # Print
+    print(by_deg)
+    print(by_deg_sem)

### 2017-02-12 ResearchNotes/stat_exps.tex
Renamed exp 001 to stat 004, since it is a stat

@@
 
-% =============================================================================
-\section{Experiments}
-
-\subsection{\texttt{EXP\_001} - Plotting Activity of Students over the
+\subsection{\texttt{STAT\_004} - Plotting Activity of Students over the
 Semester}
@@
 and the number of \textbf{C}reate activities, \textbf{R}ead activities, etc.
-Figure~\ref{fig:exp_001_all} show these plots for two courses. As it can be
+Figure~\ref{fig:stat_004_all} show these plots for two courses. As it can be
 seen, the number of Read activities is far greater then the others, which is to
@@
 The second plot made to each course is similar, but only the Create, Update,
-and Delete activities are plotted. Figure~\ref{fig:exp_001_cud} show two
+and Delete activities are plotted. Figure~\ref{fig:stat_004_cud} show two
 examples these plots. Notice that the number of creates is not always greater
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_0_CTE-UMCED-2_all}
-        \label{subfig:exp_001_0_all}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_0_CTE-UMCED-2_all}
+        \label{subfig:stat_004_0_all}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_3_INF7194_2_all}
-        \label{subfig:exp_001_3_all}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_3_INF7194_2_all}
+        \label{subfig:stat_004_3_all}
     \end{subfigure}
@@
     \caption
-        [Exp 001 plots for all activities]
-        {Exp 001 plots for all activities.}
+        [Stat 004 plots for all activities]
+        {Stat 004 plots for all activities.}
 
-    \label{fig:exp_001_all}
+    \label{fig:stat_004_all}
 \end{figure}
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_0_CTE-UMCED-2_cud}
-        \label{subfig:exp_001_0_cud}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_0_CTE-UMCED-2_cud}
+        \label{subfig:stat_004_0_cud}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/exp_001_results/fig_3_INF7194_2_cud}
-        \label{subfig:exp_001_3_cud}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_3_INF7194_2_cud}
+        \label{subfig:stat_004_3_cud}
     \end{subfigure}
@@
     \caption
-        [Exp 001 plots for C, U, and D activities]
-        {Exp 001 plots for C, U, and D activities.}
+        [Stat 004 plots for C, U, and D activities]
+        {Stat 004 plots for C, U, and D activities.}
 
-    \label{fig:exp_001_cud}
+    \label{fig:stat_004_cud}
 \end{figure}
+
+% =============================================================================
+\section{Experiments}

### 2017-02-12 src/stat_004.py

@@ -3,7 +3,7 @@
 # =============================================================================
 # Plot activities of students over the semester.
 #
-# This experiment aims at making different plots in order to visualize the #
+# This statistic aims at making different plots in order to visualize the
 # activity of students over the semester. He pick the six courses with the most
 # activities registered in the logs. We then plot those activities over the
 # whole semester and see the relation between the overall number of activities,
@@ -24,7 +24,7 @@ import data
 import util
 
 
-output_dir = "exp_001_results"
+output_dir = "stat_004_results"
 
 
 def outputFileName(index, courses, plot_type):
@@ -97,7 +97,7 @@ if __name__ == "__main__":
     # Load data
     df = data.load(data.MoodleLogs)
 
-    # Select only useful columns for this experience
+    # Select only useful columns for this statistic
     df_redux = df[[
         "CourseCodeMoodle",
         "CRUD",

### 2017-02-12 src/stat_001.py
Adds code for stats 5 and 6

@@ -15,8 +15,7 @@ import data
 import util
 
 
-output_dir = "stat_001_results"
-
+output_dir = util.outputDirName(1)
 
 if __name__ == "__main__":
     df = data.load(data.CoursesGeneral)

### 2017-02-12 src/stat_004.py

@@ -24,7 +24,7 @@ import data
 import util
 
 
-output_dir = "stat_004_results"
+output_dir = util.outputDirName(4)
 
 
 def outputFileName(index, courses, plot_type):

### 2017-02-12 src/stat_005.py

@@ -1,0 +1,55 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Plot total number of activities per week and mean number of activities per
+# course per week. Both plots are in the same graph.
+# =============================================================================
+
+
+import os
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+
+import data
+import util
+
+
+output_dir = util.outputDirName(5)
+
+
+if __name__ == "__main__":
+    # Load data
+    df = data.load(data.MoodleLogs)
+
+    # Get maximum of weeks
+    max_weeks = df["Week"].max()
+
+    # Select only useful columns for this experience
+    orig_cols = df[[
+        "CourseCodeMoodle",
+        "Week",
+        "NumberOfActivitiesPerWeek"
+    ]]
+
+    # Group by course code and week number, having the total number of
+    # activities for that week and for that course.
+    group = orig_cols.groupby(["CourseCodeMoodle", "Week"], as_index=False).agg("sum")
+    group_sum = group.groupby(by="Week").agg("sum")
+    group_mean = group.groupby(by="Week").agg("mean")
+
+    sum_values = np.array(group_sum["NumberOfActivitiesPerWeek"].values.tolist())
+    mean_values = np.array(group_mean["NumberOfActivitiesPerWeek"].values.tolist())
+
+    fig, p1 = pyplot.subplots()
+    p1.plot(range(max_weeks+1), sum_values)
+    p2 = p1.twinx()
+    p2.plot(range(max_weeks+1), mean_values, color="red")
+
+    util.makeDir(output_dir)
+
+    fig.tight_layout()
+    pyplot.title("Active Courses vs Number of Activities")
+    pyplot.savefig(os.path.join(output_dir, "plot"))

### 2017-02-12 src/stat_006.py

@@ -1,19 +1,10 @@
 #!/usr/bin/env python3
 
 # =============================================================================
-# Mean of activities per week. It falls. Write this conclusion
-
-"""
-The mean of the activities per week drops. But the number of courses which end
-drops with it.
-
-Is the number of activities is consistent along the semester? Or does it drop?
-
-Make further experiments.
-"""
-
+# Number of active courses vs total number of activities
 # =============================================================================
 
+import os
 
 import numpy as np
 import pandas
@@ -21,14 +12,15 @@ from matplotlib import pyplot
 
 
 import data
+import util
 
 
-def numberOfActs(alunos_moodle_data):
-    # Get maximum of weeks
-    max_weeks = alunos_moodle_data["Week"].max()
+output_dir = util.outputDirName(6)
 
+
+def numberOfActs(df):
     # Select only useful columns for this experience
-    orig_cols = alunos_moodle_data[[
+    orig_cols = df[[
         "CourseCodeMoodle",
         "Week",
         "NumberOfActivitiesPerWeek"
@@ -40,52 +32,58 @@ def numberOfActs(alunos_moodle_data):
     group_sum = group.groupby(by="Week").agg("sum")
     sum_values = group_sum["NumberOfActivitiesPerWeek"].values.tolist()
 
-    pyplot.plot(range(len(sum_values)), np.array(sum_values))
-    pyplot.show()
+    return np.array(sum_values)
 
 
-def weeks(alunos_moodle_data):
-    # Get maximum of weeks
-    max_weeks = alunos_moodle_data["Week"].max()
-
+def weeks(max_weeks, df):
     # Select only useful columns for this experience
-    orig_cols = alunos_moodle_data[[
+    orig_cols = df[[
         "CourseCodeMoodle",
         "Week"
     ]]
+
+    # Get maximum of weeks
+    max_weeks = df["Week"].max()
 
     # Number of courses with activities until this point
     group = orig_cols.groupby(["CourseCodeMoodle"], as_index=False).agg("max")
 
     counts = group.groupby(["Week"]).count()
 
-    max_week_per_course = []
-    for i in range(max_weeks+1):
-        if i in counts.index:
-            max_week_per_course.append(counts.loc[i].values[0])
-        else:
-            max_week_per_course.append(0)
[…]

### 2017-02-12 src/util.py
New functions in util

@@ -5,3 +5,7 @@ def makeDir(dir_name):
         os.mkdir(dir_name)
     except FileExistsError:
         pass
+
+
+def outputDirName(num):
+    return "stat_%03d_results" % (num)

### 2017-02-12 ResearchNotes/stat_exps.tex
Coomented stats 5 and 6

@@
 
+\subsection{\texttt{STAT\_005} - Total Number of Activities vs Mean of
+Activities per Week}
+
+This analysis attempts to understand if the activities of courses remain the
+same as the semester goes on. So the mean of the activities per course and week
+is plotted, as it is displayed in figure~\ref{fig:stat_005}. The total number
+of activities is also plotted in figure~\ref{fig:stat_005}.
+
+It is visible that, although the number of activities drops as the semester
+goes on, the mean number of activities stayed relatively the same despite
+varying a lot. This is due to the fact that some courses stop being active as
+the semester goes on, as it will be shown in section~\ref{sec:stat_006}, but
+despite that, the courses that do remain active don't register a drop of
+activity as time goes on.
+
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../src/stat_005_results/plot}
+
+    \caption
+        [Total Number of Activities vs Mean of Activities per Week]
+        {Total Number of Activities vs Mean of Activities per Week.}
+
+    \label{fig:stat_005}
+\end{figure}
+
+\subsection{\texttt{STAT\_006} - Number of Active Courses vs Number of
+Activities}
+\label{sec:stat_006}
+
+A correlation is found between the total number of activities and the number of
+active courses per week. The total number of active courses in a week is
+plotted as a bars chart in figure~\ref{fig:stat_006}, along with the total
+number of activities.
+
+An course ceases to be active on a given week where it no longer has activities
+after that week. For example, suppose a course has activities on weeks, 0, 1,
+4, and 8. Then the course is only active until week 8.
+
+The number of activities goes down with the number of active courses, as it
+would be expected. But around weeks 30 to 50, it is visible that the number of
+activities significantly drops from the number of courses, as opposed to the
+first weeks where the numbers are much more similar. This is +probably due to
+students incising lack of interest in courses' activities as the semester goes
+on, which is normal behaviour.
+
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../src/stat_006_results/plot}
+
+    \caption
+        [Active Courses vs Number of Activities]
+        {Active Courses vs Number of Activities.}
+
+    \label{fig:stat_006}
+\end{figure}
+
 % =============================================================================

### 2017-02-12 src/stat_005.py

@@ -44,12 +44,16 @@ if __name__ == "__main__":
     mean_values = np.array(group_mean["NumberOfActivitiesPerWeek"].values.tolist())
 
     fig, p1 = pyplot.subplots()
-    p1.plot(range(max_weeks+1), sum_values)
+    sum_plot = p1.plot(range(max_weeks+1), sum_values)
     p2 = p1.twinx()
-    p2.plot(range(max_weeks+1), mean_values, color="red")
+    mean_plot = p2.plot(range(max_weeks+1), mean_values, color="red")
 
     util.makeDir(output_dir)
 
     fig.tight_layout()
-    pyplot.title("Active Courses vs Number of Activities")
+    pyplot.title("Total Number of Activities vs Mean of Activities per Week")
+    pyplot.figlegend(
+        (sum_plot[0], mean_plot[0]),
+        ("Total", "Mean"),
+        "upper right")
     pyplot.savefig(os.path.join(output_dir, "plot"))

### 2017-02-14 src/exp_001.py
Made dataset creation for exp 1

@@ -1,0 +1,135 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# TOOD
+# =============================================================================
+
+import os
+
+import numpy as np
+import pandas as pd
+from matplotlib import pyplot
+
+import data
+import util
+
+
+def makeDataset(approvedBinary=True):
+    df_stu = data.load(data.Students)
+    df_cge = data.load(data.CoursesGeneral)
+    df_log = data.load(data.MoodleLogs)
+    df_res = data.load(data.Results)
+
+    # Join Students and results to obtain a dataset with an entry for each
+    # student and actual result, set CourseCodeSiiue to be an index along with
+    # StudentsNumber
+    df_stu.set_index("StudentsNumber", inplace=True)
+    df_res.set_index("StudentsNumber", inplace=True)
+    df_stu_res = df_stu.join(df_res)
+
+    df_stu_res.set_index("CourseCodeSiiue", inplace=True, append=True)
+    df_stu_res = df_stu_res[["EnrollementCount", "StudentType", "Grade"]]
+
+    # Get the logs and add CourseCodeSiiue to them. Then take only the useful
+    # columns
+    df_log = df_log.merge(
+            df_cge[["CourseCodeMoodle", "CourseCodeSiiue"]],
+            on = "CourseCodeMoodle",
+            how = "left") \
+        [["StudentsNumber",
+          # "CourseCodeSiiue",
+          "CourseCodeMoodle",
+          "CRUD",
+          "NumberOfActivitiesPerWeek"]]
+
+    # Perform group-bys in order to get count of CRUD activities
+    df_acts_all = df_log \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
+        .agg("sum")
+
+    df_acts_c = df_log \
+        .loc[lambda df: df["CRUD"] == "c"] \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
+        .agg("sum")
+
+    df_acts_r = df_log \
+        .loc[lambda df: df["CRUD"] == "r"] \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
+        .agg("sum")
+
+    df_acts_u = df_log \
+        .loc[lambda df: df["CRUD"] == "u"] \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
[…]

### 2017-02-24 ResearchNotes/data.tex
Moved experiments to this repository
Adds statistics
Adds text on statistics
Adds er model of data to notes
Updated Research Notes

--- 2017-02-11 data.tex
@@
 
-\TODO{Some information is missing for this sheet}
-
 This sheet contains some analysis made to the data. To begin with, there are
@@
     \item \texttt{id\_c} - Id of the course
-    \item \texttt{n\_resrc\_c} - Number of resources in the course
-    \item \texttt{n\_actv\_c} - Number of activities in the course
-    \item \texttt{id\_s} - Id of the student
-    \item \texttt{n\_cons\_s} - \TODO{}
-    \item \texttt{n\_subm\_s} - \TODO{}
+    \item \texttt{n\_resrc\_c} - Number of resources (files, urls, etc)
+    \item \texttt{n\_actv\_c} - Number of activities (forums, quizzes, etc)
+    \item \texttt{id\_s} - Student's id
+    \item \texttt{n\_cons\_s} - Number of views
+    \item \texttt{n\_subm\_s} - Number of submissions
     \item \texttt{result\_s} - Results of that student in that course
@@
 \subsection{Sheet 6 - \texttt{ResultadosUCMoodle}}
-
-\TODO{Some information is missing for this sheet}
 
@@
 
-\TODO{Fields in between}
-
 The sheet finally specifies the number of students who enrolled in a course and
@@
     \item \texttt{int} - Interval between first and last activity of course
-    \item \texttt{actsub}
-    \item \texttt{actcons}
-    \item \texttt{nactsubd}
-    \item \texttt{consal}
-    \item \texttt{consdoc}
-    \item \texttt{subal}
-    \item \texttt{subdoc}
-    \item \texttt{nalu}
-    \item \texttt{ndoc}
+    \item \texttt{actsub} - Number of Submission Activities(Same value ad above)
+    \item \texttt{actcons} - Number of views (Same as resources)
+    \item \texttt{nactsubd} - Number of different activities
+    \item \texttt{consal} - Student views
+    \item \texttt{consdoc} - Professor views
+    \item \texttt{subal} - Student submission
+    \item \texttt{subdoc} - Professor submission
+    \item \texttt{nalu} - Number of students in the course
+    \item \texttt{ndoc} - Number of professors in the course
     \item \texttt{InscSIIUE} - Number of students who enrolled
@@
 \section{Objects, Entities, and Features}
+\label{sec:objs_ents_feat}
 
@@
 \subsection{Courses}
+\label{sec:data_courses}
 
@@
 
+Courses have different values for credits. The statistic program
+\texttt{stat\_001} lists number of courses with a given value for credits.
+Table~\ref{tab:courses_credits} shows that count.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{| l | r |}
+        \hline
+        Credits & Number of Courses \\ \hline
+        1       & 3                 \\ \hline
+        2       & 3                 \\ \hline
+        3       & 3                 \\ \hline
+        5       & 6                 \\ \hline
+        6       & 46                \\ \hline
+        7.5     & 12                \\ \hline
+        12      & 2                 \\ \hline
+        15      & 2                 \\ \hline
+    \end{tabular}
+
+    \caption
+        [Number of courses group by credits]
+        {Number of courses group by credits.}
+
+    \label{tab:courses_credits}
+\end{table}
+
 \subsection{Moodle Users - Students, and Professors}
+\label{ref:moodle_users}
 
@@
 \subsection{Results}
+\label{sec:results}
 
@@
 \subsection{Moodle Logs}
+\label{sec:logs}
 
@@
 
-% CoursesGeneral.csv
-% CoursesEntriesOverview.csv
-% MoodleUsers.csv
-% Students.csv
-% CoursesStudents.csv
-% Results.csv
-% MoodleLogs.csv
-% CoursesNoMoodle.csv
+The data stored in CSV files from which any statistics are calculated and
+experiments are executed. The data is also organized following a relational
+paradigm, even if the data is not stored in a relational database.
+
+The following sections describe which datasets store the data of the various
+objects as described in section~\ref{sec:objs_ents_feat}. In the end of this
+section, figure\ref{fig:er} shows the relational model of the various datasets.
+
+\subsection{Courses Datasets}
+
+Data on the courses is mostly store in the following dataset files:
+
+\begin{itemize}
+    \item \texttt{CoursesEntriesOverview}
+    \item \texttt{CoursesGeneral}
+    \item \texttt{CoursesNoMoodle}
+    \item \texttt{CoursesStudents}
+\end{itemize}
+
+The first dataset contains mostly what is found in the original
+\texttt{ResultadosUCMoodle} dataset. The second dataset contains general
+information on each course, namely the three codes that identify each course,
+the course name, regime, credits, etc.
+
+The \texttt{CoursesNoMoodle} dataset simply contains a listing of courses that
+have no moodle. These courses aren't very important since no other information
+on them is available. But they are still kept in the processed datasets.
+
+The \texttt{CoursesStudents} dataset contains one entry for each course. Each
+entry states the total number of students of that course, the total number of
+approved students, and the approval ration for that course. This information
+would easily be calculated from other datasets in this repository, but it is
+kept here since it was present in the original datasets.
+
+\subsection{Moodle Users Datasets}
+
+The following three datasets store information on the Moodle users:
+
+\begin{itemize}
+    \item \texttt{MoodleUsers}
+    \item \texttt{Students}
+\end{itemize}
+
+Each Moodle user has a username, an actual name, and a role. That information
+is stored in \texttt{MoodleUsers}. Students have the fields stated in
+section~\ref{ref:moodle_users}, and those are stored in the \texttt{Students}
+dataset.
+
+\subsection{Results and Profiling Datasets}
+
+Results are mostly stored in the \texttt{Results} dataset, but some other
+information is stored in \texttt{Profile}.
+
+The results consist of the fields stated in section~\ref{sec:results}.
+
+\TODO{Talk about the profile dataset}
+
+\subsection{Moodle Logs Dataset}
+
+The logs dataset in \texttt{MoodleLogs} contain the entries as described in
+section~\ref{sec:logs}.
+
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../ER/er.pdf}
+
+    \caption
+        [Relational model for processed data repository]
+        {Relational model for processed data repository.}
+
+    \label{fig:er}
+\end{figure}

### 2017-02-24 ResearchNotes/introduction.tex
Updated Research Notes

\chapter{Introduction}

The objective of this project is to mine Moodle's data to find if user
activities explains their success. The hypothesis is that some behaviours in
courses during a semester may lead to success or lack there off.

This document contains the research notes made during the development of this
project. It contains a section on background, which has comments done to
papers, a listing of relevant journals and books, and some state of the art
regarding Data Mining, Educational Data Mining, and Moodle Mining.

The document also contains a section dedicate to the data which is used in this
project and a section detailing statistical analysis and experiments done to
that data.

### 2017-02-24 ResearchNotes/papers.tex
Updated Research Notes

@@
 This section contain long comments about the first papers which were explored
-for this project.
+for this project. Some of the comments are structures as a five paragraph essay
+having the sections:
+
+\begin{itemize}
+    \item Introduction
+    \item Exposure
+    \item Arguments for
+    \item Arguments against
+    \item Conclusions
+\end{itemize}
+
+Other simply have one big comment.
+
+As time went on, the necessity for analysing individual papers dropped, so the
+number of papers that have individual comments in here is lesses than the total
+number of papers in the project.
+
+The papers listed in here are about Data Mining done to Moodle data. Each paper
+is identified by the code \texttt{IND\_XXX}, where \texttt{XXX} is a number
+with 3 leading zeros. Each paper is also identified by a reference in the
+references section. As some papers reference each other, these references are
+kept in the comments. The reason to identify a particular paper with a code is
+because it makes it easy to store them in disk due to their long names.
+
+Note that the number in \texttt{IND\_XXX} doesn't imply an order in the papers
+listed in the document. Also, just because there is a paper in
+\texttt{IND\_904} and a paper \texttt{IND\_906}, that doesn't imply that a
+paper \texttt{IND\_905} needs to exists in the document.
 

### 2017-02-24 ResearchNotes/stat_exps.tex
Updated Research Notes

@@
 \section{Statistics}
+
+This section contains simple statistical analysis done to the data to better
+understand it.
 
@@
 \section{Experiments}
+
+This section contains Data Mining tasks done to the data. Each task doesn't
+just use one of the datasets (\texttt{CSV} files), rather, it creates datasets
+from existing ones and also calculates new fields.
+
+\subsection{\texttt{EXP\_001} - \textit{Pending Name}}
+
+\TODO{This experiment isn't finished yet.}
+
+In this experiment datasets are build to represent a student's activity in a
+given course and it's final outcome. The first dataset simply as a binary class
+representing if the student was approved or not. The second has the following
+classes:
+
+\begin{itemize}
+    \item $ [0, 10[ $
+    \item $ [10, 12[ $
+    \item $ [12, 14[ $
+    \item $ [14, 16[ $
+    \item $ [16, 18[ $
+    \item $ [18, 20] $
+\end{itemize}
+
+The datasets used to build the experiment sets are \textit{Students},
+\textit{CoursesGeneral}, \textit{MoodleLogs}, and \textit{Results}. The
+tables~\ref{tab:exp_001} shows the fields of the built datasets, their types,
+and origins. Note that the origin of some fields is \textit{Calculated}, this
+is not a dataset, but rather is used to indicate that that field was especially
+computed for this specific experiment.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{| l | l | l |}
+        \hline
+        \textbf{Field}    & \textbf{Source Dataset}    & \textbf{Type} \\ \hline
+        StudentsNumber    & Student                    & Number        \\ \hline
+        EnrollementCount  & Student                    & Number        \\ \hline
+        StudentType       & Student                    & Enumerated    \\ \hline
+        CourseCodeSiiue   & CoursesGeneral             & String        \\ \hline
+        Semester          & CoursesGeneral             & Enumerated    \\ \hline
+        Season            & CoursesGeneral             & Enumerated    \\ \hline
+        TotalActsOnCourse & Calculated From MoodleLogs & Number        \\ \hline
+        CActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        RActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        UActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        DActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        Class             & Calculated From Results    & Binary        \\ \hline
+    \end{tabular}
+
+    \caption
+        []
+        {}
+
+    \label{tab:exp_001}
+\end{table}
+
+% (A quick first review)
+%
+% The first dataset has a flaw. It groups students and courses together to predicts grades from those courses. That doesn't really make sense....
+%
+% Student's usage data about a single course should predict grades on that course alone.
+%
+% The same goes for clustering. Doesn't make sense to cluster this.
+%
+% Check the number of students per course. If a given course has a lot of students with approvals, run a classification for that course only.
+%
+% Also, resample.
+
+\subsection{\texttt{EXP\_002} - \textit{Pending Name}}
+
+\TODO{This experiment isn't finished yet.}
+
+\subsection{\texttt{EXP\_003} - \textit{Pending Name}}
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{| l | l | l |}
+        \hline
+        \textbf{Field}    & \textbf{Source Dataset}    & \textbf{Type} \\ \hline
+        StudentsNumber    & Student                    & Number        \\ \hline
+        EnrolledCourses   & Calculated From Results    & Number        \\ \hline
+        ApprovedCourses   & Calculated From Results    & Number        \\ \hline
+        TotalActsOnCourse & Calculated From MoodleLogs & Number        \\ \hline
+        CActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        RActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        UActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+        DActsOnCourse     & Calculated From MoodleLogs & Number        \\ \hline
+    \end{tabular}
+
+    \caption
+        []
+        {}
+
+    \label{tab:exp_003}
+\end{table}
+
+\TODO{This experiment isn't finished yet.}

### 2017-02-24 src/data.py

@@ -1,7 +1,8 @@
 import os
 
 
-import pandas
+import pandas as pd
+from sklearn import preprocessing
 
 
 data_root = "../../Data/Processed"
@@ -15,5 +16,265 @@ MoodleUsers            = os.path.join(data_root, "MoodleUsers.csv")
 Results                = os.path.join(data_root, "Results.csv")
 Students               = os.path.join(data_root, "Students.csv")
 
+
 def load(sheet):
-    return pandas.read_csv(sheet)
+    return pd.read_csv(sheet)
+
+
+def makeDataset(approvedBinary=True):
+    df_stu = load(Students)
+    df_cge = load(CoursesGeneral)
+    df_log = load(MoodleLogs)
+    df_res = load(Results)
+
+    # Join Students and results to obtain a dataset with an entry for each
+    # student and actual result, set CourseCodeSiiue to be an index along with
+    # StudentsNumber
+    df_stu.set_index("StudentsNumber", inplace=True)
+    df_res.set_index("StudentsNumber", inplace=True)
+    df_stu_res = df_stu.join(df_res)
+
+    df_stu_res.set_index("CourseCodeSiiue", inplace=True, append=True)
+    df_stu_res = df_stu_res[["EnrollementCount", "StudentType", "Grade"]]
+
+    # Get the logs and add CourseCodeSiiue to them. Then take only the useful
+    # columns
+    df_log = df_log.merge(
+            df_cge[["CourseCodeMoodle", "CourseCodeSiiue"]],
+            on = "CourseCodeMoodle",
+            how = "left") \
+        [["StudentsNumber",
+          # "CourseCodeSiiue",
+          "CourseCodeMoodle",
+          "CRUD",
+          "NumberOfActivitiesPerWeek"]]
+
+    # Perform group-bys in order to get count of CRUD activities
+    df_acts_all = df_log \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
+        .agg("sum")
+
+    df_acts_c = df_log \
+        .loc[lambda df: df["CRUD"] == "c"] \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
+        .agg("sum")
+
+    df_acts_r = df_log \
+        .loc[lambda df: df["CRUD"] == "r"] \
+        .groupby(["StudentsNumber", "CourseCodeMoodle"]) \
+        .agg("sum")
+
+    df_acts_u = df_log \
[…]

### 2017-02-24 src/stat_007.py

@@ -1,0 +1,48 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Get number of students per course and number of approved students per course
+# =============================================================================
+
+import os
+
+import numpy as np
+import pandas
+from matplotlib import pyplot
+
+
+import data
+import util
+
+
+output_dir = util.outputDirName(6)
+
+
+if __name__ == "__main__":
+    # Load data
+    df = data.load(data.Results)[[
+        "CourseCodeSiiue",
+        "StudentsNumber",
+        "Grade"]]
+
+    # Number of students per course
+    g1 = df.groupby("CourseCodeSiiue") \
+        .agg("count") \
+        .sort("StudentsNumber", ascending=False)
+
+    # Number of approved students per course
+    g2 = df \
+        .loc[lambda x: (x["Grade"] >= 10) & (x["Grade"] <= 20)] \
+        .groupby("CourseCodeSiiue").agg("count") \
+        .sort("StudentsNumber", ascending=False)
+
+    print(g1)
+    print(g2)
+
+    g1_values = g1[["StudentsNumber"]].values.flatten()
+    g2_values = g2[["StudentsNumber"]].values.flatten()
+
+    pyplot.bar(range(len(g1_values)), g1_values, width=1.0, facecolor="lightgray", edgecolor="lightgray")
+    pyplot.bar(range(len(g2_values)), g2_values, width=1.0, facecolor="blue", edgecolor="blue")
+
+    pyplot.show()

### 2017-02-28 ResearchNotes/books.tex
Added info on books

\section{Books}

Here is a listing of books which were relevant for this project. Each section
describes a single book, and how that particular book was relevant in this
project.

\subsection{``Machine Learning''}

This book is~\cite{book_ml} and is titled ``Machine Learning''. Although this
is a Data Mining project, many of the algorithms and methods exposed by this
book are very relevant in Data Mining, namely chapter 3 which is on Decision
Trees and chapter 8 which is on Instance Based Learning.

\subsection{``Data Mining and Knowledge Discovery Handbook''}

Book is~\cite{book_dm_kdh}.~\TODO{}

\subsection{``Introduction to Data Mining''}

Book is~\cite{book_intro_dm}.~\TODO{}

\subsection{``Data Mining: Practical Machine Learning Tools and Techniques''}

Book is~\cite{book_dm_practical}.~\TODO{}

### 2017-03-01 ResearchNotes/books.tex
Adds papers and books descriptions to notes

@@
 
-Book is~\cite{book_dm_practical}.~\TODO{}
+This Book~\cite{book_dm_practical} is very practical not only because it talk
+about many models, but also because is talk about how Machine Learning and data
+Mining relate. Such notion might be important in the thesis text or, at the
+very least, it might be important is someone asks why I cite sources about
+Machine Learning.
+
+The usual classifier models, such as decision trees, and usual clustering
+methods are specified in this book.

### 2017-03-01 ResearchNotes/papers.tex
Adds papers and books descriptions to notes

@@
-\section{Initial Papers in Detail}
+\section{Papers}
 
@@
 presented and discussed. The results seemed to be quite good.
+
+% =============================================================================
+\subsection{\texttt{IND\_010} - E-learning: emerging uses, empirical results
+and future directions}
+
+Paper is~\cite{ind_010}.
+
+What is taken from this paper is how e-learning has been growing in popularity
+as a way to deliver training online. Organizations such as Universities have
+been using E-Learning recently due to advances in technology and the increasing
+use of the Internet.
+
+One other point taken from this paper is how E-Learning platforms are used to
+track users in terms of their usage of the platforms, their habits, etc.
+Studying this data can be useful to the organizations that make use of
+E-Learning.
+
+% =============================================================================
+\subsection{\texttt{IND\_011} - A Learning Management System Including
+Laboratory Experiments on Measurement Instrumentation}
+
+Paper is~\cite{ind_011}.
+
+This paper supports the notion that~\gls{lms} systems are platforms for content
+distribution, and that the content they distribute have authors and learners.
+These two roles are basically the Professors and Students, respectively, in
+Moodle.
+
+Like in Moodle, the authors will create and publish content that is relevant
+for a training/learning process in the scope of college courses or any other
+E-Learning scenario. The learners will consume that content when performing
+their training or when studying courses.
+
+% =============================================================================
+\subsection{\texttt{IND\_012} - M-Learning - a New Stage Of E-Learning}
+
+Paper is~\cite{ind_012}.
+
+This paper introduces the notion of M-Learning, with the M standing for mobile.
+M-Learning is like E-Learning but aims at delivering content and training to
+mobile devices which are connected to the Internet.
+
+M-Learning is presented as a subset of E-Learning, that is, a more limited
+notion of E-Learning, since platforms for it would be developed aiming for
+tablet PCs, mobile phones, and other such devices. Content would also be
+produced, organized, and published in such a way as to fit the devices in
+question and the usage of those devices.
+
+This paper doesn't really have anything relevant for this project, and it also
+doesn't cite any relevant sources for it.
+
+% =============================================================================
+\subsection{\texttt{IND\_013} - Can E-Learning Replace Classroom Learning?}
+
+Paper is~\cite{ind_013}.
+
+A comparison between traditional classroom learning and E-Learning is made in
+this paper. The most relevant thing for this project is the fact that the paper
+supports the notion that E-Learning as been on the rise for life-long learning,
+on-the-job learning, and distance learning.
+
+% =============================================================================
+\subsection{\texttt{IND\_014} - Moodle: Using learning communities to create an
+open source course management system}
+
+Paper is~\cite{ind_014}.
+
+The paper describes PhD research about data generated in a Moodle instance in
+the first few experiments made when the platform was starting out to be
+developed and used.
+
+The main question the paper looks into the answering is following question:
+
+``How can internet software successfully support social constructionist
+epistemologies of teaching and learning?''
+
+It doesn't go into much detail regarding the actual data that Moodle stores,
+but it talks about the design of Moodle, which is a~\gls{lms}, even if it
+doesn't use that term.
+
+The kind of data that the study looks into is relevant in terms of building a
+successful~\gls{lms}, but none of it is relevant in terms of~\gls{edm}.
+\gls{edm} is about the mining of data generated by a~\gls{lms} in order to find
+patterns in training methods. This study is about looking into the usage of the
+actual platform itself.

### 2017-03-10 ResearchNotes/stat_exps.tex
Adds new statistic

@@
 
+\subsection{\texttt{STAT\_007} - Number of Students per Course and Number of
+Approved Students per Course}
+
+We plot what the title states. The grey bars show the number of students
+overall. The blue bars show the number of approved students. Each bar is one
+course. Figure~\ref{fig_stat_007} shows the results of this statistics.
+
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../src/stat_007_results/plot}
+
+    \caption
+        [Number of Students per Course and Number of Approved Students per
+        Course]
+        {Number of Students per Course and Number of Approved Students per
+        Course.}
+
+    \label{fig:stat_007}
+\end{figure}
+
 % =============================================================================

### 2017-03-10 src/stat_007.py

@@ -15,7 +15,7 @@ import data
 import util
 
 
-output_dir = util.outputDirName(6)
+output_dir = util.outputDirName(7)
 
 
 if __name__ == "__main__":
@@ -45,4 +45,6 @@ if __name__ == "__main__":
     pyplot.bar(range(len(g1_values)), g1_values, width=1.0, facecolor="lightgray", edgecolor="lightgray")
     pyplot.bar(range(len(g2_values)), g2_values, width=1.0, facecolor="blue", edgecolor="blue")
 
-    pyplot.show()
+    util.makeDir(output_dir)
+    pyplot.xticks([])
+    pyplot.savefig(os.path.join(output_dir, "plot"))

### 2017-03-12 src/exp_001.py
Adds experiments code

@@ -1,135 +1,63 @@
 #!/usr/bin/env python3
 
 # =============================================================================
-# TOOD
+# TODO
 # =============================================================================
 
 import os
 
 import numpy as np
 import pandas as pd
-from matplotlib import pyplot
+from sklearn import tree
+from sklearn import metrics
+from sklearn.naive_bayes import MultinomialNB
+from sklearn.model_selection import KFold
 
 import data
 import util
 
 
-def makeDataset(approvedBinary=True):
-    df_stu = data.load(data.Students)
-    df_cge = data.load(data.CoursesGeneral)
-    df_log = data.load(data.MoodleLogs)
-    df_res = data.load(data.Results)
+def execute(model, classes_arg):
+    # Get dataset and make the necessary numpy arrays
+    features, classes = data.dat_001(classes=classes_arg)
 
-    # Join Students and results to obtain a dataset with an entry for each
-    # student and actual result, set CourseCodeSiiue to be an index along with
-    # StudentsNumber
-    df_stu.set_index("StudentsNumber", inplace=True)
-    df_res.set_index("StudentsNumber", inplace=True)
-    df_stu_res = df_stu.join(df_res)
+    # Create folds for training and testing
+    kf = KFold(n_splits=10)
+    kf.get_n_splits(features)
 
-    df_stu_res.set_index("CourseCodeSiiue", inplace=True, append=True)
-    df_stu_res = df_stu_res[["EnrollementCount", "StudentType", "Grade"]]
+    # Train and test with different folds
+    for train_index, test_index in kf.split(features):
+        # Fit
+        if model == "tree":
+            clf = tree.DecisionTreeClassifier()
+        elif model == "bayesian":
+            clf = MultinomialNB()
 
-    # Get the logs and add CourseCodeSiiue to them. Then take only the useful
-    # columns
-    df_log = df_log.merge(
-            df_cge[["CourseCodeMoodle", "CourseCodeSiiue"]],
-            on = "CourseCodeMoodle",
-            how = "left") \
-        [["StudentsNumber",
-          # "CourseCodeSiiue",
-          "CourseCodeMoodle",
-          "CRUD",
-          "NumberOfActivitiesPerWeek"]]
[…]

### 2017-03-12 src/exp_002.py

@@ -1,0 +1,46 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# TODO
+# =============================================================================
+
+import os
+
+import numpy as np
+import pandas as pd
+from sklearn import metrics
+from sklearn.cluster import KMeans
+from sklearn.model_selection import KFold
+
+import data
+import util
+
+
+def execute(model):
+    # Get dataset and make the necessary numpy arrays
+    features, _ = data.dat_001()
+
+    # Create folds for training and testing
+    kf = KFold(n_splits=10)
+    kf.get_n_splits(features)
+
+    # Try different number of clusters
+    for cluster_number in range(2, 10):
+        print("Number of Clusters: %s" % (cluster_number))
+
+        # Train and test with different folds
+        for train_index, test_index in kf.split(features):
+            kmeans = KMeans(n_clusters=cluster_number, random_state=0)
+            kmeans.fit(features[train_index])
+
+            # And get it's silhouette score
+            m = metrics.silhouette_score(
+                features[train_index],
+                kmeans.labels_)
+
+            print("\t%0.2f" % (m))
+
+
+if __name__ == "__main__":
+    print("K-Means")
+    execute("kmeans")

### 2017-03-12 src/data.py
Correts data.py

@@ -21,7 +21,7 @@ def load(sheet):
     return pd.read_csv(sheet)
 
 
-def makeDataset(approvedBinary=True):
+def prepare_dat_001(classes):
     df_stu = load(Students)
     df_cge = load(CoursesGeneral)
     df_log = load(MoodleLogs)
@@ -44,7 +44,6 @@ def makeDataset(approvedBinary=True):
             on = "CourseCodeMoodle",
             how = "left") \
         [["StudentsNumber",
-          # "CourseCodeSiiue",
           "CourseCodeMoodle",
           "CRUD",
           "NumberOfActivitiesPerWeek"]]
@@ -111,36 +110,58 @@ def makeDataset(approvedBinary=True):
             how="left")
 
     # Map the class
-    def mapBinary(x):
-        if x >= 10 and x <= 20:
-            return True
-        return False
-
     def mapClasses(x):
-        if x >= 10 and x < 12:
-            return "10-12"
-        elif x >= 12 and x < 14:
-            return "12-14"
-        elif x >= 14 and x < 16:
-            return "14-16"
-        elif x >= 16 and x < 18:
-            return "16-18"
-        elif x >= 18 and x <= 20:
-            return "18-20"
-        else:
-            return "failed"
+        if classes == "binary":
+            if x >= 10 and x <= 20:
+                return True
+            return False
 
-    if approvedBinary:
-        df_final["Grade"] = df_final["Grade"].map(mapBinary)
+        if classes == "2":
+            if x >= 10 and x < 12:
+                return "10-12"
+            elif x >= 12 and x < 14:
+                return "12-14"
+            elif x >= 14 and x < 16:
+                return "14-16"
+            elif x >= 16 and x < 18:
+                return "16-18"
+            elif x >= 18 and x <= 20:
+                return "18-20"
+            else:
+                return "failed"
+
+        if classes == "4":
+            if x >= 10 and x < 14:
+                return "10-14"
+            elif x >= 14 and x < 18:
+                return "14-18"
+            elif x >= 18 and x <= 20:
+                return "18-20"
+            else:
+                return "failed"
+
+    if classes == "binary":
+        df_final["Grade"] = df_final["Grade"].map(mapClasses)
         df_final.rename(columns={"Grade": "Approved"}, inplace=True)
     else:
         df_final["Grade"] = df_final["Grade"].map(mapClasses)
[…]

### 2017-03-12 src/stat_007.py
Adds stat 008

@@ -36,9 +36,6 @@ if __name__ == "__main__":
         .groupby("CourseCodeSiiue").agg("count") \
         .sort("StudentsNumber", ascending=False)
 
-    print(g1)
-    print(g2)
-
     g1_values = g1[["StudentsNumber"]].values.flatten()
     g2_values = g2[["StudentsNumber"]].values.flatten()
 

### 2017-03-12 src/stat_008.py

@@ -1,0 +1,54 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# Get number of enrolled courses and approved courses per student
+# =============================================================================
+
+import os
+
+import numpy as np
+import pandas as pd
+from matplotlib import pyplot
+
+
+import data
+import util
+
+
+output_dir = util.outputDirName(8)
+
+
+def displayStats(label, series):
+    print(label)
+    print("\tMin %d" % (series.min()))
+    print("\tMax %d" % (series.max()))
+    print("\tMean %.02f" % (series.values.mean()))
+    print("\tMedian %.02f" % (np.median(series.values)))
+    print("\tTotal %d" % (len(series)))
+
+
+if __name__ == "__main__":
+    # Load data
+    df = data.load(data.Results)[["StudentsNumber", "Results"]]
+
+    # Number of enrolled courses
+    enrolled = df.groupby("StudentsNumber") \
+        .agg("count") \
+        .rename(columns={"Results": "EnrolledCount"})
+
+    # Number of approved courses
+    approved = df \
+        .loc[lambda x: x["Results"] == "Aprovado"] \
+        .groupby("StudentsNumber") \
+        .agg("count") \
+        .rename(columns={"Results": "ApproveCount"})
+
+    displayStats("Enrolled", enrolled[["EnrolledCount"]])
+    displayStats("Approved", approved[["ApproveCount"]])
+
+    pyplot.boxplot([enrolled[["EnrolledCount"]].values,
+                    approved[["ApproveCount"]].values
+        ], vert=False)
+
+    util.makeDir(output_dir)
+    pyplot.savefig(os.path.join(output_dir, "plot"))

### 2017-03-12 src/data.py
Finishes a few experiments and makes changes to src

@@ -1,4 +1,5 @@
 import os
+import math
 
 
 import pandas as pd
@@ -116,7 +117,7 @@ def prepare_dat_001(classes):
                 return True
             return False
 
-        if classes == "2":
+        elif classes == "2":
             if x >= 10 and x < 12:
                 return "10-12"
             elif x >= 12 and x < 14:
@@ -130,7 +131,7 @@ def prepare_dat_001(classes):
             else:
                 return "failed"
 
-        if classes == "4":
+        elif classes == "4":
             if x >= 10 and x < 14:
                 return "10-14"
             elif x >= 14 and x < 18:
@@ -196,7 +197,7 @@ def dat_001(classes=None):
             df[["GradeClass"]].values)
 
 
-def makeDataset2():
+def prepare_dat_002():
     df_stu = load(Students)
     df_log = load(MoodleLogs)
     df_res = load(Results)
@@ -298,4 +299,33 @@ def makeDataset2():
     df_merged["ActSumU"] = df_merged["ActSumU"].map(removeNanFloat)
     df_merged["ActSumD"] = df_merged["ActSumD"].map(removeNanFloat)
 
-    return df_merged
+    return df_merged.drop(["StudentsNumber"], axis=1)
+
+
+def dat_002(classes=None, approvedJoin=False):
+    # Parse arguments
+    arguments = [None, "2", "4", "6"]
+
+    if classes is not None and classes not in arguments:
+        raise Exception("Argument no in " + str(arguments))
+
+    df = prepare_dat_002()
+
+    def mapClasses(x):
+        if x == 12:
+            x = 11
+        iclass = 13 // int(classes)
+        return x // math.floor(iclass)
+
+    if classes is not None:
+        df["ApprovedCourses"] = df["ApprovedCourses"].map(mapClasses)
+
+    # Approved field will be just another features. There will be no class
+    if approvedJoin:
+        return (df.values, None)
+
+    # Approved field will be the class of the dataset
+    else:
+        return (
+            df.drop(["ApprovedCourses"], axis=1).values,
+            df[["ApprovedCourses"]].values)

### 2017-03-12 src/exp_001.py

@@ -4,16 +4,9 @@
 # TODO
 # =============================================================================
 
-import os
-
-import numpy as np
-import pandas as pd
-from sklearn import tree
-from sklearn import metrics
-from sklearn.naive_bayes import MultinomialNB
-from sklearn.model_selection import KFold
 
 import data
+import tasks
 import util
 
 
@@ -45,19 +38,19 @@ def execute(model, classes_arg):
 
 if __name__ == "__main__":
     print("Decision Tree Classifier, Binary Class")
-    execute("tree", "binary")
+    tasks.supervised(data.dat_001("binary"), "tree")
 
     print("Decision Tree Classifier, Grades Classes 2 by 2")
-    execute("tree", "2")
+    tasks.supervised(data.dat_001("2"), "tree")
 
     print("Decision Tree Classifier, Grades Classes 4 by 4")
-    execute("tree", "4")
+    tasks.supervised(data.dat_001("4"), "tree")
 
     print("Bayesian Network, Binary Class")
-    execute("bayesian", "binary")
+    tasks.supervised(data.dat_001("binary"), "bayesian")
 
     print("Bayesian Network, Grades Classes 2 by 2")
-    execute("bayesian", "2")
+    tasks.supervised(data.dat_001("2"), "bayesian")
 
     print("Bayesian Network, Grades Classes 4 by 4")
-    execute("bayesian", "4")
+    tasks.supervised(data.dat_001("4"), "bayesian")

### 2017-03-12 src/exp_003.py

@@ -1,0 +1,36 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# TODO
+# =============================================================================
+
+
+import data
+import tasks
+import util
+
+
+if __name__ == "__main__":
+    print("Decision Tree Classifier, Original Classes")
+    tasks.supervised(data.dat_002(approvedJoin=False), "tree")
+
+    print("Decision Tree Classifier, 2 Classes")
+    tasks.supervised(data.dat_002(classes="2", approvedJoin=False), "tree")
+
+    print("Decision Tree Classifier, 4 Classes")
+    tasks.supervised(data.dat_002(classes="4", approvedJoin=False), "tree")
+
+    print("Decision Tree Classifier, 6 Classes")
+    tasks.supervised(data.dat_002(classes="6", approvedJoin=False), "tree")
+
+    print("Bayesian Network, Original Classes")
+    tasks.supervised(data.dat_002(approvedJoin=False), "bayesian")
+
+    print("Bayesian Network, 2 Classes")
+    tasks.supervised(data.dat_002(classes="2", approvedJoin=False), "bayesian")
+
+    print("Bayesian Network, 4 Classes")
+    tasks.supervised(data.dat_002(classes="4", approvedJoin=False), "bayesian")
+
+    print("Bayesian Network, 6 Classes")
+    tasks.supervised(data.dat_002(classes="6", approvedJoin=False), "bayesian")

### 2017-03-12 src/exp_004.py

@@ -12,4 +12,4 @@ import util
 
 if __name__ == "__main__":
     print("K-Means")
-    tasks.unsupervised(data.dat_001())
+    tasks.unsupervised(data.dat_002(approvedJoin=False))

### 2017-03-12 src/stat_008.py

@@ -43,12 +43,53 @@ if __name__ == "__main__":
         .agg("count") \
         .rename(columns={"Results": "ApproveCount"})
 
+    # Group of enrolled
+    group_enrolled = enrolled.reset_index() \
+        .groupby("EnrolledCount") \
+        .agg("count")
+
+    # Group of approved
+    group_approved = approved.reset_index() \
+        .groupby("ApproveCount") \
+        .agg("count")
+
+    # Display results
     displayStats("Enrolled", enrolled[["EnrolledCount"]])
     displayStats("Approved", approved[["ApproveCount"]])
 
-    pyplot.boxplot([enrolled[["EnrolledCount"]].values,
+    # Plot graphs
+    def prepareForPlot(values, max_len):
+        ret = []
+
+        cur_index = 0
+        while cur_index < len(values):
+            if values[cur_index][0] > len(ret):
+                ret.append(0)
+                continue
+
+            ret.append(values[cur_index][1])
+            cur_index += 1
+
+        while len(ret) < max_len:
+            ret.append(0)
+
+        return ret
+
+    group_enrolled = prepareForPlot(group_enrolled.reset_index().values, 13)
+    group_approved = prepareForPlot(group_approved.reset_index().values, 13)
+
+    fig, p1 = pyplot.subplots()
+
+    p1.plot(range(13), group_enrolled)
+    p1.plot(range(13), group_approved)
+
+    p2 = p1.twinx()
+
+    p2.boxplot([enrolled[["EnrolledCount"]].values,
                     approved[["ApproveCount"]].values
         ], vert=False)
 
+
     util.makeDir(output_dir)
+    fig.tight_layout()
     pyplot.savefig(os.path.join(output_dir, "plot"))

### 2017-03-12 src/tasks.py

@@ -1,0 +1,56 @@
+from sklearn import tree
+from sklearn import metrics
+from sklearn.naive_bayes import MultinomialNB
+from sklearn.cluster import KMeans
+from sklearn.model_selection import KFold
+
+
+def supervised(data, model):
+    # Get dataset and make the necessary numpy arrays
+    features, classes = data
+
+    # Create folds for training and testing
+    kf = KFold(n_splits=10)
+    kf.get_n_splits(features)
+
+    # Train and test with different folds
+    for train_index, test_index in kf.split(features):
+        # Fit
+        if model == "tree":
+            clf = tree.DecisionTreeClassifier()
+        elif model == "bayesian":
+            clf = MultinomialNB()
+
+        clf.fit(features[train_index], classes[train_index])
+
+        # And get it's accuracy
+        m = metrics.accuracy_score(
+            classes[test_index],
+            clf.predict(features[test_index]))
+
+        print("\t%0.2f" % (m))
+
+
+def unsupervised(data):
+    # Get dataset and make the necessary numpy arrays
+    features, _ = data
+
+    # Create folds for training and testing
+    kf = KFold(n_splits=10)
+    kf.get_n_splits(features)
+
+    # Try different number of clusters
+    for cluster_number in range(2, 10):
+        print("Number of Clusters: %s" % (cluster_number))
+
+        # Train and test with different folds
+        for train_index, test_index in kf.split(features):
+            kmeans = KMeans(n_clusters=cluster_number, random_state=0)
+            kmeans.fit(features[train_index])
+
+            # And get it's silhouette score
+            m = metrics.silhouette_score(
+                features[train_index],
+                kmeans.labels_)
+
+            print("\t%0.2f" % (m))

### 2017-03-14 ThesisText/data.tex
Wrote data chapter

--- 2017-02-10 data.tex
@@
-\chapter{Data}
+\chapter{University of Évora's Moodle Data}
 
@@
 like Python with Pandas, they were exported to a CSV format. Some
-transformations may be applied to these files in other to clean useless points
-or make the data easier to process.
+transformations were applied to these files in other to clean useless data
+points or make the data easier to process.
 
-To begin, each sheet is analysed individually. Conclusions on each list are
-registered in the fist section of this chapter.
+The provided sheets contain information of courses, students, their activities
+in Moodle, and their results. The data comes from Moodle databases and from the
+SIIUE system. Both systems are integrated, making Moodle plus SIIUE the
+University's own~\gls{lms}.
 
-\subsection{Sheet 1 - \texttt{UCs\_SIIUE\_Moodle.csv}}
-
-This sheet lists a few courses (referred to in their native name as
-\textit{Unidades Curriculares}). Some courses have a Moodle page while others
-do not.
-
-The sheets contain four non blank columns. The first and second column are a
-list of courses represented has their Course Id and Moodle Id, which is
-different.
-
-The third column is just a calculated copy of the first. It doesn't add
-anything and it might have been a loose processing step at some point.
-
-The fourth column contains a list of course with no Moodle and therefore no
-Moodle Id.
-
-Only the following columns are relevant:
+The sheets are the following:
 
 \begin{itemize}
-    \item \texttt{Cod UC} - Course's code
-    \item \texttt{No moodle} - Course's code in the Moodle
+    \item \texttt{UCs\_SIIUE\_Moodle.csv}
+    \item \texttt{complete\_rate.csv}
+    \item \texttt{completed.csv}
+    \item \texttt{Notas UCS SIIUE.csv}
+    \item \texttt{Profile}
+    \item \texttt{ResultadosUCMoodle}
+    \item \texttt{Correlação}
+    \item \texttt{Alunos SIIUE E-L}
+    \item \texttt{AlunosMoodle}
+    \item \texttt{export\_weekly}
+    \item \texttt{export\_activity}
 \end{itemize}
 
-\subsection{Sheet 2 - \texttt{complete\_rate.csv}}
+Some sheets contained information that was the result of calculations from the
+established data. Analysis like these were removed from the data for two
+reasons. First, some analysis are not useful so them being in the data only
+makes the data more complex. By removing such analysis we get a simpler data
+model. Second, even if some analysis are in fact useful, they are still easily
+reproduced by the Python tools, with the difference that they are only
+reproduced when needed for specific experiments.
 
-This sheet contain a listing of the completion rate of each course. The list
-has tree columns. A column for course id in the Moodle (\texttt{c\_moodleid}),
-the number of students in that course (\texttt{n\_students}), and the
-completion rate of those courses (\texttt{complete\_rate}).
+Some analysis were made and are described in section~\ref{set:stats}
 
-So a course with 100 students having a completion rate of 75\% means that 75
-students actually completed that course.
-
-\subsection{Sheet 3 - \texttt{completed.csv}}
-
-Contains a listing of which student was approved at which course. The sheet has
-three columns. The first column is the course id of the Moodle
-(\texttt{c\_moodleid}), the second is the user id of the Moodle
-(\texttt{u\_moodleid}), and the third is a column (\texttt{completed}) which
-has the value 1 if the student completed that course, or 0 if not.
-
-The user id may be assigned to a professor, but in this sheet every user id is
-of a student.
-
-\subsection{Sheet 4 - \texttt{Notas UCS SIIUE.csv}}
-
-There is a lot of information in this sheet.
-
-\TODO{Explore}
-
-\subsection{Sheet 5 - \texttt{Profile}}
-
-This sheet contains some analysis made to the data. To begin with, there are
-the following columns:
+The sheets that remained after removing analysis data are:
 
 \begin{itemize}
-    \item \texttt{name\_c} - Name of the course
-    \item \texttt{name\_s} - Name of the student
-    \item \texttt{id\_c} - Id of the course
-    \item \texttt{n\_resrc\_c} - Number of resources in the course
-    \item \texttt{n\_actv\_c} - Number of activities in the course
-    \item \texttt{id\_s} - Id of the student
-    \item \texttt{n\_cons\_s} - \TODO{}
-    \item \texttt{n\_subm\_s} - \TODO{}
-    \item \texttt{result\_s} - Results of that student in that course
+    \item \texttt{UCs\_SIIUE\_Moodle.csv}
+    \item \texttt{Notas UCS SIIUE.csv}
+    \item \texttt{Profile}
+    \item \texttt{ResultadosUCMoodle}
+    \item \texttt{Alunos SIIUE E-L}
+    \item \texttt{AlunosMoodle}
 \end{itemize}
 
-In order for a student to get approved, some creteria must be followed.
-
-\subsection{Sheet 6 - \texttt{ResultadosUCMoodle}}
-
-This sheet contains a lot of information about the courses. To begin with, we
-get a relation between the three ways of identifying a course in the data. A
-course is identified by the Course's Code (\texttt{Cod Siiue}), Course's Code
-in the Moodle (\texttt{Cod Moodle}), or by a course id (\texttt{course}). The
-course id is an integer number, while the codes contain information about the
-department.
-
-The department (\texttt{Dep}) is calculate from the course's code (\texttt{Cod
-Siiue}) by taking the first three characters from that course. So a course with
-code \texttt{INF7185\_1} will be from the department \texttt{INF}.
-
-The sheet presents the first and last register of an activity in the Moodle,
-and also the interval in days from the first activity to the last. The fields
-are \texttt{min}, \texttt{max}, and \texttt{int}.
-
-\TODO{Fields in between}
-
-The sheet finally specifies the number of students who enrolled in a course and
-the number of students who were approved on that course.
-
-\begin{itemize}
-    \item \texttt{course} - Course's id
-    \item \texttt{Cod Moodle} - Course's Code in the Moodle
-    \item \texttt{Cod Siiue} - Course's Code
-    \item \texttt{min} - First activity of course
-    \item \texttt{max} - Last activity of course
-    \item \texttt{int} - Interval between first and last activity of course
-    \item \texttt{actsub}
-    \item \texttt{actcons}
-    \item \texttt{nactsubd}
-    \item \texttt{consal}
-    \item \texttt{consdoc}
-    \item \texttt{subal}
-    \item \texttt{subdoc}
-    \item \texttt{nalu}
-    \item \texttt{ndoc}
-    \item \texttt{InscSIIUE} - Number of students who enrolled
-    \item \texttt{Aprov} - Number of students who got approved
-    \item \texttt{Dep} - Department of the course
-\end{itemize}
-
-\subsection{Sheet 7 - \texttt{Correlação}}
-
-This is a sheet that implements the calculation of correlation between the
-variables of the previous sheet \texttt{ResultadosUCMoodle}. The data it self
-is coppied from the previous sheet and a new table to the right of the first
-table is added with the correlation itself.
-
-The data in here doesn't need to be preprocessed to a new sheet, but the notion
-of correlation, which can be calculated using alternative methods, may be used
-later.
-
-\subsection{Sheet 8 - \texttt{Alunos SIIUE E-L}}
-
-There are a few issues with this sheet. First of all, the sheet appears to have
-two tables which are not really related, maybe? Anyway. The first table has a
-student number, his name, and what appears to be a count of it's enrollments.
-It also has a calculated fields stating his/her's first and last name and
-another column with the students number which is a calculated copy of the
-first.
-
-The second table has Moodle usernames. A first and last name, another count
-(which I assume is a count of enrollments but I can't be sure because the field
-simply states ``count'').
-
-Oh! The second number in the first table is used for a calculated value in the
-second table.
-
-So the last three columns of the second table are all calculated values. The
-first appears to fetch a number from the first. It goes like this. If the name
-of the student in the second table is equal to the first and last name of the
-first, then fetch that number. Else, it is an error.
-
-Next, we get a value, is the username in the Moodle contains a number. If not,
-we get an error.
-
-Lastly. We try to get the first calculated number (the fourth column of the
-second table) if it exists. If it doesn't, we try to get the second calculated
-value (the fifth column of the second table). If that doesn't exist too, we
-write 0.
-
-Up to this point I'm not sure what the purpose of this sheet is. I believe it
-gives me a relation between a student in SIIUE and in Moodle, but I have
-serious doubts about the correctiveness of that relation.
-
-After examining the sheet a second, I now believe that the first table has a
-listing o students and their SIIUE internal id or student number. The second
-table simply has usernames of Moodle. So the second table has students and
-professors, while the first one has students only.
-
-About the counts. The count of the first table is indeed a number of courses
-where the students is enrolled in. The count of the second table is the total
-number of log entries which is a copy (a copy-paste values only copy) of the
-information in the \texttt{AlunosMoodle} sheet.
-
-I don't think this sheet alone is much useful. But it did give insight on how
-this data repository is organized.
-
-\subsection{Sheet 9 - \texttt{AlunosMoodle}}
-
-These are the actual logs of the accesses to Moodle. This is a big table that
-contains the following fields:
-
-\begin{itemize}
-    \item \texttt{c\_id} - Course ID in the Moodle
-    \item \texttt{c\_cod} - Course ID
-    \item \texttt{c\_role} - Role of the user
-    \item \texttt{c\_user} - User's code
-    \item \texttt{c\_nome} - User's name
-    \item \texttt{c\_uid} - User's ID in the Moodle
-    \item \texttt{crud} - \gls{crud} activity
-    \item \texttt{semana} - Number of the week in which the events happened
-    \item \texttt{n} - Number of time the event happened on that week
-    \item \texttt{numAluno} - Number of the student, which is calculated from
-        elsewhere, even when the user is not a student
-    \item \texttt{SNot} - If the student was approved, or failed, or something
-        else
-\end{itemize}
-
-Basically, we get a list of entries specifying activities in Moodle. Each entry
-contains information on the course, the user, and the activity itself.
-
-For the course information we have course id and course code (Moodle code).
-
-For user, we get the user's role, username, actual name of user, and some
-number.
-
-Finally, we get what kind of CRUD activity it was (if create activity, delete,
-etc), the week in which said activity occurred, and number of times the
-activity occurred in that week. The activities are aggregated by type and week.
-
-This table then contains the fields \texttt{numAluno} and \texttt{SNot}. These
-are calculate fields. The first attempts to take a student's number from the
-username, if possible. The second calculates if that particular student was
-approved in the course for which the entries refers to. Both of these fields
-are useless.
-
-There is a second table. This table gives a username, actual name, and total
-number of activities of that user in the log. It is useful, but it should be in
-it's own sheet. The fields of that table have no name, but in order to be
-consistent with the table on the left, we get:
-
-\begin{itemize}
-    \item \texttt{c\_user}
-    \item \texttt{c\_nome}
-    \item \texttt{count}
-\end{itemize}
-
-\subsection{Sheet 10 - \texttt{CalculosAlunos}}
-
-I'm not sure what his sheet is. It has a few values copied from other tables,
-has ``\textit{Calculos}'', on the name, but it doesn't really calculate
-anything. All the values are literal.
-
-It may be discarded.
-
-\subsection{Sheet 11 - \texttt{export\_weekly}}
-
-This table seems to have a line for each pain of course and user ids. It then
-has a few columns of log activities. For each intersection of pair of course
-and user ids we get a 1 or 0 for one of the columns. Presumably, we get a true,
-that is, 1, if a user A made that activity in course B in that week.
-
-This isn't really useful because the same can be easily calculated using other
-tools when needed.
-
-\subsection{Sheet 12 - \texttt{export\_activity}}
-
-This table seems to have a few more calculations, but I don't fully understand
-them. It has at least a calculation of a standard deviation, and little else.
-
-Again, this isn't really useful because the same can be easily calculated using
-other tools when needed.
+The rest of the sheets contain the objects which are going to be used in the
+Data Mining tasks. Those objects and their features are fairly unorganized in
+the original data, so an analysis on of the available features had to be done.
+Section~\ref{sec:objs_feat} states the objects found in the data and details
+their features. Section~\ref{sec:preproc_dataset} shows how the final dataset
+is organized.
 
 % =============================================================================
-\section{Objects, Entities, and Features}
+\section{Objects and Features}
+\label{sec:objs_feat}
 
-After a first analysis of the data, the objects in the data are identified.
-These objects are the courses, students, results, etc. Data mining tasks are
-applied over these objects which are described by their features.
+After a first analysis of the data, the objects are identified. These objects
+are courses, students, results, etc. Data mining tasks are applied over these
+objects which are described by their features.
 
@@
 \subsection{Courses}
+\label{sec:data_courses}
 
 In the data, a course is identified by one of three unique ids. They are, the
-Course ID, SIIUE's Course's Code, and Moodle Course's Code. The Course id is
+Course Id, SIIUE's Course's Code, and Moodle Course's Code. The Course Id is
 simply a unique positive integer number which is probably assigned as an
@@
 
-Table~\ref{tab:course_code_example} shows an example of the three ids.
-Notice how \texttt{CourseCodeMoodle} is either the same as \texttt{CourseCodeSiiue} or has just minimal differences (in this case there is an underscore follwed by a number).
+Table~\ref{tab:course_code_example} shows an example of the three ids. Notice
+how \texttt{CourseCodeMoodle} is either the same as \texttt{CourseCodeSiiue} or
+has just minimal differences (in this case there is an underscore followed by a
+number).
 
@@
 are enumerated types because they have a limited domain. For example, the
-semester may either be ``\textit{Par}'' or ``\textit{Ímpar}'' (Even or Odd),
-the season may be ``\textit{Normal}'' or ``\textit{Special}''.
+semester may either be ``\textit{Par}'' or ``\textit{Ímpar}'' (Even or Odd).
+The type only has one possible value which is \texttt{Normal}. Despite this,
+the field was kept in the data for completeness. Table~\ref{tab:courses_enum}
+shows the domain of each enumerated type.
 
@@
 
-Some courses have no Moodle. They are therefore useless for our these
-experiments.
-
-Some statistics are taken from the courses data as shown in
-table~\ref{tab:courses_stats}.
-
-The domain of the enumerated fields are listed in table~\ref{tab:courses_enum}.
-The field ``type'' only has the domain ``\textit{Normal}'', so it is a field
-which is useless for further analysis. The field ``Semester'' has a missing
-values. In the actual data there is only one course which doesn't have a value
-for this field.
-
-\begin{table}[h!]
-    \centering
-
-    \begin{tabular}{| l | r |}
-        \hline
-        Number of Courses     & 77 \\ \hline
-        Number of Departments & 9  \\ \hline
-        Number of Degrees     & 11 \\ \hline
-    \end{tabular}
-
-    \caption
-        [Statistics on the Courses]
-        {Statistics on the Courses.}
-
-    \label{tab:courses_stats}
-\end{table}
+Some courses have no Moodle and, as such, have no other associated information.
+They are therefore useless for these experiments and were simply removed.
 
@@
         Regime   & \{ A, O, S \}                            \\ \hline
-        Semester & \{ ``\textit{Par}'', ``\textit{Ímpar}'', n/a \}  \\ \hline
-        Season   & \{ ``\textit{Normal}'', ``\textit{Especial}'' \} \\ \hline
-        Type     & \{ ``\textit{Normal}'' \}                        \\ \hline
+        Semester & \{ \textit{Par}, \textit{Ímpar}, n/a \}  \\ \hline
+        Season   & \{ \textit{Normal}, \textit{Especial} \} \\ \hline
+        Type     & \{ \textit{Normal} \}                    \\ \hline
     \end{tabular}
@@
 \subsection{Moodle Users - Students, and Professors}
+\label{ref:moodle_users}
 
@@
 
-The values for the \textit{StudentType} field are:
+The Name field contains the actual name of the student. This information is
+useless for any analysis, but the name was kept for completeness.
 
-\begin{itemize}
-    \item \textit{Normal}
-    \item \textit{Interno}
-\end{itemize}
+The values for the \textit{StudentType} field are \textit{Normal} or
+\textit{Interno} (Normal or Internal).
 
@@
 The data contains a field called \textit{Role} which states if that Moodle user
-is a professor or a student. Supposedly, if the user is a professor the role
-has a value of 5. If it is a student, it has a value of 3. However, we see
-entries for professors with a role value of 3 and vice versa. It is unknown if
-this role actually state the if a user is a professor or student or something
-else.
+is a professor or a student. If the user is a professor the role has a value of
+5. If it is a student, it has a value of 3. A few users with a value of 4 are
+present. These users have the role of editor, but the Moodle logs only have 11
+entries for a user of this role, so they are simply discarded.
 
@@
 
-Table~\ref{tab:moodle_users_stats} shows some stats regarding Moodle users.
-
-\begin{table}[h!]
-    \centering
-
-    \begin{tabular}{| l | r |}
-        \hline
-        Number of Students   & 145 \\ \hline
-        Number of Professors & 87  \\ \hline
-        Total                & 232 \\ \hline
-    \end{tabular}
-
-    \caption
-        [Professor features]
-        {Professor features.}
-
-    \label{tab:moodle_users_stats}
-\end{table}
-
 \subsection{Results}
+\label{sec:results}
 
@@
 
-Table~\ref{tab:results_stats} shows some statistics on the results data.
+\subsection{Profile}
+\label{sec:profile}
 
-\begin{table}[h!]
-    \centering
+Profile objects are objects that, for each pair of course and student, state
+the following indicators:
 
-    \begin{tabular}{| l | r |}
-        \hline
-        Number of results           & 716 \\ \hline
-        Number of non-empty results & 57  \\ \hline
-        Number of approved students & 334 \\ \hline
-    \end{tabular}
+\begin{itemize}
+    \item \texttt{NumberOfResources}
+    \item \texttt{NumberOfActivities}
+    \item \texttt{NumberOfViews}
+    \item \texttt{NumberOfSubmissions}
+\end{itemize}
 
-    \caption
-        [Statistics on the Results]
-        {Statistics on the Results.}
+A resource in this context is any file or link that a professor may have placed
+in a Moodle page. Profile objects state the number of resources which were seen
+by a particular student in a particular course.
 
-    \label{tab:results_stats}
-\end{table}
+An activity is any interaction in a Moodle forum or quizzed. A view is a view
+of any page, resource, etc. A submission is when a students submits a project.
 
 \subsection{Moodle Logs}
+\label{sec:logs}
 
@@
 \section{Preprocessed Datasets}
+\label{sec:preproc_dataset}
 
@@
 
-% CoursesGeneral.csv
-% CoursesEntriesOverview.csv
-% MoodleUsers.csv
-% Students.csv
-% CoursesStudents.csv
-% Results.csv
-% MoodleLogs.csv
-% CoursesNoMoodle.csv
+The data stored in CSV files from which any statistics are calculated and
+experiments are executed. The data is also organized following a relational
+paradigm.~\footnote{Despite the fact that the data is not kept in a Relational
+Database, having it organized in a relational paradigm simplifies the process
+of working with it using the Python tools}
+
+The following sections describe which datasets store the data of the various
+objects as described in section~\ref{sec:objs_feat}. In the end of this
+section, figure~\ref{fig:er} shows the relational model of the various datasets.
+
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../ER/er.pdf}
+
+    \caption
+        [Relational model for processed data repository]
+        {Relational model for processed data repository.}
+
+    \label{fig:er}
+\end{figure}
+
+\subsection{Courses Datasets}
+
+Data on the courses is store in the following dataset files:
+
+\begin{itemize}
+    \item \texttt{CoursesEntriesOverview}
+    \item \texttt{CoursesGeneral}
+    \item \texttt{CoursesNoMoodle}
+    \item \texttt{CoursesStudents}
+\end{itemize}
+
+The first dataset contains mostly what is found in the original
+\texttt{ResultadosUCMoodle} dataset. The second dataset contains general
+information on each course, namely the three codes that identify each course,
+the course name, regime, credits, etc.
+
+The \texttt{CoursesNoMoodle} dataset simply contains a listing of courses that
+have no moodle. These courses aren't very important since no other information
+on them is available. But they are still kept in the processed datasets.
+
+The \texttt{CoursesStudents} dataset contains one entry for each course. Each
+entry states the total number of students of that course, the total number of
+approved students, and the approval ration for that course. This information
+would easily be calculated from other datasets in this repository, but it is
+kept here since it was present in the original datasets as a non calculated
+field.
+
+\subsection{Moodle Users Datasets}
+
+Each Moodle user has a username, an actual name, and a role. That information
+is stored in \texttt{MoodleUsers}. This dataset contains users which are both
+professors and students. They are identified by their role.
+
+Students have the fields stated in section~\ref{ref:moodle_users}, and those
+are stored in the \texttt{Students} dataset.
+
+\subsection{Results}
+
+Results are stored in the \texttt{Results} dataset. The results consist of the
+fields stated in table~\ref{tab:results_fetures} in section~\ref{sec:results}.
+
+\subsection{Profiling Datasets}
+
+The \texttt{Profile} dataset contains the information stated in
+section~\ref{sec:profile}. The fields are shown in
+table~\ref{tab:profile_fields}.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{l l}
+        Feature                      & Data Type \\ \hline
+        \texttt{CourseCodeMoodle}    & String    \\
+        \texttt{CourseId}            & String    \\
+        \texttt{StudentsId}          & Number    \\
+        \texttt{NumberOfResources}   & Number    \\
+        \texttt{NumberOfActivities}  & Number    \\
+        \texttt{NumberOfViews}       & Number    \\
+        \texttt{NumberOfSubmissions} & Number    \\
+    \end{tabular}
+
+    \caption
+        [Profile features]
+        {Profile features.}
+
+    \label{tab:profile_features}
+\end{table}
+
+\subsection{Moodle Logs Dataset}
+
+The logs dataset in \texttt{MoodleLogs} contain the entries as described in
+section~\ref{sec:logs}.

### 2017-03-14 src/exp_001.py
Adds detail to experiment 1

@@ -5,38 +5,14 @@
 # =============================================================================
 
 
+import sys
+
+
 import data
 import tasks
-import util
 
 
-def execute(model, classes_arg):
-    # Get dataset and make the necessary numpy arrays
-    features, classes = data.dat_001(classes=classes_arg)
-
-    # Create folds for training and testing
-    kf = KFold(n_splits=10)
-    kf.get_n_splits(features)
-
-    # Train and test with different folds
-    for train_index, test_index in kf.split(features):
-        # Fit
-        if model == "tree":
-            clf = tree.DecisionTreeClassifier()
-        elif model == "bayesian":
-            clf = MultinomialNB()
-
-        clf.fit(features[train_index], classes[train_index])
-
-        # And get it's accuracy
-        m = metrics.accuracy_score(
-            classes[test_index],
-            clf.predict(features[test_index]))
-
-        print("\t%0.2f" % (m))
-
-
-if __name__ == "__main__":
+def doAll():
     print("Decision Tree Classifier, Binary Class")
     tasks.supervised(data.dat_001("binary"), "tree")
 
@@ -54,3 +30,16 @@ if __name__ == "__main__":
 
     print("Bayesian Network, Grades Classes 4 by 4")
     tasks.supervised(data.dat_001("4"), "bayesian")
+
+
+if __name__ == "__main__":
+    if len(sys.argv) == 1:
+        doAll()
+
+    elif len(sys.argv) == 3:
+        tasks.supervised(data.dat_001(sys.argv[2]), sys.argv[1])
+
+    else:
+        print("Usage:")
+        print("\texp_001.py")
+        print("\texp_001.py [tree | bayesian] [binary | 2 | 4]")

### 2017-03-14 src/exp_001a.py

@@ -1,0 +1,44 @@
+#!/usr/bin/env python3
+
+import sys
+
+import data
+import tasks
+import util
+
+
+if __name__ == "__main__":
+    limit = 100
+
+    if len(sys.argv) == 1:
+        model = "tree"
+
+    elif len(sys.argv) in [2, 3]:
+        model = sys.argv[1]
+
+        if len(sys.argv) == 3:
+            limit = int(sys.argv[2])
+
+    else:
+        print("Usage:")
+        print("\texp_001a.py")
+        print("\texp_001a.py [tree | bayesian] {limit}")
+
+    df = data.dat_001("binary")
+    clf, accuracy = tasks.supervised(df, model)
+
+    # Display True Positives
+    util.title("True Positives")
+    util.printByCondition(lambda _, c, p: c and p[0], df, clf, limit)
+
+    # Display True Negatives
+    util.title("True Negatives")
+    util.printByCondition(lambda _, c, p: not c and not p[0], df, clf, limit)
+
+    # Display False Positives
+    util.title("False Positives")
+    util.printByCondition(lambda _, c, p: c != p[0] and p[0], df, clf)
+
+    # Display False Negatives
+    util.title("False Negatives")
+    util.printByCondition(lambda _, c, p: c != p[0] and not p[0], df, clf)

### 2017-03-14 src/exp_001b.py

@@ -1,0 +1,38 @@
+#!/usr/bin/env python3
+
+import sys
+
+import data
+import tasks
+import util
+
+
+if __name__ == "__main__":
+    limit = 100
+
+    if len(sys.argv) == 1:
+        model = "tree"
+        classes = "2"
+
+    elif len(sys.argv) in [3, 4]:
+        model = sys.argv[1]
+        classes = sys.argv[2]
+
+        if len(sys.argv) == 4:
+            limit = int(sys.argv[2])
+
+    else:
+        print("Usage:")
+        print("\texp_001a.py")
+        print("\texp_001a.py [tree | bayesian] [2 | 4] {limit}")
+
+    df = data.dat_001(classes)
+    clf, accuracy = tasks.supervised(df, model)
+
+    # Display True Positives
+    util.title("Correctly classified")
+    util.printByCondition(lambda _, c, p: c == p[0], df, clf, limit)
+
+    # Display True Negatives
+    util.title("Incorrectly Classified")
+    util.printByCondition(lambda _, c, p: c != p[0], df, clf, limit)

### 2017-03-14 src/exp_002.py
Made experiment 2

@@ -7,9 +7,9 @@
 
 import data
 import tasks
-import util
 
 
 if __name__ == "__main__":
-    print("K-Means")
-    tasks.unsupervised(data.dat_001())
+    for i in range(2, 10):
+        print("K-Means, %s" % (i))
+        tasks.unsupervised(data.dat_001(), i)

### 2017-03-14 src/exp_002a.py

@@ -1,0 +1,51 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# TODO
+# =============================================================================
+
+import sys
+
+import data
+import tasks
+
+
+def getObjects(df, kmeans, cluster, limit):
+    count = 0
+
+    for i in df[0]:
+        res_cluster = kmeans.predict([i])
+
+        if cluster == res_cluster:
+            sys.stdout.write("[");
+            for j in i[:-1]:
+                sys.stdout.write("%.02f, " % (j));
+            sys.stdout.write("%.02f]\n" % (i[-1]));
+
+            count += 1
+
+        if count == limit:
+            return
+
+
+if __name__ == "__main__":
+    limit = 50
+
+    if len(sys.argv) == 1:
+        clusters_num = 2
+    elif len(sys.argv) in [2, 3]:
+        clusters_num = int(sys.argv[1])
+
+        if len(sys.argv) == 3:
+            limit = int(sys.argv[2])
+    else:
+        print("Usage:")
+        print("\texp_002a.py")
+        print("\texp_002a.py clusters_num {limit}")
+
+    df = data.dat_001()
+    (kmeans, silhouette) = tasks.unsupervised(df, clusters_num)
+
+    for i in range(clusters_num):
+        print("Cluster %d" %(i))
+        getObjects(df, kmeans, i, limit)

### 2017-03-15 src/exp_001.py
Adds experiment 3

@@ -43,3 +43,4 @@ if __name__ == "__main__":
         print("Usage:")
         print("\texp_001.py")
         print("\texp_001.py [tree | bayesian] [binary | 2 | 4]")
+        exit(1)

### 2017-03-15 src/exp_001a.py

@@ -23,6 +23,7 @@ if __name__ == "__main__":
         print("Usage:")
         print("\texp_001a.py")
         print("\texp_001a.py [tree | bayesian] {limit}")
+        exit(1)
 
     df = data.dat_001("binary")
     clf, accuracy = tasks.supervised(df, model)

### 2017-03-15 src/exp_003.py

@@ -5,12 +5,13 @@
 # =============================================================================
 
 
+import sys
+
 import data
 import tasks
-import util
 
 
-if __name__ == "__main__":
+def doAll():
     print("Decision Tree Classifier, Original Classes")
     tasks.supervised(data.dat_002(approvedJoin=False), "tree")
 
@@ -34,3 +35,23 @@ if __name__ == "__main__":
 
     print("Bayesian Network, 6 Classes")
     tasks.supervised(data.dat_002(classes="6", approvedJoin=False), "bayesian")
+
+
+if __name__ == "__main__":
+    if len(sys.argv) == 1:
+        doAll()
+
+    elif len(sys.argv) == 3:
+        classes_op = None
+
+        if sys.argv[2] in ["2", "4", "6"]:
+            classes_op = sys.argv[2]
+
+        df = data.dat_002(classes=classes_op, approvedJoin=False)
+        tasks.supervised(df, sys.argv[1])
+
+    else:
+        print("Usage:")
+        print("\texp_003.py")
+        print("\texp_003.py [tree | bayesian] [original | 2 | 4 | 6]")
+        exit(1)

### 2017-03-15 src/exp_003a.py

@@ -9,25 +9,27 @@ import util
 
 if __name__ == "__main__":
     limit = None
+    classes_op = None
 
     if len(sys.argv) == 1:
         model = "tree"
-        classes = "2"
 
     elif len(sys.argv) in [3, 4]:
         model = sys.argv[1]
-        classes = sys.argv[2]
+
+        if sys.argv[2] in ["2", "4", "6"]:
+            classes_op = sys.argv[2]
 
         if len(sys.argv) == 4:
-            limit = int(sys.argv[2])
+            limit = int(sys.argv[3])
 
     else:
         print("Usage:")
-        print("\texp_001a.py")
-        print("\texp_001a.py [tree | bayesian] [2 | 4] {limit}")
+        print("\texp_003a.py")
+        print("\texp_003a.py [tree | bayesian] [original | 2 | 4 | 6] {limit}")
         exit(1)
 
-    df = data.dat_001(classes)
+    df = data.dat_002(classes=classes_op, approvedJoin=False)
     clf, accuracy = tasks.supervised(df, model)
 
     # Display True Positives

### 2017-03-15 src/tasks.py

@@ -14,6 +14,9 @@ def supervised(data, model):
     kf.get_n_splits(features)
 
     # Train and test with different folds
+    best_clf = None
+    best_accuracy = 0
+
     for train_index, test_index in kf.split(features):
         # Fit
         if model == "tree":
@@ -24,14 +27,21 @@ def supervised(data, model):
         clf.fit(features[train_index], classes[train_index])
 
         # And get it's accuracy
-        m = metrics.accuracy_score(
+        accuracy = metrics.accuracy_score(
             classes[test_index],
             clf.predict(features[test_index]))
 
-        print("\t%0.2f" % (m))
+        # Keep best
+        if accuracy > best_accuracy:
+            best_clf = clf
+            best_accuracy = accuracy
+
+        print("\t%0.2f" % (accuracy))
+
+    return (best_clf, best_accuracy)
 
 
-def unsupervised(data):
+def unsupervised(data, cluster_number):
     # Get dataset and make the necessary numpy arrays
     features, _ = data
 
@@ -39,18 +49,24 @@ def unsupervised(data):
     kf = KFold(n_splits=10)
     kf.get_n_splits(features)
 
-    # Try different number of clusters
-    for cluster_number in range(2, 10):
-        print("Number of Clusters: %s" % (cluster_number))
+    # Train and test with different folds
+    best_kmeans = None
+    best_silhouette = 0
 
-        # Train and test with different folds
-        for train_index, test_index in kf.split(features):
-            kmeans = KMeans(n_clusters=cluster_number, random_state=0)
-            kmeans.fit(features[train_index])
+    for train_index, test_index in kf.split(features):
+        kmeans = KMeans(n_clusters=cluster_number, random_state=0)
+        kmeans.fit(features[train_index])
 
-            # And get it's silhouette score
-            m = metrics.silhouette_score(
-                features[train_index],
-                kmeans.labels_)
+        # And get it's silhouette score
+        silhouette = metrics.silhouette_score(
+            features[train_index],
+            kmeans.labels_)
 
-            print("\t%0.2f" % (m))
+        # Keep best
+        if silhouette > best_silhouette:
+            best_kmeans = kmeans
+            best_silhouette = silhouette
+
+        print("\t%0.2f" % (silhouette))
+
+    return (best_kmeans, best_silhouette)

### 2017-03-15 src/exp_004.py
Adds experient 4

@@ -7,9 +7,9 @@
 
 import data
 import tasks
-import util
 
 
 if __name__ == "__main__":
-    print("K-Means")
-    tasks.unsupervised(data.dat_002(approvedJoin=False))
+    for i in range(2, 10):
+        print("K-Means, %s" % (i))
+        tasks.unsupervised(data.dat_002(approvedJoin=False), i)

### 2017-03-15 src/exp_004a.py

@@ -1,0 +1,51 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# TODO
+# =============================================================================
+
+import sys
+
+import data
+import tasks
+
+
+def getObjects(df, kmeans, cluster, limit):
+    count = 0
+
+    for i in df[0]:
+        res_cluster = kmeans.predict([i])
+
+        if cluster == res_cluster:
+            sys.stdout.write("[");
+            for j in i[:-1]:
+                sys.stdout.write("%.02f, " % (j));
+            sys.stdout.write("%.02f]\n" % (i[-1]));
+
+            count += 1
+
+        if count == limit:
+            return
+
+
+if __name__ == "__main__":
+    limit = 50
+
+    if len(sys.argv) == 1:
+        clusters_num = 2
+    elif len(sys.argv) in [2, 3]:
+        clusters_num = int(sys.argv[1])
+
+        if len(sys.argv) == 3:
+            limit = int(sys.argv[2])
+    else:
+        print("Usage:")
+        print("\texp_004a.py")
+        print("\texp_004a.py clusters_num {limit}")
+
+    df = data.dat_002(approvedJoin=False)
+    (kmeans, silhouette) = tasks.unsupervised(df, clusters_num)
+
+    for i in range(clusters_num):
+        print("Cluster %d" %(i))
+        getObjects(df, kmeans, i, limit)

### 2017-03-15 ResearchNotes/stat_exps.tex
Adds information on experiments to research notes

--- 2017-02-12 ResearchNotes/stat_exps.tex
@@
-\chapter{Statistics and Experiments}
+\chapter{Statistics, Datasets, and Experiments}
 
@@
 
+\subsection{\texttt{STAT\_007} - Number of Students per Course and Number of
+Approved Students per Course}
+
+We plot what the title states. The grey bars show the number of students
+overall. The blue bars show the number of approved students. Each bar is one
+course. Figure~\ref{fig_stat_007} shows the results of this statistics.
+
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../src/stat_007_results/plot}
+
+    \caption
+        [Number of Students per Course and Number of Approved Students per
+        Course]
+        {Number of Students per Course and Number of Approved Students per
+        Course.}
+
+    \label{fig:stat_007}
+\end{figure}
+
+% =============================================================================
+\section{Datasets}
+
+\subsection{\texttt{DAT\_001} - Students, Courses, Logs, and Results}
+\label{sec:dat_001}
+
+This dataset, which features is presented in table~\ref{tar:dat_001}, takes
+information from the following datasets:
+
+\begin{itemize}
+    \item \texttt{Students}
+    \item \texttt{CoursesGeneral}
+    \item \texttt{MoodleLogs}
+    \item \texttt{Results}
+\end{itemize}
+
+Each object of the dataset contains information referencing a single student in
+a single course. The objects contain the overall enrollment count and type of
+student. It also contain the identification of a course along with the semester
+and season of that course. Finally, the object contains a count of CRUD
+activities of that student in that course.
+
+The dataset can be generated with different classes. There are three options.
+Option one, the classes is a binary features which has the value true if the
+student was approved on that course, and false otherwise. Option two, the class
+has one of the following values which are intervals of grades:
+
+\begin{itemize}
+    \item $ [0, 10[ $
+    \item $ [10, 12[ $
+    \item $ [12, 14[ $
+    \item $ [14, 16[ $
+    \item $ [16, 18[ $
+    \item $ [18, 20] $
+\end{itemize}
+
+Option three, the class has one of the following values:
+
+\begin{itemize}
+    \item $ [0, 10[ $
+    \item $ [10, 14[ $
+    \item $ [14, 18[ $
+    \item $ [18, 20] $
+\end{itemize}
+
+With option two, we get a dataset which groups every records with grades below
+10, but separates the positive grades in groups of two. Option three does the
+same but separates grades into groups of four, with the last group having only
+three values instead of 4.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{| l | l | l |}
+        \hline
+        \textbf{Field}             & \textbf{Source Dataset}    & \textbf{Type} \\ \hline
+        \texttt{EnrollementCount}  & Student                    & Number        \\ \hline
+        \texttt{StudentType}       & Student                    & Enumerated    \\ \hline
+        \texttt{CourseCodeSiiue}   & CoursesGeneral             & String        \\ \hline
+        \texttt{Semester}          & CoursesGeneral             & Enumerated    \\ \hline
+        \texttt{Season}            & CoursesGeneral             & Enumerated    \\ \hline
+        \texttt{TotalActsOnCourse} & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{CActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{RActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{UActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{DActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{Class}             & Calculated From Results    & Binary        \\ \hline
+    \end{tabular}
+
+    \caption
+        [Dataset 001]
+        {Dataset 001}
+
+    \label{tab:dat_001}
+\end{table}
+
+\subsection{\texttt{DAT\_002} - Results and Logs}
+\label{sec:dat_002}
+
+This dataset only contains features which a number type. Each object is
+calculated from the records of a single student and relates the number of
+enrolled courses, number of approved courses, number of each CRUD activity and
+overall number of activities. The records come from the \texttt{Results} and
+\texttt{MoodleLogs} datasets. Table~\ref{tab_dat_002} shows the features of
+this dataset.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{| l | l | l |}
+        \hline
+        \textbf{Field}           & \textbf{Source Dataset}    & \textbf{Type} \\ \hline
+        \texttt{EnrolledCourses} & Calculated From Results    & Number        \\ \hline
+        \texttt{ApprovedCourses} & Calculated From Results    & Number        \\ \hline
+        \texttt{ActSumAll}       & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{ActSumC}         & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{ActSumR}         & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{ActSumU}         & Calculated From MoodleLogs & Number        \\ \hline
+        \texttt{ActSumD}         & Calculated From MoodleLogs & Number        \\ \hline
+    \end{tabular}
+
+    \caption
+        [Dataset 002]
+        {Dataset 002}
+
+    \label{tab:dat_002}
+\end{table}
+
 % =============================================================================
 \section{Experiments}
+
+This section gives a very brief overview of the experiments run for this
+project. The actual explanation and interpretation of them is written in the
+thesis text.
+
+\subsection{\texttt{EXP\_001} - Predicting Students Grades Based on Moodle
+Usage}
+
+Experiment uses \texttt{DAT\_001}, described in section~\ref{sec:dat_001}.
+
+\begin{verbatim}
+Tree Binary and Bayesian Binary
+Conclusion: More activity -> approved
+
+Tree 4 and Bayesian 4
+Conclusion: More activity -> approved
+Conclusion: More number of CRUD -> better grade
+Conclusion: More updates -> better grades
+
+Tree 2 and Bayesian 2
+Conclusion: More activity -> approved
+Conclusion: More number of CRUD -> better grade
+Conclusion: More updates -> better grades
+\end{verbatim}
+
+\subsection{\texttt{EXP\_002} - Clustering Moodle Usage Data}
+
+Experiment uses \texttt{DAT\_001}, described in section~\ref{sec:dat_001}.
+
+\begin{verbatim}
+Two clusters
+Conclusion: Clusters are found in users with different number of CRUD
+activities. Silhouette score isn't that great
+\end{verbatim}
+
+\subsection{\texttt{EXP\_003} - Predicting Number of Approved Courses Based on
+Moodle Usage}
+
+Experiment uses \texttt{DAT\_002}, described in section~\ref{sec:dat_002}.
+
+\begin{verbatim}
+Tree and Bayesian with original, 2, 4, or 6
+Conclusion: Based on the activity and number of enrolled courses we can
+predict the number of approved courses with great accuracy!
+\end{verbatim}
+
+\subsection{\texttt{EXP\_004} - clustering Number of Approved Courses Based on
+Moodle Usage}
+
+Experiment uses \texttt{DAT\_002}, described in section~\ref{sec:dat_002}.
+
+\begin{verbatim}
+Results are similar to EXP_002
+\end{verbatim}

### 2017-03-15 ResearchNotes/data.tex
Correction in notes

@@
 
-After a first analysis of the data, the objects in the data are identified.
-These objects are the courses, students, results, etc. Data mining tasks are
-applied over these objects which are described by their features.
+After a first analysis of the data, the objects are identified.  These objects
+are the courses, students, results, etc. Data mining tasks are applied over
+these objects which are described by their features.
 
@@
 In the data, a course is identified by one of three unique ids. They are, the
-Course ID, SIIUE's Course's Code, and Moodle Course's Code. The Course id is
+Course Id, SIIUE's Course's Code, and Moodle Course's Code. The Course Id is
 simply a unique positive integer number which is probably assigned as an
@@
 semester may either be ``\textit{Par}'' or ``\textit{Ímpar}'' (Even or Odd),
-the season may be ``\textit{Normal}'' or ``\textit{Special}''.
+the season may be ``\textit{Normal}'' or ``\textit{Especial}'' (Normal or
+Special).
 
@@
 
-Some courses have no Moodle. They are therefore useless for our these
-experiments.
+Some courses have no Moodle. They are therefore useless for these experiments.
 
@@
 
-The values for the \textit{StudentType} field are:
-
-\begin{itemize}
-    \item \textit{Normal}
-    \item \textit{Interno}
-\end{itemize}
+The values for the \textit{StudentType} field are \textit{Normal} or
+\textit{Interno} (Normal or Internal).
 
@@
 
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../ER/er.pdf}
+
+    \caption
+        [Relational model for processed data repository]
+        {Relational model for processed data repository.}
+
+    \label{fig:er}
+\end{figure}
+
 \subsection{Courses Datasets}
@@
 
-\TODO{Talk about the profile dataset}
+The \texttt{Profile} dataset contains the following calculated fields:
+
+\begin{itemize}
+    \item \texttt{NumberOfResources}
+    \item \texttt{NumberOfActivities}
+    \item \texttt{NumberOfViews}
+    \item \texttt{NumberOfSubmissions}
+\end{itemize}
 
@@
 section~\ref{sec:logs}.
-
-\begin{figure}[h!]
-    \centering
-
-    \includegraphics[width=\linewidth]{../ER/er.pdf}
-
-    \caption
-        [Relational model for processed data repository]
-        {Relational model for processed data repository.}
-
-    \label{fig:er}
-\end{figure}

### 2017-03-15 ResearchNotes/introduction.tex
Correction in notes

@@
 
-The objective of this project is to mine Moodle's data to find if user
-activities explains their success. The hypothesis is that some behaviours in
-courses during a semester may lead to success or lack there off.
+This document contains the research notes made during the development of the
+thesis called ``Educational Data Mining applied to Moodle data from the
+University of Évora''.
 
-This document contains the research notes made during the development of this
-project. It contains a section on background, which has comments done to
-papers, a listing of relevant journals and books, and some state of the art
-regarding Data Mining, Educational Data Mining, and Moodle Mining.
+Initially, there are a few comments made to papers and books. Some journals are
+also mentioned.
 
-The document also contains a section dedicate to the data which is used in this
-project and a section detailing statistical analysis and experiments done to
-that data.
+The document has a detailed explanation about the original data and how it was
+processed into a data repository for experiments.
+
+Information about statistics, processed datasets, and experiments are in their
+own section.
+
+Note that this document doesn't have all the information for the thesis. Some
+text in the thesis was copy-pasted from this initial document, but a lot of
+details are only added or corrected in the thesis text.

### 2017-03-16 src/exp_002.py
Adds information to tasks output

@@ -10,6 +10,6 @@ import tasks
 
 
 if __name__ == "__main__":
-    for i in range(2, 10):
+    for i in range(2, 6):
         print("K-Means, %s" % (i))
         tasks.unsupervised(data.dat_001(), i)

### 2017-03-16 src/exp_004.py

@@ -10,6 +10,6 @@ import tasks
 
 
 if __name__ == "__main__":
-    for i in range(2, 10):
+    for i in range(2, 7):
         print("K-Means, %s" % (i))
         tasks.unsupervised(data.dat_002(approvedJoin=False), i)

### 2017-03-16 src/tasks.py

@@ -15,7 +15,9 @@ def supervised(data, model):
 
     # Train and test with different folds
     best_clf = None
-    best_accuracy = 0
+    max_accuracy = 0
+    min_accuracy = 0
+    sum_accuracy = 0
 
     for train_index, test_index in kf.split(features):
         # Fit
@@ -31,14 +33,26 @@ def supervised(data, model):
             classes[test_index],
             clf.predict(features[test_index]))
 
-        # Keep best
-        if accuracy > best_accuracy:
+        # Keep info
+        if best_clf is None:
             best_clf = clf
-            best_accuracy = accuracy
+            max_accuracy = accuracy
+            min_accuracy = accuracy
+
+        elif accuracy > max_accuracy:
+            best_clf = clf
+            max_accuracy = accuracy
+
+        elif accuracy < min_accuracy:
+            min_accuracy = accuracy
+
+        sum_accuracy += accuracy
 
         print("\t%0.2f" % (accuracy))
 
-    return (best_clf, best_accuracy)
+    print("%0.2f, %0.2f, %0.2f" % (max_accuracy, min_accuracy, sum_accuracy / 10))
+
+    return (best_clf, max_accuracy)
 
 
 def unsupervised(data, cluster_number):
@@ -51,7 +65,9 @@ def unsupervised(data, cluster_number):
 
     # Train and test with different folds
     best_kmeans = None
-    best_silhouette = 0
+    max_silhouette = 0
+    min_silhouette = 0
+    sum_silhouette = 0
 
     for train_index, test_index in kf.split(features):
         kmeans = KMeans(n_clusters=cluster_number, random_state=0)
@@ -62,11 +78,24 @@ def unsupervised(data, cluster_number):
             features[train_index],
             kmeans.labels_)
 
-        # Keep best
-        if silhouette > best_silhouette:
+        # Keep info
+        if best_kmeans is None:
             best_kmeans = kmeans
-            best_silhouette = silhouette
+            max_silhouette = silhouette
+            min_silhouette = silhouette
+
+        elif silhouette > max_silhouette:
+            best_kmeans = kmeans
+            max_silhouette = silhouette
+
+        elif silhouette < min_silhouette:
+            best_kmeans = kmeans
+            min_silhouette = silhouette
+
+        sum_silhouette += silhouette
 
         print("\t%0.2f" % (silhouette))
 
-    return (best_kmeans, best_silhouette)
[…]

### 2017-03-16 src/exp_002.py
Adds affinity propagation to tasks

@@ -12,4 +12,9 @@ import tasks
 if __name__ == "__main__":
     for i in range(2, 6):
         print("K-Means, %s" % (i))
-        tasks.unsupervised(data.dat_001(), i)
+        tasks.unsupervised(data.dat_001(), "kmeans", i)
+
+    # Does not get good results and is very slow
+    # for i in range(2, 6):
+    #     print("Affinity Propagation, %s" % (i))
+    #     tasks.unsupervised(data.dat_001(), "affinity", i)

### 2017-03-16 src/exp_004.py

@@ -12,4 +12,8 @@ import tasks
 if __name__ == "__main__":
     for i in range(2, 7):
         print("K-Means, %s" % (i))
-        tasks.unsupervised(data.dat_002(approvedJoin=False), i)
+        tasks.unsupervised(data.dat_002(approvedJoin=False), "kmeans", i)
+
+    for i in range(2, 7):
+        print("Affinity Propagation, %s" % (i))
+        tasks.unsupervised(data.dat_002(approvedJoin=False), "affinity", i)

### 2017-03-16 src/tasks.py

@@ -2,6 +2,7 @@ from sklearn import tree
 from sklearn import metrics
 from sklearn.naive_bayes import MultinomialNB
 from sklearn.cluster import KMeans
+from sklearn.cluster import AffinityPropagation
 from sklearn.model_selection import KFold
 
 
@@ -25,6 +26,8 @@ def supervised(data, model):
             clf = tree.DecisionTreeClassifier()
         elif model == "bayesian":
             clf = MultinomialNB()
+        else:
+            raise Exception("Model %s doesn't exist" % (model))
 
         clf.fit(features[train_index], classes[train_index])
 
@@ -55,7 +58,7 @@ def supervised(data, model):
     return (best_clf, max_accuracy)
 
 
-def unsupervised(data, cluster_number):
+def unsupervised(data, model, cluster_number):
     # Get dataset and make the necessary numpy arrays
     features, _ = data
 
@@ -64,32 +67,39 @@ def unsupervised(data, cluster_number):
     kf.get_n_splits(features)
 
     # Train and test with different folds
-    best_kmeans = None
+    best_clf = None
     max_silhouette = 0
     min_silhouette = 0
     sum_silhouette = 0
 
     for train_index, test_index in kf.split(features):
-        kmeans = KMeans(n_clusters=cluster_number, random_state=0)
-        kmeans.fit(features[train_index])
+        # Fit
+        if model == "kmeans":
+            clf = KMeans(n_clusters=cluster_number, random_state=0)
+        elif model == "affinity":
+            clf = AffinityPropagation()
+        else:
+            raise Exception("Model %s doesn't exist" % (model))
+
+        clf.fit(features[train_index])
 
         # And get it's silhouette score
         silhouette = metrics.silhouette_score(
             features[train_index],
-            kmeans.labels_)
+            clf.labels_)
 
         # Keep info
-        if best_kmeans is None:
-            best_kmeans = kmeans
+        if best_clf is None:
+            best_clf = clf
             max_silhouette = silhouette
             min_silhouette = silhouette
 
         elif silhouette > max_silhouette:
-            best_kmeans = kmeans
+            best_clf = clf
             max_silhouette = silhouette
 
         elif silhouette < min_silhouette:
-            best_kmeans = kmeans
+            best_clf = clf
             min_silhouette = silhouette
 
         sum_silhouette += silhouette
[…]

### 2017-03-20 src/data.py
Adds experiment

@@ -16,6 +16,7 @@ MoodleLogs             = os.path.join(data_root, "MoodleLogs.csv")
 MoodleUsers            = os.path.join(data_root, "MoodleUsers.csv")
 Results                = os.path.join(data_root, "Results.csv")
 Students               = os.path.join(data_root, "Students.csv")
+Profile                = os.path.join(data_root, "Profile.csv")
 
 
 def load(sheet):
@@ -188,13 +189,13 @@ def dat_001(classes=None):
     df["DActsOnCourse"] = df["DActsOnCourse"].map(replaceNan)
 
     if classes == "binary":
-        return (
-            df.drop(["Approved"], axis=1).values,
-            df[["Approved"]].values)
+        class_name = "Approved"
     else:
-        return (
-            df.drop(["GradeClass"], axis=1).values,
-            df[["GradeClass"]].values)
+        class_name = "GradeClass"
+
+    return (
+        df.drop([class_name], axis=1).values,
+        df[[class_name]].values)
 
 
 def prepare_dat_002():
@@ -329,3 +330,101 @@ def dat_002(classes=None, approvedJoin=False):
         return (
             df.drop(["ApprovedCourses"], axis=1).values,
             df[["ApprovedCourses"]].values)
+
+
+def prepare_dat_003(classes):
+    df_stu = load(Students)
+    df_cge = load(CoursesGeneral)
+    df_pro = load(Profile)
+    df_res = load(Results)
+
+    # Drop unnecessary fields from datasets
+    df_stu = df_stu[["StudentsNumber", "StudentsId"]]
+    df_cge = df_cge[["CourseId", "CourseCodeSiiue"]]
+    df_pro = df_pro.drop(["Results"], axis=1)
+
+    # Join Profile with Students and with Courses General to get other students
+    # and courses id's
+    df_pro_ids = df_pro.merge(df_stu, how="left")
+    df_pro_ids = df_pro_ids.merge(df_cge, how="left")
+
+    # Join results
+    df_all = df_pro_ids.merge(df_res, how="left")
+
+    # Keep only useful fields
+    df_final = df_all[[
+            "CourseCodeSiiue",
+            "NumberOfResources",
+            "NumberOfActivities",
+            "NumberOfViews",
+            "NumberOfSubmissions",
+            "Grade"]].copy()
+
+    # Map the class
+    def mapClasses(x):
+        if classes == "binary":
+            if x >= 10 and x <= 20:
+                return True
[…]

### 2017-03-20 src/exp_005.py

@@ -1,0 +1,46 @@
+#!/usr/bin/env python3
+
+# =============================================================================
+# TODO
+# =============================================================================
+
+
+import sys
+
+
+import data
+import tasks
+
+
+def doAll():
+    print("Decision Tree Classifier, Binary Class")
+    tasks.supervised(data.dat_003("binary"), "tree")
+
+    print("Decision Tree Classifier, Grades Classes 2 by 2")
+    tasks.supervised(data.dat_003("2"), "tree")
+
+    print("Decision Tree Classifier, Grades Classes 4 by 4")
+    tasks.supervised(data.dat_003("4"), "tree")
+
+    print("Bayesian Network, Binary Class")
+    tasks.supervised(data.dat_003("binary"), "bayesian")
+
+    print("Bayesian Network, Grades Classes 2 by 2")
+    tasks.supervised(data.dat_003("2"), "bayesian")
+
+    print("Bayesian Network, Grades Classes 4 by 4")
+    tasks.supervised(data.dat_003("4"), "bayesian")
+
+
+if __name__ == "__main__":
+    if len(sys.argv) == 1:
+        doAll()
+
+    elif len(sys.argv) == 3:
+        tasks.supervised(data.dat_003(sys.argv[2]), sys.argv[1])
+
+    else:
+        print("Usage:")
+        print("\texp_005.py")
+        print("\texp_005.py [tree | bayesian] [binary | 2 | 4]")
+        exit(1)

### 2017-03-20 ResearchNotes/stat_exps.tex
Adds notes about new dataset and experiment 5

@@
 
-\subsection{\texttt{DAT\_002} - Results and Logs}
+\subsection{\texttt{DAT\_002} - Number of Courses and Logs}
 \label{sec:dat_002}
@@
 overall number of activities. The records come from the \texttt{Results} and
-\texttt{MoodleLogs} datasets. Table~\ref{tab_dat_002} shows the features of
+\texttt{MoodleLogs} datasets. Table~\ref{tab:dat_002} shows the features of
 this dataset.
@@
     \label{tab:dat_002}
+\end{table}
+
+\subsection{\texttt{DAT\_003} - Student/Course Profile and Results}
+\label{sec:dat_003}
+
+In this dataset we relate the profiling of a student in a courses and his
+results on that course. To begin constructing this dataset we use the Profiling
+dataset, as shown in section~\ref{sec:data_profiling}.
+
+The Profiling dataset contains a results field, however, the meaning of each
+value in this field is unknown. So initially, this field is dropped. To get the
+actual results of student we need the Results dataset.
+
+Each object of Profile contains the features \texttt{CourseCodeMoodle},
+\texttt{CourseId}, and \texttt{StudentsId}. But the Results dataset has
+\texttt{CourseCodeSiiue} and \texttt{StudentsNumber}. In order to relate both,
+the Students dataset and the CoursesGeneral dataset are needed.
+
+These datasets are related and some fields are dropped in order to produce a
+dataset with the features listed in table~\ref{tab:dat_003}.
+
+Like in dataset~\ref{sec:dat_001}, the class of this dataset may be binary, in
+which the student is approved or not, or divided in groups of 2 or 4.
+
+\begin{table}[h!]
+    \centering
+
+    \begin{tabular}{| l | l | l |}
+        \hline
+        \textbf{Field}               & \textbf{Source Dataset} & \textbf{Type} \\ \hline
+        \texttt{CourseCodeSiiue}     & CoursesGeneral          & String        \\ \hline
+        \texttt{NumberOfResources}   & Profile                 & Number        \\ \hline
+        \texttt{NumberOfActivities}  & Profile                 & Number        \\ \hline
+        \texttt{NumberOfViews}       & Profile                 & Number        \\ \hline
+        \texttt{NumberOfSubmissions} & Profile                 & Number        \\ \hline
+        \texttt{Class}               & Results                 & Binary        \\ \hline
+    \end{tabular}
+
+    \caption
+        [Dataset 003]
+        {Dataset 003}
+
+    \label{tab:dat_003}
 \end{table}
@@
 
-\subsection{\texttt{EXP\_004} - clustering Number of Approved Courses Based on
+\subsection{\texttt{EXP\_004} - Clustering Number of Approved Courses Based on
 Moodle Usage}
@@
 \end{verbatim}
+
+\subsection{\texttt{EXP\_005} - Predicting Students Grades From Course
+Profiling}
+
+Experiment uses \texttt{DAT\_003}, described in section~\ref{sec:dat_003}.
+
+\begin{verbatim}
+Results are similar to EXP_001, but performs worse. Meaning that looking into
+Moodle usage data produces more accurate results then course profiling.
+\end{verbatim}

### 2017-03-20 ResearchNotes/stat_exps.tex
Made minor corrections

@@
 value in this field is unknown. So initially, this field is dropped. To get the
-actual results of student we need the Results dataset.
+actual results of students we need the Results dataset.
 
@@
 
-\begin{verbatim}
-Two clusters
-Conclusion: Clusters are found in users with different number of CRUD
-activities. Silhouette score isn't that great
-\end{verbatim}
+This experiment isn't correct because there are nominal fields. Clustering may
+be used in nominal fields, but the algorithms used in there may not.
 
@@
 
-\subsection{\texttt{EXP\_005} - Predicting Students Grades From Course
+\subsection{\texttt{EXP\_005} - Predicting Students Grades From Student/Course
 Profiling}

### 2017-03-21 ResearchNotes/stat_exps.tex
Adds exp 6

@@
 
-\begin{verbatim}
-Results are similar to EXP_002
-\end{verbatim}
+Results are similar to \texttt{EXP\_002}.
 
@@
 
-\begin{verbatim}
-Results are similar to EXP_001, but performs worse. Meaning that looking into
-Moodle usage data produces more accurate results then course profiling.
-\end{verbatim}
+Results are similar to \texttt{EXP\_001}, but performs worse. Meaning that
+looking into Moodle usage data produces more accurate results then course
+profiling.
+
+\subsection{\texttt{EXP\_006} - Clustering Students Grades From Student/Course
+Profiling}
+
+Experiment uses \texttt{DAT\_003}, described in section~\ref{sec:dat_003}.
+
+This experiment drops the SIIUE Courses Code from the dataset. No significant
+results were achieved.

### 2017-03-21 src/data.py

@@ -428,3 +428,9 @@ def dat_003(classes=None):
     return (
         df.drop([class_name], axis=1).values,
         df[[class_name]].values)
+
+
+def dat_004(classes=None):
+    df = prepare_dat_003("binary")
+    df = df.drop(["CourseCodeSiiue"], axis=1)
+    return df.values, []

### 2017-03-22 ThesisText/statistics.tex
Adds statistics and experiments to repository

\chapter{Statistics on the Data}
\label{sec:stats}

Before applying any Data Mining methods, it is important to understand how the
data is constituted. Knowing things like, how many courses are there, how many
students, how many students per course, etc, is important in order to train
models and interpret the results.

\section{Courses}

To begin exploring the courses some counts are made. We count the number of
overall courses, the number of departments, and number of degrees.
Table~\ref{tab:courses_stats} shows these counts. We see 77 courses which are
part of 9 different departments and make up 11 degrees. While the number of
courses is much greater then some experiments mentioned in the background, it
is important to understand how many students does each of these courses have.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Number of Courses     & 77 \\ \hline
        Number of Departments & 9  \\ \hline
        Number of Degrees     & 11 \\ \hline
    \end{tabular}

    \caption
        [Courses in numbers]
        {Courses in numbers.}

    \label{tab:courses_stats}
\end{table}

If some courses only have one or two students who fail, they might not be
useful for training of the model because they won't add any useful information
to them. To understand the distribution of students in the courses, we plot the
number of students for each course, and the number of approved students.
Figure~\ref{fig:stat_007} shows that plot. Each bar is a course, and courses
are sorted by total number of students.

\begin{figure}[h!]
    \centering

    \includegraphics[width=\linewidth]{../src/stat_007_results/plot}

    \caption
        [Number of Students per Course and Number of Approved Students per
        Course]
        {Number of Students per Course and Number of Approved Students per
        Course. The grey bars show the number of students, while the blue bars
        show the number of approved students. The graph}

    \label{fig:stat_007}
\end{figure}

Courses have different values for credits. Table~\ref{tab:courses_credits}
shows how many courses are there for each number of credits. We see that the
majority of courses has 6 credits.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Credits & Number of Courses \\ \hline
        1       & 3                 \\ \hline
        2       & 3                 \\ \hline
        3       & 3                 \\ \hline
        5       & 6                 \\ \hline
        6       & 46                \\ \hline
        7.5     & 12                \\ \hline
        12      & 2                 \\ \hline
        15      & 2                 \\ \hline
    \end{tabular}

    \caption
        [Number of courses group by credits]
        {Number of courses group by credits.}

    \label{tab:courses_credits}
\end{table}

We take the number of courses for each degree. Table~\ref{tab:stat_003_res}.
The actual name of the degrees is not available, so we can only identify them
by their codes. Looking into the counts we see that the degree with code
\texttt{B\_M\_REST\_EINF\_E (578)} has the majority of courses by far. From
this table we also see that some degrees have a low number of degrees and some
even have have courses only on one semester.

It was to be expected that every degree would have a similar number of courses,
or at least that every course have curses for both semesters. It is unknown why
there are degrees present that have such characteristics, whether it is just
missing data or if there really are such degrees.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | r | r |}
        \hline
        Degree                             & Semester       & Number of Courses & Total \\ \hline
        \texttt{B\_M\_EHEA (478)}          & \textit{Par}   & 3                 & 7     \\
                                           & \textit{Ímpar} & 4                 &       \\ \hline
        \texttt{B\_M\_REST\_EINF\_E (578)} & \textit{Par}   & 14                & 28    \\
                                           & \textit{Ímpar} & 14                &       \\ \hline
        \texttt{B\_M\_REST\_TET (457)}     & \textit{Par}   & 2                 & 7     \\
                                           & \textit{Ímpar} & 5                 &       \\ \hline
        \texttt{B\_PD\_M\_E (580)}         & \textit{Par}   & 3                 & 8     \\
                                           & \textit{Ímpar} & 5                 &       \\ \hline
        \texttt{PG\_B\_EGNEG (554)}        & \textit{Par}   & 4                 & 9     \\
                                           & \textit{Ímpar} & 5                 &       \\ \hline
        \texttt{PG\_B\_AE (438)}           & \textit{Par}   & 5                 & 10    \\
                                           & \textit{Ímpar} & 5                 &       \\ \hline
        \texttt{FC\_EDM (564)}             & \textit{Par}   & 1                 & 1     \\ \hline
        \texttt{CCDND\_CPM (344)}          & \textit{Ímpar} & 1                 & 1     \\ \hline
        \texttt{FC\_CVP\_FDOC (513)}       & \textit{Ímpar} & 2                 & 2     \\ \hline
        \texttt{FC\_CVP\_UAM (514)}        & \textit{Ímpar} & 3                 & 3     \\ \hline
    \end{tabular}

    \caption
        [Number of Courses per Degree and Semester]
        {Number of Courses per Degree and Semester.}

    \label{tab:stat_003_res}
\end{table}

\section{Students}

Table~\ref{tab:moodle_users_stats} shows some stats regarding Moodle users.
Some of the explored projects listed in the background, like~\TODO{cite}, have
a number of students around 800 to 900 which is greater then the number of
students in this dataset, but they also have a lower number of courses, for
example~\TODO{cite} has 7, where this dataset has 77.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Number of Students   & 145 \\ \hline
        Number of Professors & 87  \\ \hline
        Total                & 232 \\ \hline
    \end{tabular}

    \caption
        [Professor features]
        {Professor features.}

    \label{tab:moodle_users_stats}
\end{table}

Observing the behaviour of students in most courses in any University, it is
common to find that the number of students decreases over time has the course
progresses. This is due to students giving up on that course for that
particular semester for various reasons. Although the data doesn't have any
explicit data point stating that a student gave up on a course, we can still
observe student activity to verify if it decreases or not as the semester
progresses over the weeks.

The activity on a course is measured using the Moodle logs which contain the
CRUD activities. CRUD activities, as stated in~\ref{sec:logs}, are aggregated
by courses, students, and weeks. To have a sense of how activity changes over a
course we take the logs dataset and aggregate every object only by course. So,
each entry will have the sum of the activities aggregated by course and week.

To get a sense of activities over the semester, we take six courses. Two of the courses are the ones with the greatest number of activities, other two are the ones with a median number of activities, and the last two are courses with the least number of activities.

We take the two courses with the most activity overall and plot a graph which x
axis is the weeks and y axis is the number of activities.
Figure~\ref{fig:stat_004_all} shows the graph for these courses. The number of
read activities and number of total activities is much greater then the number
of create, update, and delete activities. Because of this, we plot a graph
which only contains create, update, and delete activities for the same courses
as shown in figure~\ref{fig:stat_004_cud}.

\begin{figure}[h!]
    \centering

    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_all}
        \caption[]{}
        \label{subfig:stat_most_004_0_all}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_1_eDocOn-T1_all}
        \caption[]{}
        \label{subfig:stat_most_004_1_all}
    \end{subfigure}

    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_middle_0_MAT7514_1_all}
        \caption[]{}
        \label{subfig:stat_middle_004_0_all}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_middle_1_GES11656_all}
        \caption[]{}
        \label{subfig:stat_middle_004_1_all}
    \end{subfigure}

    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_0_HIS10755_all}
        \caption[]{}
        \label{subfig:stat_least_004_0_all}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_1_MAT7515_all}
        \caption[]{}
        \label{subfig:stat_least_004_1_all}
    \end{subfigure}

    \caption
        [Number of CRUD activities for courses]
        {Number of CRUD activities for two courses with greater activity
        (\ref{subfig:stat_most_004_0_all} and
        \ref{subfig:stat_most_004_1_all}), median activity
        (\ref{subfig:stat_middle_004_0_all} and
        \ref{subfig:stat_middle_004_1_all}), and least activity
        (\ref{subfig:stat_least_004_0_all} and
        \ref{subfig:stat_least_004_1_all}).}

    \label{fig:stat_004_all}
\end{figure}

\begin{figure}[h!]
    \centering

    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_cud}
        \caption[]{}
        \label{subfig:stat_most_004_0_cud}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_1_eDocOn-T1_cud}
        \caption[]{}
        \label{subfig:stat_most_004_1_cud}
    \end{subfigure}

    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_middle_0_MAT7514_1_cud}
        \caption[]{}
        \label{subfig:stat_middle_004_0_cud}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_middle_1_GES11656_cud}
        \caption[]{}
        \label{subfig:stat_middle_004_1_cud}
    \end{subfigure}

    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_0_HIS10755_cud}
        \caption[]{}
        \label{subfig:stat_least_004_0_cud}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_1_MAT7515_cud}
        \caption[]{}
        \label{subfig:stat_least_004_1_cud}
    \end{subfigure}

    \caption
        [Number of Create, Update, and Delete activities for courses]
        {Number of Create, Update, and Delete activities for two courses with
        greater activity (\ref{subfig:stat_most_004_0_cud} and
        \ref{subfig:stat_most_004_1_cud}), median activity
        (\ref{subfig:stat_middle_004_0_cud} and
        \ref{subfig:stat_middle_004_1_cud}), and least activity
        (\ref{subfig:stat_least_004_0_cud} and
        \ref{subfig:stat_least_004_1_cud}).}

    \label{fig:stat_004_cud}
\end{figure}

From figures~\ref{fig:stat_004_all} and~\ref{fig:stat_004_cud} we see that the
activities of courses remain generally the same during the semester. With the
exception of courses with the least activity, it is observable that while there
are a few peaks during the semester, there is no significant drop in activities
towards the end of them. The examples shown have peaks in various places during
the semester, however, they don't tend to be at the start or at the end of the
semester.

This leads to the conclusion not only that the levels of activity remain
relatively constant during the semester, but also leads to the conclusion that
many students don't quit the courses even i they fail.

To get an overall view of how activities change during the whole semester we
make another plot. This time, we see the total number of activities for every
course during the semester and the mean of activities.
Figure~\ref{fig:stat_005} shows this plot.

The number of total activities per week drops significantly. This is because
some courses are longer then others. For example, some courses may have their
last evaluation in the first weeks of the exam season, while others may still
have ongoing activity after exam season.

\TODO{Explain what are active courses and show the plot}

The mean of activities remains the same. The mean is calculated for each week
only on courses that have any activity on that week. So looking at the last few
weeks we get a mean of activities only for courses which are still active. With
this in mind, we again see that courses remain relatively the same in terms of
activities while the semester is going.

\begin{figure}[h!]
    \centering

    \includegraphics[width=\linewidth]{../src/stat_005_results/plot}

    \caption
        [Total Number of Activities vs Mean of Activities per Week]
        {Total Number of Activities vs Mean of Activities per Week.}

    \label{fig:stat_005}
\end{figure}

\section{Results}

Table~\ref{tab:results_stats} shows some statistics on the results data.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r |}
        \hline
        Number of results           & 716 \\ \hline
        Number of non-empty results & 57  \\ \hline
        Number of approved students & 334 \\ \hline
    \end{tabular}

    \caption
        [Statistics on the Results]
        {Statistics on the Results.}

    \label{tab:results_stats}
\end{table}

\section{Profile}

\section{Moodle Logs}

### 2017-03-22 src/exp_004.py

@@ -10,10 +10,27 @@ import tasks
 
 
 if __name__ == "__main__":
+    print("Original Classes")
     for i in range(2, 7):
         print("K-Means, %s" % (i))
-        tasks.unsupervised(data.dat_002(approvedJoin=False), "kmeans", i)
+        tasks.unsupervised(data.dat_002(approvedJoin=True), "kmeans", i)
 
+    print("2 Classes")
     for i in range(2, 7):
-        print("Affinity Propagation, %s" % (i))
-        tasks.unsupervised(data.dat_002(approvedJoin=False), "affinity", i)
+        print("K-Means, %s" % (i))
+        tasks.unsupervised(data.dat_002(classes="2", approvedJoin=True), "kmeans", i)
+
+    print("4 Classes")
+    for i in range(2, 7):
+        print("K-Means, %s" % (i))
+        tasks.unsupervised(data.dat_002(classes="4", approvedJoin=True), "kmeans", i)
+
+    print("6 Classes")
+    for i in range(2, 7):
+        print("K-Means, %s" % (i))
+        tasks.unsupervised(data.dat_002(classes="6", approvedJoin=True), "kmeans", i)
+
+    # Results are terrible
+    # for i in range(2, 7):
+    #     print("Affinity Propagation, %s" % (i))
+    #     tasks.unsupervised(data.dat_002(approvedJoin=False), "affinity", i)

### 2017-03-22 src/stat_004.py

@@ -17,7 +17,7 @@
 import os
 
 import numpy as np
-import pandas
+import pandas as pd
 from matplotlib import pyplot
 
 import data
@@ -27,8 +27,8 @@ import util
 output_dir = util.outputDirName(4)
 
 
-def outputFileName(index, courses, plot_type):
-    file_name = "fig_%d_%s_%s" % (index, courses, plot_type)
+def outputFileName(group_type, index, courses, plot_type):
+    file_name = "fig_%s_%d_%s_%s" % (group_type, index, courses, plot_type)
     return os.path.join(output_dir, file_name)
 
 
@@ -40,7 +40,7 @@ def takeCRUD(df, crud):
     return df.loc[lambda df: df["CRUD"] == crud]
 
 
-def plotCourse(index, course, groups):
+def plotCourse(group_type, index, course, groups):
     # Get this course's groups
     group_all = groups["all"].loc[course]
     group_C   = groups["C"].loc[course]
@@ -69,13 +69,13 @@ def plotCourse(index, course, groups):
     plot_U, = pyplot.plot(weeks_U, act_U)
     plot_D, = pyplot.plot(weeks_D, act_D)
 
-    pyplot.title("%d - %s" % (index, course))
+    # pyplot.title("%d - %s" % (index, course))
     pyplot.figlegend(
         (plot_all[0], plot_C, plot_R, plot_U, plot_D),
         ("All", "C", "R", "U", "D"),
         "upper right")
 
-    pyplot.savefig(outputFileName(index, course, "all"));
+    pyplot.savefig(outputFileName(group_type, index, course, "all"));
     pyplot.clf()
 
     # Plot the C, U, and D only
@@ -83,13 +83,13 @@ def plotCourse(index, course, groups):
     plot_U, = pyplot.plot(weeks_U, act_U)
     plot_D, = pyplot.plot(weeks_D, act_D)
 
-    pyplot.title("%d - %s" % (index, course))
+    # pyplot.title("%d - %s" % (index, course))
     pyplot.figlegend(
         (plot_C, plot_U, plot_D),
         ("C", "U", "D"),
         "upper right")
 
-    pyplot.savefig(outputFileName(index, course, "cud"));
+    pyplot.savefig(outputFileName(group_type, index, course, "cud"));
     pyplot.clf()
 
 
@@ -105,9 +105,16 @@ if __name__ == "__main__":
         "NumberOfActivitiesPerWeek"
     ]]
 
[…]

### 2017-03-23 ResearchNotes/stat_exps.tex

@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_0_CTE-UMCED-2_all}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_all}
         \label{subfig:stat_004_0_all}
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_3_INF7194_2_all}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_1_eDocOn-T1_all}
         \label{subfig:stat_004_3_all}
@@
     \caption
-        [Stat 004 plots for all activities]
-        {Stat 004 plots for all activities.}
+        [Number of CRUD activities for two courses with greater activity]
+        {Number of CRUD activities for two courses with greater activity.}
 
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_0_CTE-UMCED-2_cud}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_cud}
         \label{subfig:stat_004_0_cud}
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_3_INF7194_2_cud}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_most_1_eDocOn-T1_cud}
         \label{subfig:stat_004_3_cud}
@@
     \caption
-        [Stat 004 plots for C, U, and D activities]
-        {Stat 004 plots for C, U, and D activities.}
+        [Number of Create, Update, and Delete activities for two courses with
+        greater activity]
+        {Number of Create, Update, and Delete activities for two courses with
+        greater activity.}
 

### 2017-03-23 ResearchNotes/tools.tex

\section{Tools}

Python 3.5.1 was used for everything. The tools used were:

\begin{itemize}
    \item Numpy 1.11.3~\cite{web_numpy}
    \item SciPy 0.18.1~\cite{web_scipy}
    \item SciKit-Learn 0.18.1~\cite{web_scikit}
    \item Pandas 0.19.2~\cite{web_pandas}
    \item Matplotlib 1.5.4~\cite{web_matplotlib}
\end{itemize}

### 2017-03-23 ThesisText/background.tex

\chapter{Background}

% =============================================================================
\section{E-Learning and Learning Management Systems}
\label{sec:elearning_and_lms}

Recently, E-Learning has been rising in popularity as a solution for the
delivery of training. Institutions such as universities and companies have
successfully implemented E-Learning to bring course content and evaluation to
students, and to deliver training to employees remotely. This has been made
possible by the advances in technology and the increasing use of the
Internet.

E-Learning is usually provided over the Internet using platforms
called~\gls{lms}. Examples of~\gls{lms} include Moodle, Blackboard Learn, D2L,
among others. These platforms organize users by separating them into roles such
as Professors and Students or Authors and Learners. The Professors or Authors
are the users who organize courses in universities or training in companies,
create and publish content, perform evaluation of students, etc. Students or
Learners are the users which actually perform the training and follow courses
in the platforms.~\cite{ind_010, ind_011, ind_013}

In the specific case of Moodle, content is divided into courses. Each Moodle
course represents a course from a University or similar learning institution.

The contents of a course may be resources such as PDF files, links, media such
as images, videos, etc. These resources should be guides for students achieve
their objectives in learning, or should be resources needed for the student to
perform certain tasks. A course may also have activities. Activities may be,
for example, projects that a student needs to complete before a given deadline.
Activities may also be quizzes made to students over Moodle. These quizzes may
be used for evaluation.

Course also have forums, in which professors and students can exchange messages
which are relevant to the topics at hand.

A course is overseen by a group of Professors who will publish resources and
create activities. Students will use the resources and perform the
activities.~\cite{ind_014, ind_015}

% =============================================================================
\section{Data in Learning Management Systems}
\TODO{This section should just be part of the previous one.}

A typical~\gls{lms} contains various data entries which are stored in a
database. Information on users, such as their names, usage logs, roles, forum
messages, are generated by the systems and kept in their databases.

Resources and activities, as well as the logs detailing their usage, are also
kept in a database. The usage of these systems by professors and students
generates entries.

% =============================================================================
\section{Data Mining}

Current computational systems produce and store large amounts of data. This
data may represent anything, such as the sales records in a large commercial
chain, the usage telemetry on a popular web service, or even the registers of
some scientific survey. Data is usually rich in features. For example, the data
from the sales records would contain information on which items were sold and
in what quantity, to whom they were sold, on what date, and so on.

Data Mining aims at finding patterns in data by looking at the different
features of the data entries. These patterns allows then to make non trivial
assumptions in data.~\cite{book_dm_practical}

To find these patterns we build and train model such as Decision Trees,
Bayesian Networks, etc, from the data.

Data is usually stored in databases which may follow different paradigms and
have different architectures. But generally, we refer to data entries as
datasets. We also refer to a group of datasets as data repository.

\TODO{Talk about specific models}

% =============================================================================
\section{Educational Data Mining}

\TODO{Talk about clustering (See ind\_007 for things about this)}

Educational Data Mining, \gls{edm}, refers to the act of applying Data Mining
techniques to educational data. Educational data refers to data which origin is
in the act of teaching and training, in the activities of professors and
students, results of evaluation, and so on. In this project, the origin of data
is the University of Évora's Moodle.

The main objective of~\gls{edm} is to extract meaningful information from
educational data like web logs. It is not trivial to do so without advance Data
Mining techniques. \gls{edm}, therefore, aims at training models based on
education data for a variety of tasks.

One such example of~\gls{edm} Data Mining tasks is too predict student grades
based features from the students themselves, the courses they are undertaking,
and their activity in a~\gls{lms}. By training such models, professors can make
decisions about how courses are organized regarding students success. For
example, a certain number of activities in a given course may be beneficial to
it's success because it keeps students engaged, but any number of activities
more might not make a difference.~\cite{ind_001, ind_002, ind_005}

These models also detect students which may have greater difficulty completing
certain courses, or be used to predict how many students will complete a
certain class.~\cite{ind_007, ind_008}.

In~\cite{ind_001}, some models are trained in order to predict student grades.
The tasks described in the paper are classification tasks, in which a student
is assigned to a given class corresponding to their success in courses based on
their usage of a~\gls{lms} platform. The classes available are Excellent, Good,
Pass, and Fail.

The dataset used in~\cite{ind_001} contained the fields in
table~\ref{tab:ind_001_004_fields} and was made from registers of 438 students
in 7 different courses.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Field name                       & Description                                    \\ \hline
        \texttt{course}                  & Identification number of the course.           \\
        \texttt{n\_assigment}            & Number of assignments done.                    \\
        \texttt{n\_quiz}                 & Number of quizzes taken.                       \\
        \texttt{n\_quiz\_a}              & Number of quizzes passed.                      \\
        \texttt{n\_quiz\_s}              & Number of quizzes failed.                      \\
        \texttt{n\_posts}                & Number of messages sent to the forum.          \\
        \texttt{n\_read}                 & Number or messages read on the forum.          \\
        \texttt{total\_time\_assignment} & Total time used on assignments.                \\
        \texttt{total\_time\_quiz}       & Total time used on quizzes.                    \\
        \texttt{total\_time\_forum}      & Total time used on forum.                      \\
        \texttt{mark}                    & Final mark the student obtained in the course. \\
    \end{tabular}

    \caption
        [Features in experiment~\cite{ind_001} and~\cite{ind_004}]
        {Features in experiment~\cite{ind_001} and~\cite{ind_004}.}

    \label{tab:ind_001_004_fields}
\end{table}

The used models are Statistical Classifiers, Decision Trees, Association Rules,
and Neural Networks. After training, performance measures are calculated using
test data. The best models were found to be decision trees according to the
calculated global percentage of correctly classified students.

The same experiment described in~\cite{ind_001} is also described
in~\cite{ind_004} with the same models, data, and results.

Similar models to before are used in~\cite{ind_005}. It is observed again that
Decision Trees have the best performance. Table~\ref{tab:ind_005_fields} shows
the data from~\cite{ind_005}. The experiments were made on data from 824
students in 11 courses.

\begin{table}[h!]
    \centering

    \begin{tabular}{l l}
        Field name                & Description                                              \\ \hline
        \texttt{UserName}         & Name of User                                             \\
        \texttt{CourseName}       & Name of the Course                                       \\
        \texttt{ResourceView}     & Number of Coursware and Other Supporting Materials Views \\
        \texttt{VirtualClassroom} & Number of Virtual Classroom Participations               \\
        \texttt{ArchiveView}      & Number of Archive Views                                  \\
        \texttt{ForumRead}        & Number of Forum Reads                                    \\
        \texttt{ForumPost}        & Number of Forum Posts                                    \\
        \texttt{DiscussionRead}   & Number of Discussion Reads                               \\
        \texttt{DiscussionPost}   & Number of Discussion Responses                           \\
        \texttt{AssignmentView}   & Number of Assignments Views                              \\
        \texttt{AssignmentUpload} & Number of Assignment Answer Uploads                      \\
        \texttt{FinalGrade}       & Final Grades                                             \\
    \end{tabular}

    \caption
        [Features in experiment~\cite{ind_005}]
        {Features in experiment~\cite{ind_005}.}

    \label{tab:ind_005_fields}
\end{table}

Neural networks are used in~\cite{ind_003} and~\cite{ind_006} for similar
purposes. In~\cite{ind_003} a neural network is used on a dataset with 116
entries. Each entry is made from 25 features and represents the results of
placements tests made to students. There are 5 outputs to the network, one for
each possible grade. The paper claims correctly classified rates of over 90\%
in some trained networks while most are just above 80\%. This shows how neural
networks can be used in this type of problems.

In~\cite{ind_006}, networks are trained from Moodle logs. The data from the
logs are structured in a slightly different way then the previous studies. The
features of this data are shown in table~\ref{tab:ind_006_fields}.

\begin{table}[h!]
    \centering

    \begin{tabular}{l}
        Description \\ \hline
        Full name of the student. \\
        Number of times that has been officially registered in the subject. \\
        Number of examination sessions. \\
        Mark in numerical format. \\
        Total of accesses (of any type) made to the Moodle system during course 2005/2006. \\
        Total of accesses to “resource view” \\
        Percentage of “resource view” accesses from the total accesses. \\
        Number of different “resource view”of each type (theoretical, examples, etc.) have been visited. \\
        Percentage of the resources of each type (theoretical, examples, etc.) sights. \\
        Segmentation of the number of accesses in every month of the year. \\
        Segmentation per month of the percentage of accesses. \\
    \end{tabular}

    \caption
        [Features in experiment~\cite{ind_006}]
        {Features in experiment~\cite{ind_006}.}

    \label{tab:ind_006_fields}
\end{table}

The data is extracted from 240 students. The number of courses is not
specified. Some trained networks are able to achieve 80\% of correctly
classified instances, while most achieve a value over 70\%.

% =============================================================================
\section{Data Mining Tools}

The practical component of this project was developed entirely using Python and
some Python libraries which are relevant in. The version of CPython used was
version 3.5.2. The python libraries Pandas, version 0.19.2, was used for
statistical analysis, exploration, and preprocessing. The library SciKit-Learn,
version 0.18.1, was used to perform Data Mining tasks. SciKit-Learn uses Numpy
and Scikit, in versions 1.11.3 and 0.18.1 respectively. To output some graphs,
the python library Mat Plot Lib, version 1.5.4, was used.

\TODO{State the actual methods and algorithms in SciKit-Learn that were used}

### 2017-03-23 ThesisText/experiments.tex

\chapter{Experiments}
\label{sec:exps}

The Data Mining experiments made in this project consist in organizing the data
from the processed datasets described in section~\ref{sec:preproc_dataset} into
a new dataset useful for a set of experiments. This chapter is organized by
sections on each dataset. Each dataset is used in several experiments with
different algorithms.

\section{Students, Courses, Logs, and Results}
\label{sec:exp_001_002}

In an initial analysis, we look into how Moodle usage relates with student's
success. A dataset is build with the features presented in
table~\ref{tab:dat_001}. The new dataset takes information from the following
datasets:

\begin{itemize}
    \item \texttt{Students}
    \item \texttt{CoursesGeneral}
    \item \texttt{MoodleLogs}
    \item \texttt{Results}
\end{itemize}

Each object of the dataset contains information referencing a single student in
a single course. The objects contain the overall enrollment count and type of
student. They also contain the identification of a course along with the
semester and season of that course. Finally, the object contains a count of
CRUD activities of that student in that course.

The dataset can be generated with different classes. There are three options.
Option one, the class is binary and has the value true if the student was
approved on that course, and false otherwise. Option two, the class has one of
the following values which are intervals of grades:

\[
    [0, 10[\;,\, [10, 12[\;,\, [12, 14[\;,\, [14, 16[\;,\, [16, 18[\;,\, [18, 20]
\]

Option three, the class has one of the following values:

\[
    [0, 10[\;,\, [10, 14[\;,\, [14, 18[\;,\, [18, 20]
\]

With option two, we get a dataset which groups every records with grades below
10, but separates the positive grades in groups of two. Option three does the
same but separates grades into groups of four, with the last group having only
three values instead of four.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | l |}
        \hline
        \textbf{Field}             & \textbf{Source Dataset}    & \textbf{Type} \\ \hline
        \texttt{EnrollementCount}  & Student                    & Number        \\ \hline
        \texttt{StudentType}       & Student                    & Enumerated    \\ \hline
        \texttt{CourseCodeSiiue}   & CoursesGeneral             & String        \\ \hline
        \texttt{Semester}          & CoursesGeneral             & Enumerated    \\ \hline
        \texttt{Season}            & CoursesGeneral             & Enumerated    \\ \hline
        \texttt{TotalActsOnCourse} & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{CActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{RActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{UActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{DActsOnCourse}     & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{Class}             & Calculated From Results    & Binary        \\ \hline
    \end{tabular}

    \caption
        [Students, Courses, Logs, and Results]
        {Students, Courses, Logs, and Results. This dataset joins information
        from four processed datasets in order to relate a student, a course,
        usage data from that student on that course, and the final grade of it}

    \label{tab:dat_001}
\end{table}

\subsection{Predicting Students Grades Based on Moodle Usage}
% EXP_001

A first experiment on this dataset aims at predicting the students grades based
on their Moodle usage data and in their other characteristics present on this
dataset. In this experiment six tasks are made. Both Decision Trees and Bayesian
Networks are trained using the dataset with binary classes, classes with
intervals of two, and classes with intervals of four.

Some fields on dataset~\ref{tab:dat_001} are strings and enumerated types,
therefore, these are nominal fields. Because of this characteristic, Decision
Trees and Bayesian Networks are used.~\TODO{Citation is needed on claim that
Trees and BNets are good for nominal data}

The models are trained using cross-validation with 10 splits. This means that
the data is divided into 10 parts. Nine parts are used to train to models and
the remaining part is used to test the models.

Training is executed 10 times, each time with different splits for training and
testing. The accuracy of each model is taken. Table~\ref{tab:exp_001_res} shows
the maximum accuracy of the trained models, the minimum, and the average.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree    & Binary Class          & 0.92 & 0.34 & 0.70 \\
                         & Grades Classes 4 by 4 & 0.84 & 0.22 & 0.59 \\
                         & Grades Classes 2 by 2 & 0.74 & 0.34 & 0.55 \\ \hline
        Bayesian Network & Binary Class          & 0.95 & 0.53 & 0.74 \\
                         & Grades Classes 4 by 4 & 0.73 & 0.36 & 0.57 \\
                         & Grades Classes 2 by 2 & 0.67 & 0.28 & 0.52 \\ \hline
    \end{tabular}

    \caption
        [Predicting students grades based on Moodle usage results]
        {Predicting students grades based on Moodle usage results. Two models
        are trained 10 times on three different variations of the same dataset.
        The maximum, minimum, and average accuracy is displayed for those ten
        executions.}

    \label{tab:exp_001_res}
\end{table}

It is observable that for binary classes, Bayesian Networks performed slightly
better then decision trees. But for grade classes two by two and four by four,
decision trees were better. In particular, training models for binary classes,
understandably, yields better results on average.

These results show that such models can be trained to predict grades of
students with success.

\subsection{Clustering Students Grades Based on Moodle Usage}
% EXP_002

With the same dataset, clustering using the K-Means algorithm is performed,
however, the class feature is removed. The training and testing method is
similar to the previous experiment. Cross validation is used and a model is
trained 10 times, each time using nine different folds for training and one
fold for testing.

K-Means is trained and tested for a number of clusters between two and five.
Testing is done using silhouette score. Like before, the maximum, minimum, and
average score is taken. Table~\ref{tab:exp_002_res} shows the results.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r | r | r |}
        \hline
        \textbf{Number of Clusters} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        2 & 0.72 & 0.68 & 0.69 \\ \hline
        3 & 0.67 & 0.65 & 0.66 \\ \hline
        4 & 0.68 & 0.64 & 0.66 \\ \hline
        5 & 0.67 & 0.65 & 0.66 \\ \hline
    \end{tabular}

    \caption
        []
        {}

    \label{tab:exp_002_res}
\end{table}

In general, the model fails to get good results having a maximum score of 0.72
for two clusters. Numbers greater them five clusters were tried but they,
understandably, got lower scores.

To understand what the algorithm is separating, that is, which type of objects
are being placed in which cluster, we look in detail to an execution of K-Means
with two clusters.

Table~\ref{tab:exp_002a_res} has a listing of objects that are placed in each
of the two clusters. It is observable that the major difference is in the
number of CRUD activities. Different values for activities remain the same for
different courses and enrollment numbers, but greater values are placed in one
cluster while lower values are placed in another.

Besides K-Means, Affinity Propagation was executed with this dataset, but the
algorithm performed poorly have scores below 0.5.

\section{Number of Courses and Logs}

In similar experiments to~\ref{sec:exp_001_002}, we make a dataset in which we
relate a students total number of enrolled courses,  number of approved
courses, and Moodle usage data.

Each object of this dataset will have information on a single student. As
mentioned, an object has the number of enrolled and approved courses and the
sum of each CRUD activity for every course of that single student. Each entry
does not have the identification of the student to which it refers too because
that information is irrelevant for our learning objectives.

This dataset only contains features which have a number type. The records come
from the \texttt{Results} and \texttt{MoodleLogs} datasets.
Table~\ref{tab_dat_002} shows the features of this dataset.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | l |}
        \hline
        \textbf{Field}           & \textbf{Source Dataset}    & \textbf{Type} \\ \hline
        \texttt{EnrolledCourses} & Calculated From Results    & Number        \\ \hline
        \texttt{ApprovedCourses} & Calculated From Results    & Number        \\ \hline
        \texttt{ActSumAll}       & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{ActSumC}         & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{ActSumR}         & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{ActSumU}         & Calculated From MoodleLogs & Number        \\ \hline
        \texttt{ActSumD}         & Calculated From MoodleLogs & Number        \\ \hline
    \end{tabular}

    \caption
        [Number of Courses and Logs]
        {Number of courses and logs. Each object in this dataset contains the
        number of enrolled courses, number of approved courses and Moodle usage
        data for a single student.}

    \label{tab:dat_002}
\end{table}

Some classification tasks use the \texttt{ApprovedCourses} feature as a class. In this case, feature may be left in it's original form or it may be replaced by classes. There are three ways to do this replacement.~\TODO{Explain this and mentioned statistics}

\subsection{Predicting Number of Approved Courses Based on Moodle Usage}
% EXP_003

Like before both Decision Trees and Bayesian Networks are used.
Table~\ref{tab:exp_003_res} shows the maximum, minimum, and average accuracy
for each pair of model and classes.

The models are trained using cross validation with 10 splits like before.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree    & 2 Classes        & 1.00 & 0.64 & 0.83 \\
                         & 4 Classes        & 1.00 & 0.29 & 0.68 \\
                         & 6 Classes        & 1.00 & 0.21 & 0.68 \\
                         & Original Classes & 0.86 & 0.14 & 0.52 \\ \hline
        Bayesian Network & 2 Classes        & 1.00 & 0.47 & 0.70 \\
                         & 4 Classes        & 0.87 & 0.29 & 0.56 \\
                         & 6 Classes        & 0.86 & 0.07 & 0.43 \\
                         & Original Classes & 0.86 & 0.00 & 0.39 \\ \hline
    \end{tabular}

    \caption
        [Predicting number of approved courses based on Moodle usage results]
        {Predicting number of approved courses based on Moodle usage results.
        These are the results of two models being trained 10 times for 4
        different variants of the same dataset. The maximum, minimum, and
        average accuracy is shown.}

    \label{tab:exp_003_res}
\end{table}

It is observable that Decision Trees reached an accuracy of 1 producing, no
false negatives or false positives in testing, for all datasets except for the
one with the original classes. On average, Decision Trees, models achieve high
levels of accuracy for every dataset.

Bayesian Networks are achieve hight level of accuracy, but not as high as
Decision Trees. For the dataset with the original classes, Bayesian Networks
perform similarly in the best case scenario, given that the values for maximum
accuracy are similar. But the values for minimum and average are noticeably
lower.

Because the models trained with the original classes achieve high values for
accuracy we conclude that it is possible to predict the exact number of
approved courses with great accuracy given only the initial number of enrolled
courses and the Moodle usage data, especially if such tasks are executed with
Decision Trees.

\subsection{Clustering Number of Approved Courses Based on Moodle Usage}
% EXP_004

Clustering methods are applied to look for groups of similar objects within
this data. In this experiment, every field is a feature, so the
\texttt{ApprovedCourses} field is a feature and not a class like before.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r | r | r |}
        \hline
        \textbf{Number of Clusters} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        2 & 0.83 & 0.80 & 0.81 \\ \hline
        3 & 0.78 & 0.75 & 0.76 \\ \hline
        4 & 0.78 & 0.57 & 0.73 \\ \hline
        5 & 0.68 & 0.55 & 0.61 \\ \hline
        6 & 0.57 & 0.55 & 0.56 \\ \hline
    \end{tabular}

    \caption
        [Clustering number of approved courses based on Moodle usage results]
        {Clustering number of approved courses based on Moodle usage results. The table displays the results for the silhouette score for clustering.}

    \label{tab:exp_004_res}
\end{table}

\TODO{These values are the same for every variant of the dataset. This needs to be explained}

\section{Student/Course Profile and Results}

In this dataset we relate the profiling of a student in a courses and his
results on that course. To begin constructing this dataset we use the Profiling
dataset, as shown in section~\ref{sec:profile}.

The Profiling dataset contains a results field, however, the meaning of each
value in this field is unknown. So initially, this field is dropped. To get the
actual results of students we need the Results dataset.

Each object of Profile contains the features \texttt{CourseCodeMoodle},
\texttt{CourseId}, and \texttt{StudentsId}. But the Results dataset has
\texttt{CourseCodeSiiue} and \texttt{StudentsNumber}. In order to relate both,
the Students dataset and the CoursesGeneral dataset are needed.

These datasets are related and some fields are dropped in order to produce a
dataset with the features listed in table~\ref{tab:dat_003}.

Like in dataset~\ref{sec:exp_001_002}, the class of this dataset may be binary, in
which the student is approved or not, or divided in groups of 2 or 4.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | l |}
        \hline
        \textbf{Field}               & \textbf{Source Dataset} & \textbf{Type} \\ \hline
        \texttt{CourseCodeSiiue}     & CoursesGeneral          & String        \\ \hline
        \texttt{NumberOfResources}   & Profile                 & Number        \\ \hline
        \texttt{NumberOfActivities}  & Profile                 & Number        \\ \hline
        \texttt{NumberOfViews}       & Profile                 & Number        \\ \hline
        \texttt{NumberOfSubmissions} & Profile                 & Number        \\ \hline
        \texttt{Class}               & Results                 & Binary        \\ \hline
    \end{tabular}

    \caption
        [Student/Course Profile and Results]
        {Student/Course Profile and Results.}

    \label{tab:dat_003}
\end{table}

\subsection{Predicting Students Grades From Student/Course Profiling}
% EXP_005

Using the described dataset, we train models to predict student grades based on
profiling data. Table~\ref{tab:exp_005_res} shows the results of six tasks.
Three for each Model. Each tasks is done in a variant of the dataset.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree    & Binary Class          & 0.91 & 0.43 & 0.73 \\
                         & Grades Classes 4 by 4 & 0.65 & 0.31 & 0.44 \\
                         & Grades Classes 2 by 2 & 0.56 & 0.18 & 0.34 \\ \hline
        Bayesian Network & Binary Class          & 0.97 & 0.49 & 0.77 \\
                         & Grades Classes 4 by 4 & 0.91 & 0.31 & 0.55 \\
                         & Grades Classes 2 by 2 & 0.71 & 0.26 & 0.46 \\ \hline
    \end{tabular}

    \caption
        [Predicting students grades from student/course profiling results]
        {Predicting students grades from student/course profiling results. The
        results show the maximum, minimum, and average for each trained model
        and each dataset variant.}

    \label{tab:exp_005_res}
\end{table}

We observe from the results that Bayesian Networks perform better then Decision
Trees achieving greater levels of accuracy for each dataset variant.

\subsection{Clustering Students Grades From Student/Course Profiling}
% EXP_006

Taking the same dataset, we remove the field \texttt{CourseCodeSiiue} and apply
clustering like in previous experiments. Table~\ref{tab:exp_006_res} shows the
results of this experiment.

\begin{table}[h!]
    \centering

    \begin{tabular}{| l | r | r | r |}
        \hline
        \textbf{Number of Clusters} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        2 & 0.64 & 0.60 & 0.63 \\ \hline
        3 & 0.58 & 0.56 & 0.57 \\ \hline
        4 & 0.57 & 0.53 & 0.54 \\ \hline
        5 & 0.51 & 0.50 & 0.50 \\ \hline
    \end{tabular}

    \caption
        []
        {}

    \label{tab:exp_006_res}
\end{table}

Having two clusters we get the higher values for silhouette score, but those
values are still far from 1, meaning that the executed algorithms do not find
any meaningful separation of objects into groups.

### 2017-03-23 ThesisText/statistics.tex
Minor corrections on statistics

@@
 Table~\ref{tab:moodle_users_stats} shows some stats regarding Moodle users.
-Some of the explored projects listed in the background, like~\TODO{cite}, have
-a number of students around 800 to 900 which is greater then the number of
+Some of the explored projects listed in the background, like~\cite{ind_001},
+have a number of students around 800 to 900 which is greater then the number of
 students in this dataset, but they also have a lower number of courses, for
-example~\TODO{cite} has 7, where this dataset has 77.
+example~\cite{ind_001} has 7, where this dataset has 77.
 
@@
 
-To get a sense of activities over the semester, we take six courses.
-Two of the courses are the ones with the greatest number of activities, other two are the ones with a median number of activities, and the last two are courses with the least number of activities.
+To get a sense of activities over the semester, we take six courses. Two of the
+courses are the ones with the greatest number of activities, other two are the
+ones with a median number of activities, and the last two are courses with the
+least number of activities.
 
@@
 relatively constant during the semester, but also leads to the conclusion that
-many students don't quit the courses even i they fail.
+many students don't quit the courses even if they fail.
 
@@
 
-\TODO{Explain what are active courses and show the plot}
-
 The mean of activities remains the same. The mean is calculated for each week
@@
 activities while the semester is going.
+
+The concept of active course needs to be understood. A course is active from
+week 0 until the last week for which there is any activity. For example, if the
+last activity of a course is on week 20, then the course is active from week 0
+until week 20.
+
+Figure~\ref{fig:stat_006} shows a graphs bar with the number of active courses
+per week, and again shows the total number of activities over these bars. We
+see that the number of active courses fall correlates with the fall of the
+total number of activities. This further reinforces the notion that the
+activity of courses remains relatively the same as the semester moves forward.
 
@@
     \caption
-        [Total Number of Activities vs Mean of Activities per Week]
-        {Total Number of Activities vs Mean of Activities per Week.}
+        [Total number of activities vs mean of activities per week]
+        {Total number of activities (Right) vs mean of activities per week
+        (Left).}
 
@@
 
+\begin{figure}[h!]
+    \centering
+
+    \includegraphics[width=\linewidth]{../src/stat_006_results/plot}
+
+    \caption
+        [Total number of activities vs mean of activities per week]
+        {Total number of activities (right) vs total active courses (Left).}
+
+    \label{fig:stat_006}
+\end{figure}
+
 \section{Results}
 
-Table~\ref{tab:results_stats} shows some statistics on the results data.
+Table~\ref{tab:results_stats} shows some statistics on the results data. We see
+that from the available 716 entries, 334 of the, refer to instances of a
+student getting approved on a course.
+
+As seen in section~\ref{sec:results}, the dataset containing the results has
+the fields \textit{Grade}, \textit{Results}, and \textit{FinalResult} for a
+combination of course and student. However, some objects have no values for
+those three fields. When that happens, we interpret the objects as having the
+student not getting approved on that course. From the 716 results, 57 of them
+are empty.
 
@@
         Number of results           & 716 \\ \hline
-        Number of non-empty results & 57  \\ \hline
+        Number of non-empty results & 659  \\ \hline
         Number of approved students & 334 \\ \hline
@@
 \end{table}
-
-\section{Profile}
-
-\section{Moodle Logs}

### 2017-03-23 ThesisText/background.tex
Adds information on tools

@@
 The practical component of this project was developed entirely using Python and
-some Python libraries which are relevant in. The version of CPython used was
-version 3.5.2. The python libraries Pandas, version 0.19.2, was used for
-statistical analysis, exploration, and preprocessing. The library SciKit-Learn,
-version 0.18.1, was used to perform Data Mining tasks. SciKit-Learn uses Numpy
-and Scikit, in versions 1.11.3 and 0.18.1 respectively. To output some graphs,
-the python library Mat Plot Lib, version 1.5.4, was used.
+some Python libraries. The version of CPython used was version 3.5.2. The
+python libraries Pandas, version 0.19.2, was used for statistical analysis,
+exploration, and preprocessing. The library SciKit-Learn, version 0.18.1, was
+used to perform Data Mining tasks. SciKit-Learn uses Numpy and Scikit, in
+versions 1.11.3 and 0.18.1 respectively. To output some graphs, the python
+library Mat Plot Lib, version 1.5.4, was used.
 
-\TODO{State the actual methods and algorithms in SciKit-Learn that were used}
+Decision trees and Bayesian Networks were used in this project. The algorithm
+to train decision trees, implemented in the SciKit-Learn library, is an
+optimized version of the CART algorithm~\cite{web_scikit_trees}. Bayesian
+Networks are used in Naive Bayes tasks. The algorithm to train such models,
+also implemented in SciKit-Learn, is Multinomial Naive
+Bayes~\cite{web_scikit_bayes}.
+
+Clustering is done using K-Means~\cite{web_scikit_kmeans} and Affinity
+Propagation~\cite{web_scikit_affinity}. Both algorithms are implemented in
+SciKit-Learn.

### 2017-03-27 ThesisText/background.tex
Adds parts for background

@@
 students, and to deliver training to employees remotely. This has been made
-possible by the advances in technology and the increasing use of the
-Internet.
+possible by the advances in technology and the increasing use of the Internet.
 
@@
 Learners are the users which actually perform the training and follow courses
-in the platforms.~\cite{ind_010, ind_011, ind_013}
+in the platforms.
+
+A typical~\gls{lms} contains various data entries which are stored in a
+database. Information on users, such as their names, usage logs, roles, forum
+messages, are generated by the systems and kept in their databases.
+
+Resources and activities, as well as the logs detailing their usage, are also
+kept in a database. The usage of these systems by professors and students
+generates entries.~\cite{ind_010, ind_011, ind_013}
 
@@
 % =============================================================================
-\section{Data in Learning Management Systems}
-\TODO{This section should just be part of the previous one.}
-
-A typical~\gls{lms} contains various data entries which are stored in a
-database. Information on users, such as their names, usage logs, roles, forum
-messages, are generated by the systems and kept in their databases.
-
-Resources and activities, as well as the logs detailing their usage, are also
-kept in a database. The usage of these systems by professors and students
-generates entries.
-
-% =============================================================================
 \section{Data Mining}
@@
 Data Mining aims at finding patterns in data by looking at the different
-features of the data entries. These patterns allows then to make non trivial
+features of the data entries. These patterns allows us to make non trivial
 assumptions in data.~\cite{book_dm_practical}
-
-To find these patterns we build and train model such as Decision Trees,
-Bayesian Networks, etc, from the data.
 
@@
 have different architectures. But generally, we refer to data entries as
-datasets. We also refer to a group of datasets as data repository.
+datasets. We also refer to a group of datasets as a data repository.
 
-\TODO{Talk about specific models}
+To find patterns in data, models are trained either using supervised or
+unsupervised learning. The following sections describe these concepts and
+introduce models from both.
+
+\subsection{Supervised Learning}
+
+Supervised learning is done when data made up of a collection of labeled
+objects which have a list of features and a label, usually called class. The
+idea is that a model is trained using instances from labeled data in order to
+correctly classify unlabeled instances later.
+
+There exist several supervised learning models, but in this project we use only
+two. They are Decision Trees and Naive Bayes
+Classifiers.~\cite{book_ml,book_dm_practical}
+
+\subsubsection{Decision Trees}
+
+Decision trees are a model that make use of a tree to classify an object. The
+model works by sorting an object down the tree, taking a single feature into
+account on each node of the tree. The branches on each node are the possible
+values of the feature being taken into account on that node. The leafs of the
+tree assign a class to the object.
+
+Decision trees are generally best suited for situations where the data has
+discrete values. This is the case in this project, since many features are
+either integer numbers with a limited range of values, or strings and
+enumerated types which are nominal in nature.
+
+There are a few algorithms used to train these model. They are ID3, C4.5 which
+is an optimization of ID3, and CART.~\cite{book_ml,book_dm_practical}
+
+\subsubsection{Naive Bayes Classifiers}
+
+Naive Bayes Classifiers are probabilistic models which are based on Bayesian
+Networks. A Bayesian Network is a graph that describes directional the
+dependencies between features of objects. Those features are described as
+random variables.
+
+The models that make up a Naive Bayes Classifier are a particular case of
+Bayesian Networks. In these models there is a single node of the network which
+is the class of the object. Every feature depends directly on the class and
+only on the class.
+
+Each model essentially describes a conditional probability distribution like:
+
+\begin{equation} \label{eq:prob} P(\; x_1, \; \dots, \; x_n \; | \; C),
+\end{equation}
+
+where $ C $ is the class of the object and $ x_1, \; \dots, \; x_n $ are the
+features of it. The model is able to classify an object by directly applying
+the Bayes Theorem to the distribution, obtaining a value a $ C $.
+
+There are different types of Naive Bayes Classifier. Each type is adequate for
+different types of features. For example, Guassian Naive Bayes Classifier
+describe relations between variables, like equation~\ref{eq:prob}, as normal
+distributions. Bernoulli Naive Bayes Classifier are used if the features are
+boolean.
+
+Multinomial Naive Bayes Classifier are used when the data has discrete values
+and is nominal in nature. This model is used in this project because the data
+it deals with is discrete and nominal.~\cite{book_ml,book_dm_practical}
+
+\subsection{Unsupervised Learning}
+
+Unsupervised learning deals with data that is unlabeled. One approach to
+unsupervised learning is clustering. The objective of clustering is to find
+groups of unlabeled objects. A group would ideally have objects which features
+are similar to objects within that group, and dissimilar to objects in other
+groups.
+
+One difficulty of clustering is to find the number of clusters that optimally
+divide the data. Some algorithms take a number of clusters as an arguments,
+while others are able to find an optimal value.
+
+By it's own nature, it may not be possible to find clusters in data, that is,
+there may not be any great division of objects given their features in a given
+dataset. After executing clustering algorithms a performance measure is used to
+determine if that division is good or not.
+
+In this project, the performance measure used is called Silhouette Score.
+Silhouette score yields a value between -1 a 1. The interpretation given to
+this value is that higher values mean that objects are correctly placed in
+their clusters, lower values mean the objects are in the wrong clusters. Values
+near zero indicate that the clusters overlap.~\cite{book_ml,book_dm_practical}
+
+\subsubsection{K-Means}
+
+K-Means is an clustering algorithm that find $ K $ clusters in data. The $ K $
+variable is given to the algorithm as an argument, hence the name K-Means. The
+algorithm works by creating $ K $ random points based on the features of the
+data. The algorithm then iterates updating the location of those points until
+they converge.~\cite{book_ml,book_dm_practical}
+
+\subsubsection{Affinity Propagation}
+
+Affinity Propagation is another clustering algorithm, but this one does not
+take a number of clusters as an argument. This algorithm finds the optimal
+number of clusters.~\cite{book_ml,book_dm_practical}
+
+\section{Cross Validation}
+
+The model used in this project and described in this section always go through
+two processes. Training and testing. To do this two sets of data are needed,
+one for each process. Data for the two processes is usually taken from the same
+dataset in different proportions. For example, some projects may take 75\% or
+90\% of the data to do training, and use the remainder to test.
+
+Testing a model with different objects then the ones used in training is very
+important. This is because a model, while training, will learn the very
+specific details of that data and may over fit. When over fitted, a model will
+not work for general examples of the data for which it was trained and will
+only ``remember'' training data instead.
+
+A way of dividing a dataset into training and testing sets is to do Cross
+Validation. Cross Validation divides the data into $ N $ folds. From the
+original $ N $ folds, a single one is taken for testing and the model is
+trained with the remaining $ N - 1 $ folds.
+
+A model is trained and tested $ N $ times, each time with a different fold
+taken for testing. Like this we get $ N $ different executions for the same
+dataset and same model and we are able to compare the best, worst, and average
+scenario in terms of training and testing.~\cite{book_dm_practical}
 

### 2017-03-27 ThesisText/experiments.tex
Adds parts for background

@@
 therefore, these are nominal fields. Because of this characteristic, Decision
-Trees and Bayesian Networks are used.~\TODO{Citation is needed on claim that
-Trees and BNets are good for nominal data}
+Trees and Bayesian Networks are used.
 
@@
         [Clustering number of approved courses based on Moodle usage results]
-        {Clustering number of approved courses based on Moodle usage results. The table displays the results for the silhouette score for clustering.}
+        {Clustering number of approved courses based on Moodle usage results.
+        The table displays the results for the silhouette score for clustering.
+        The values for every variant of the dataset are approximately the same,
+        so only one table is shown.}
 
@@
 
-\TODO{These values are the same for every variant of the dataset. This needs to be explained}
+Table~\ref{tab:exp_004_res} shows the results of this experiment. The
+silhouette score shown is approximately the same for each variant of the
+dataset, so only one table is shown for every variant.
+
+It is observable that for two clusters, the score is around 0.8, and for three
+clusters the score is around 0.75. This shows how these objects are able to be
+separated in groups of two or three without having those groups overlapping.
+
+For a number of clusters equal to or greater than four, we start getting scores
+bellow 0.7.
+
+The results shown were obtain with the K-Means algorithm. Affinity Propagation
+was also executed, but the results obtained were not good, having scores within
+0.0 and 0.5.
 

### 2017-03-27 ThesisText/background.tex
Adds detail about clustering

@@
 
-\TODO{Talk about clustering (See ind\_007 for things about this)}
-
 Educational Data Mining, \gls{edm}, refers to the act of applying Data Mining
@@
 
+Beyond the classification tasks, clustering may be used in educational data.
+The overall objective of clustering is to find groups of similar objects
+within data. In the context of~\gls{edm} that may mean looking for groups of
+students with similar characteristics without their Moodle usage habits.~\cite{ind_007,ind_008}
+
 % =============================================================================

### 2017-03-27 ThesisText/introduction.tex
Adds introduction

\chapter{Introduction}

In the last few years E-Learning as been rising in popularity as a way to
deliver education and training in a remote setting. Institutions such as
universities and companies have been taking advantage of E-Learning in
different ways. Universities have been developing remote learning courses in
order to reach more people, and companies have been developing remote training
solutions to reach people in across different physical locations

This rise in popularity of E-Learning is due to the some factors. The
advancements of technology, such as the Internet, have not only contributed to
the development of better and more complex systems, but have also made possible
for those systems to reach more people and obtain grater adoption from users
and institutions.

These E-Leaning platforms are called Learning Management Systems. The platforms
allow institutions not only to deliver training but also keep track of users,
courses, content, etc. In their normal operations, theses platforms generate
and store a lot of data. This data concerns the courses, the content in those
courses, activities from users, and so on. In some specific cases, like Moodle
in universities, the system may even keep track of the students grades.

The data becomes complex because these systems store a lot of information
regarding the users, courses, access logs, etc. Analysing this data, beyond
performing some statistical analysis, is not a trivial tasks. Some Data Mining
techniques may be apply to in order to make less trivial assumptions.

The field of study which deals with applying Data Mining techniques to
educational data is called Educational Data Mining. This projects applies EDM
to data from the Moodle of the University of Évora.~\cite{ind_010, ind_011,
ind_013, ind_014, ind_015}

In this document we start by exploring the background of~\gls{dm} and,
specifically,~\gls{edm} in section~\ref{sec:back}. The background section also
talks about some~\gls{lms} and the tools used in this project.

Before applying those techniques, we explore the original datasets of the
Moodle provided for this specific project. Section~\ref{sec:data} talks about
the original datasets, how they were explored, and how they were pre processed
into a data repository for use in this project.

Some statistical analysis are made to the pre processed data repository before
any Data Mining techniques are applied. This is done in order to to understand
the contents of the data and know which~\gls{dm} techniques should be applied
and to which subset of the data repository those techniques should be applied
to. Section~\ref{sec:stats} talks about these statistics and draws some
conclusions from them.

Section~\ref{sec:exps} talks about the Data Mining techniques which are applied
to the data and also draws some conclusions from them. Those techniques
reference the statistics made in the previous section.

Finally, section~\ref{sec:final} draws some conclusions.

### 2017-03-28 ThesisText/experiments.tex
Corrects small details

@@
 student. They also contain the identification of a course along with the
-semester and season of that course. Finally, the object contains a count of
-CRUD activities of that student in that course.
+semester and season of that course. Finally, the object contains a count
+of~\gls{crud} activities of that student in that course.
 
@@
 of the two clusters. It is observable that the major difference is in the
-number of CRUD activities. Different values for activities remain the same for
-different courses and enrollment numbers, but greater values are placed in one
-cluster while lower values are placed in another.
+number of~\gls{crud} activities. Different values for activities remain the
+same for different courses and enrollment numbers, but greater values are
+placed in one cluster while lower values are placed in another.
 
@@
 mentioned, an object has the number of enrolled and approved courses and the
-sum of each CRUD activity for every course of that single student. Each entry
-does not have the identification of the student to which it refers too because
-that information is irrelevant for our learning objectives.
+sum of each~\gls{crud} activity for every course of that single student. Each
+entry does not have the identification of the student to which it refers too
+because that information is irrelevant for our learning objectives.
 

### 2017-03-28 ThesisText/final.tex
Corrects small details

\chapter{Conclusions and Future Work}
\label{sec:final}

% =============================================================================
\section{This Project}
\TODO{}

In this project~\gls{edm} techniques were applied to the University of Évora's
Moodle data. The tasks applied to the data were based on similar projects shown
in the background section which also applied~\gls{edm} techniques to Moodle
data of other learning institutions.

In this project it was shown how the original data from the University of
Évora's Moodle was preprocessed into a data repository useful for analysis.
Some statistics over that data were then shown.

The techniques applied are divided into Supervised and Unsupervised learning. In supervised learning

% =============================================================================
\section{Future Work}
\TODO{}

### 2017-03-28 ThesisText/statistics.tex
Corrects small details

@@
 
-The activity on a course is measured using the Moodle logs which contain the
-CRUD activities. CRUD activities, as stated in~\ref{sec:logs}, are aggregated
-by courses, students, and weeks. To have a sense of how activity changes over a
-course we take the logs dataset and aggregate every object only by course. So,
-each entry will have the sum of the activities aggregated by course and week.
+The activity on a course is measured using the Moodle logs which contain
+the~\gls{crud} activities. \gls{crud} activities, as stated in~\ref{sec:logs},
+are aggregated by courses, students, and weeks. To have a sense of how activity
+changes over a course we take the logs dataset and aggregate every object only
+by course. So, each entry will have the sum of the activities aggregated by
+course and week.
 
@@
     \caption
-        [Total number of activities vs mean of activities per week]
-        {Total number of activities (right) vs total active courses (Left).}
+        [Total number of active courses per week (bars) vs total active courses (Left)]
+        {Total number of active courses per week (bars) vs total active courses (Left).}
 

### 2017-03-30 src/exp_003.py
Removes unused top line comment from experiments source

@@ -1,10 +1,5 @@
 #!/usr/bin/env python3
 
-# =============================================================================
-# TODO
-# =============================================================================
-
-
 import sys
 
 import data

### 2017-03-30 ThesisText/final.tex
Adds conclusions and future work

@@
 \label{sec:final}
-
-% =============================================================================
-\section{This Project}
-\TODO{}
 
@@
 Évora's Moodle was preprocessed into a data repository useful for analysis.
-Some statistics over that data were then shown.
+Some statistics over that data were then shown. The experiments were broken
+down by dataset. Each dataset was build from the pre processed original data.
 
-The techniques applied are divided into Supervised and Unsupervised learning. In supervised learning
+The techniques applied are divided into Supervised and Unsupervised learning.
+In supervised learning, Decision Trees and Naive Bayes Classifiers were
+applied. In unsupervised learning, clustering was done using the K-Means and
+the Affinity Propagation algorithms.
 
 % =============================================================================
-\section{Future Work}
-\TODO{}
+\section{Supervised Learning Tasks}
+
+Supervised learning techniques were applied to the different variants of the
+three described datasets. For every one of them, Decision Trees and Naive Bayes
+Classifiers were trained using the cross validation technique described in
+section~\ref{sec:cross_val}.
+
+It was shown how these models were generally able to achieve high values of
+accuracy. This means that it is possible to:
+
+\begin{enumerate}
+    \item Predict student grades from their usage of Moodle;
+        (Section~\ref{sec:exp_001})
+    \item Predict how many courses is a student going to complete given he's
+        usage of Moodle and initial number of enrolled courses;
+        (Section~\ref{sec:exp_003})
+    \item Predict student grades from his usage of features in Moodle.
+        (Section~\ref{sec:exp_005})
+\end{enumerate}
+
+Following cross validation, each model was trained ten times, each training is
+done with nine folds from the dataset, leaving one fold for testing. Despite
+having high values for maximum accuracy, and relatively high values for average
+accuracy, some models show very low values for minimum accuracy. This means
+that some instances of training have testing folds which are too different from
+the folds used in training, and therefore, the testing folds may contain
+outlier objects.
+
+Continuing the analysis in this project, it might be important to understand
+which are these outlier cases in order to achieve better accuracies in further
+supervised learning tasks.
+
+% =============================================================================
+\section{Unsupervised Learning Tasks}
+
+Clustering tasks were done and it was shown that some separation of objects
+exists in the explored datasets. Results for experiment~\ref{sec:exp_004},
+which are about course completion given the number of enrolled courses and
+Moodle usage data, were particularly promising because the results show a
+maximum and average silhouette score above 0.70 for two, three, and four
+clusters.
+
+Clustering was also done by dividing the data into folds of ten and excluding
+one fold for each training execution, and then using that fold to calculate the
+silhouette score. The difference between the maximum and minimum silhouette
+scores using this division method were noted to be minimal, sometimes not
+exceeding a value of 0.05.
+
+In these unsupervised learning tasks, the K-Means algorithm was used along with
+Affinity Propagation. However, only K-Means produced good results for
+silhouette scores.
+
+From these results, two things can be further done. First, the clustering
+results may be analysed in order to understand what are the characteristics of
+each cluster of the more successful tasks. In section~\ref{sec:exp_002}, for
+example, we see that data seems to be divisible into two clusters. An
+hypothesis for this binary division may be that one cluster contains objects in
+which students get approved in courses while the other cluster contains the
+opposite.
+
+In section~\ref{sec:exp_004}, like mentioned above, we see good scores for two,
+three, and four clusters. However, without further analysis, no hypothesis is
+presented at this time as to what are the reasons we see such good separation
+of objects in this dataset.
+
+Aside from those analysis, one second thing to do is to use different
+clustering algorithms. One algorithm to use would be K-Modes, which is a
+variant of K-Means useful for nominal data, which is present in this data.
+
+% =============================================================================
+\section{Applications}
+
+The results of an~\gls{edm} study may have many applications. Experiments such
+as the ones made to dataset~\ref{sec:exp_001_002} and~\ref{sec:exp_005_006}
+yield insight into how the basic usage of Moodle is a great indicator of the
+success of students. From the trained models, future applications may be
+develop which look into students and their usage of Moodle and finally predict
+their success even before a student as finished a course. With such
+applications we can detect student behaviour that might lead to his failure of
+a course and signal that student to take action to avoid failure.
+
+From the Moodle data described in this project, further classification and
+clustering analysis may also be done. Mode Datasets may be built and analyse from
+the original data and more algorithms may be used. For example, a dataset may
+be built that joins Moodle usage data (as described in
+sections~\ref{sec:exp_001_002} and~\ref{sec:exp_005_006}) and profiling data
+(as described in section~\ref{sec:exp_003_004}). A dataset with all these
+fields might yield better accuracies in the prediction of grades or more ways
+to do clustering.
+
+Finally, other types of Data Mining analysis may be done. For example, in the
+background section, Neural Networks are mentioned to be a usual model
+in~\gls{edm} for grade prediction. Association Rule Analysis, which is a topic
+not discussed in this project, may also be used with this data. Possibly,
+dependencies between course completion features might be detected.

### 2017-03-31 ThesisText/abstract.tex
Made abstract and sumary

\begin{tueABSTRACT}

E-Learning as been rising in popularity as a way to deliver training due to the
advancements of technologies, like the Internet. Institutions such as
universities and companies have been making use of E-Learning to deliver
training to remote locations extending their reach to students and employees
who are physically distant.

Systems called Learning Management Systems, like Moodle, exist to organize
E-Learning. They provide online platforms where professors and educators can
publish content, organize activities, perform evaluations, and so on, in order
for students to learn and get evaluated.

These systems generate and store lots of data regarding not only their usage,
but also regarding the grades of students. This kind of data is often referred
too as Educational Data. Data Mining techniques are applied to this data in
order to make non trivial assumptions. The techniques applied take inspiration
from similar projects within the field of Educational Data Mining. This field
consists in applying Data Mining Techniques to Education Data.

In this project, a data repository from the Moodle of the University of Évora
is explored. Supervised learning techniques are applied to this data in order
to show how it is possible to make predictions about the success of students
based on their usage of Moodle. Unsupervised learning techniques are also
applied in order to show how data is divisible.

\end{tueABSTRACT}

### 2017-03-31 ThesisText/sumario.tex
Made abstract and sumary

\begin{tueSUMARIO}

E-Learning tem vindo a ganhar popularidade como forma de transmitir educação
devido aos avanços nas tecnologias, como a Internet. Instituições como
universidades e empresas têm vindo a a usar E-Learning para a transmissão de
educação para locais remotos estendendo o seu alcance a estudantes e
colaboradores que estão fisicamente distantes.

Sistemas chamado ``Learning Management Systemas'', como o Moodle, existem para
organizar E-Learning. Eles oferecem plataformas online onde professores e
educadores podem publicar conteúdo, organizar actividades, fazer avaliações, e
afins, de modo a que estudantes possam aprender e serem avaliados.

Estes sistemas geram e guardam muitos dados relacionados não só com o seu uso,
mas também relacionados com notas de estudantes. Este tipo de dados são
frequentemente chamados de Dados Educacionais. Métodos de Data Mining são
aplicados a estes dados de modo a fazer suposições não triviais. As técnicas
aplicadas tomam inspiração de projectos semelhantes no campo da Data Mining
Educacional. Este campo consiste na aplicação de métodos de Data Mining a Dados
Educacionais.

Neste projecto, um repositório de dados do Moodle da Universidade de Évora é
explorado. Técnicas de aprendizagem supervisionada são aplicadas aos dados de
modo a mostrar como é possível prever o sucesso de estudantes a partir do seu
uso do Moodle. Métodos de aprendizagem não supervisionada são também aplicados
de modo a mostrar como há divisões nos dados.

\end{tueSUMARIO}

### 2017-03-31 ThesisText/experiments.tex

@@
 
+% =============================================================================
 \section{Students, Courses, Logs, and Results}
@@
 \[
-    [0, 10[\;,\, [10, 12[\;,\, [12, 14[\;,\, [14, 16[\;,\, [16, 18[\;,\, [18, 20]
+    [0, 10[\;,\, [10, 12[\;,\, [12, 14[\;,\, [14, 16[\;,\, [16, 18[\;,\, [18, 20].
 \]
@@
 \[
-    [0, 10[\;,\, [10, 14[\;,\, [14, 18[\;,\, [18, 20]
+    [0, 10[\;,\, [10, 14[\;,\, [14, 18[\;,\, [18, 20].
 \]
@@
 \subsection{Predicting Students Grades Based on Moodle Usage}
+\label{sec:exp_001}
 % EXP_001
@@
 on their Moodle usage data and in their other characteristics present on this
-dataset. In this experiment six tasks are made. Both Decision Trees and Bayesian
-Networks are trained using the dataset with binary classes, classes with
-intervals of two, and classes with intervals of four.
+dataset. In this experiment six tasks are made. Both Decision Trees and Naive
+Bayes Classifiers are trained using the dataset with binary classes, classes
+with intervals of two, and classes with intervals of four.
 
@@
 therefore, these are nominal fields. Because of this characteristic, Decision
-Trees and Bayesian Networks are used.
+Trees and Naive Bayes Classifiers are used.
 
@@
                       & Grades Classes 2 by 2 & 0.74 & 0.34 & 0.55 \\ \hline
-        Bayesian Network & Binary Class          & 0.95 & 0.53 & 0.74 \\
+        Naive Bayes   & Binary Class          & 0.95 & 0.53 & 0.74 \\
                       & Grades Classes 4 by 4 & 0.73 & 0.36 & 0.57 \\
@@
 
-It is observable that for binary classes, Bayesian Networks performed slightly
-better then decision trees. But for grade classes two by two and four by four,
-decision trees were better. In particular, training models for binary classes,
-understandably, yields better results on average.
+It is observable that for binary classes, Naive Bayes Classifiers performed
+slightly better then decision trees. But for grade classes two by two and four
+by four, decision trees were better. In particular, training models for binary
+classes, understandably, yields better results on average.
 
@@
 \subsection{Clustering Students Grades Based on Moodle Usage}
+\label{sec:exp_002}
 % EXP_002
@@
     \caption
-        []
-        {}
+        [Clustering Students Grades Based on Moodle Usage]
+        {Clustering Students Grades Based on Moodle Usage. The table displays
+        the results for the silhouette score for clustering. The values for
+        every variant of the dataset are approximately the same, so only one
+        table is shown.}
 
@@
 
+% =============================================================================
 \section{Number of Courses and Logs}
+\label{sec:exp_003_004}
 
@@
 Some classification tasks use the \texttt{ApprovedCourses} feature as a class.
-In this case, feature may be left in it's original form or it may be replaced by classes.
-There are three ways to do this replacement.~\TODO{Explain this and mentioned statistics}
+In this case, feature may be left in it's original form or it may be replaced
+by classes. There are three ways to do this replacement.
+
+The maximum number of approved courses is 12 and minimum is 0, so the possible
+divisions are with two classes:
+
+\[
+    [0, 6[\;,\, [6, 12],
+\]
+
+Three classes:
+
+\[
+    [0, 4[\;,\, [4, 8[\;,\, [8, 12],
+\]
+
+And four classes:
+
+\[
+    [0, 3[\;,\, [3, 6[\;,\, [6, 9[\;,\, [9, 12].
+\]
 
 \subsection{Predicting Number of Approved Courses Based on Moodle Usage}
+\label{sec:exp_003}
 % EXP_003
 
-Like before both Decision Trees and Bayesian Networks are used.
+Like before both Decision Trees and Naive Bayes Classifiers are used.
 Table~\ref{tab:exp_003_res} shows the maximum, minimum, and average accuracy
@@
                       & Original Classes & 0.86 & 0.14 & 0.52 \\ \hline
-        Bayesian Network & 2 Classes        & 1.00 & 0.47 & 0.70 \\
+        Naive Bayes   & 2 Classes        & 1.00 & 0.47 & 0.70 \\
                       & 4 Classes        & 0.87 & 0.29 & 0.56 \\
@@
 
-Bayesian Networks are achieve hight level of accuracy, but not as high as
-Decision Trees. For the dataset with the original classes, Bayesian Networks
-perform similarly in the best case scenario, given that the values for maximum
-accuracy are similar. But the values for minimum and average are noticeably
-lower.
+Naive Bayes Classifiers are achieve hight level of accuracy, but not as high as
+Decision Trees. For the dataset with the original classes, Naive Bayes
+Classifiers perform similarly in the best case scenario, given that the values
+for maximum accuracy are similar. But the values for minimum and average are
+noticeably lower.
 
@@
 \subsection{Clustering Number of Approved Courses Based on Moodle Usage}
+\label{sec:exp_004}
 % EXP_004
@@
 
+% =============================================================================
 \section{Student/Course Profile and Results}
+\label{sec:exp_005_006}
 
@@
 \subsection{Predicting Students Grades From Student/Course Profiling}
+\label{sec:exp_005}
 % EXP_005
@@
                       & Grades Classes 2 by 2 & 0.56 & 0.18 & 0.34 \\ \hline
-        Bayesian Network & Binary Class          & 0.97 & 0.49 & 0.77 \\
+        Naive Bayes   & Binary Class          & 0.97 & 0.49 & 0.77 \\
                       & Grades Classes 4 by 4 & 0.91 & 0.31 & 0.55 \\
@@
 
-We observe from the results that Bayesian Networks perform better then Decision
-Trees achieving greater levels of accuracy for each dataset variant.
+We observe from the results that Naive Bayes Classifiers perform better then
+Decision Trees achieving greater levels of accuracy for each dataset variant.
 
 \subsection{Clustering Students Grades From Student/Course Profiling}
+\label{sec:exp_006}
 % EXP_006
@@
     \caption
-        []
-        {}
+        [Clustering Students Grades From Student/Course Profiling]
+        {Clustering Students Grades From Student/Course Profiling. The values
+        for every variant of the dataset are approximately the same, so only
+        one table is shown.}
 

### 2017-03-31 ThesisText/final.tex

@@
         (Section~\ref{sec:exp_001})
-    \item Predict how many courses is a student going to complete given he's
+    \item Predict how many courses a student is going to complete given his
         usage of Moodle and initial number of enrolled courses;

### 2017-03-31 ThesisText/introduction.tex

@@
 
-These E-Leaning platforms are called Learning Management Systems. The platforms
+These E-Leaning platforms are called~\gls{lms}. The platforms
 allow institutions not only to deliver training but also keep track of users,
@@
 
-The field of study which deals with applying Data Mining techniques to
-educational data is called Educational Data Mining. This projects applies EDM
-to data from the Moodle of the University of Évora.~\cite{ind_010, ind_011,
-ind_013, ind_014, ind_015}
+The field of study which deals with applying~\gls{dm} techniques to educational
+data is called~\gls{edm}. This projects applies~\gls{edm} to data from the
+Moodle of the University of Évora.~\cite{ind_010, ind_011, ind_013, ind_014,
+ind_015}
 
@@
 Some statistical analysis are made to the pre processed data repository before
-any Data Mining techniques are applied. This is done in order to to understand
-the contents of the data and know which~\gls{dm} techniques should be applied
-and to which subset of the data repository those techniques should be applied
-to. Section~\ref{sec:stats} talks about these statistics and draws some
-conclusions from them.
+any~\gls{dm} techniques are applied. This is done in order to to understand the
+contents of the data and know which~\gls{dm} techniques should be applied and
+to which subset of the data repository those techniques should be applied to.
+Section~\ref{sec:stats} talks about these statistics and draws some conclusions
+from them.
 

### 2017-04-01 ThesisText/experiments.tex
Corrected typos

@@
 It is observable that for binary classes, Naive Bayes Classifiers performed
-slightly better then decision trees. But for grade classes two by two and four
+slightly better than decision trees. But for grade classes two by two and four
 by four, decision trees were better. In particular, training models for binary
@@
 from the \texttt{Results} and \texttt{MoodleLogs} datasets.
-Table~\ref{tab_dat_002} shows the features of this dataset.
+Table~\ref{tab:dat_002} shows the features of this dataset.
 
@@
 
-We observe from the results that Naive Bayes Classifiers perform better then
+We observe from the results that Naive Bayes Classifiers perform better than
 Decision Trees achieving greater levels of accuracy for each dataset variant.

### 2017-04-01 ThesisText/introduction.tex
Corrected typos

@@
 
-These E-Leaning platforms are called~\gls{lms}. The platforms
+These E-Learning platforms are called~\gls{lms}. The platforms
 allow institutions not only to deliver training but also keep track of users,
@@
 Moodle provided for this specific project. Section~\ref{sec:data} talks about
-the original datasets, how they were explored, and how they were pre processed
+the original datasets, how they were explored, and how they were preprocessed
 into a data repository for use in this project.
 
-Some statistical analysis are made to the pre processed data repository before
+Some statistical analysis are made to the preprocessed data repository before
 any~\gls{dm} techniques are applied. This is done in order to to understand the

### 2017-04-01 ThesisText/sumario.tex
Corrected typos

@@
 
-E-Learning tem vindo a ganhar popularidade como forma de transmitir educação
-devido aos avanços nas tecnologias, como a Internet. Instituições como
-universidades e empresas têm vindo a a usar E-Learning para a transmissão de
-educação para locais remotos estendendo o seu alcance a estudantes e
-colaboradores que estão fisicamente distantes.
+E-Learning tem vindo a ganhar popularidade como forma de transmissão de
+conhecimentos a nível educacional graças aos avanços nas tecnologias, como por
+exemplo, a Internet. Instituições como universidades e empresas têm vindo a
+usar E-Learning para a transmissão de conteúdos educacionais para locais
+remotos estendendo o seu alcance a estudantes e colaboradores que estão
+fisicamente distantes.
 
-Sistemas chamado ``Learning Management Systemas'', como o Moodle, existem para
+Sistemas chamados ``Learning Management Systems'', como o Moodle, existem para
 organizar E-Learning. Eles oferecem plataformas online onde professores e
 educadores podem publicar conteúdo, organizar actividades, fazer avaliações, e
-afins, de modo a que estudantes possam aprender e serem avaliados.
+outro tipo de ações relacionadas, de modo a que os estudantes possam aprender e
+serem avaliados.
 
@@
 
-Neste projecto, um repositório de dados do Moodle da Universidade de Évora é
+Neste projecto, um repositório de dados do Moodle da Universidade de Évora foi
 explorado. Técnicas de aprendizagem supervisionada são aplicadas aos dados de

### 2017-04-01 ThesisText/background.tex
Adds introduction
Adds conclusions and future work

@@
 \chapter{Background}
+\label{sec:back}
 
@@
 
-\section{Cross Validation}
+\subsection{Cross Validation}
+\label{sec:cross_val}
 
@@
 
-Testing a model with different objects then the ones used in training is very
+Testing a model with different objects than the ones used in training is very
 important. This is because a model, while training, will learn the very
@@
 
-Educational Data Mining, \gls{edm}, refers to the act of applying Data Mining
-techniques to educational data. Educational data refers to data which origin is
-in the act of teaching and training, in the activities of professors and
-students, results of evaluation, and so on. In this project, the origin of data
-is the University of Évora's Moodle.
+\acrfull{edm}, refers to the act of applying Data Mining techniques to
+educational data. Educational data refers to data which origin is in the act of
+teaching and training, in the activities of professors and students, results of
+evaluation, and so on. In this project, the origin of data is the University of
+Évora's Moodle.
 
@@
 
-One such example of~\gls{edm} Data Mining tasks is too predict student grades
-based features from the students themselves, the courses they are undertaking,
-and their activity in a~\gls{lms}. By training such models, professors can make
+One such example of an~\gls{edm} tasks is to predict student grades based
+features from the students themselves, the courses they are undertaking, and
+their activity in a~\gls{lms}. By training such models, professors can make
 decisions about how courses are organized regarding students success. For
@@
 In~\cite{ind_006}, networks are trained from Moodle logs. The data from the
-logs are structured in a slightly different way then the previous studies. The
+logs are structured in a slightly different way than the previous studies. The
 features of this data are shown in table~\ref{tab:ind_006_fields}.

### 2017-04-01 ThesisText/data.tex
Minor corrections on statistics
Adds introduction

@@
 \chapter{University of Évora's Moodle Data}
+\label{sec:data}
 
@@
 
-To begin with, we explore what was initially provided in terms of data. The
-repository is a spreadsheet book with 12 sheets. The sheets contain raw data
-and some calculated fields.
+To begin with, we explore the data which was initially provided. The repository
+is a spreadsheet book with 12 sheets. The sheets contain raw data and some
+calculated fields.
 
@@
 like Python with Pandas, they were exported to a CSV format. Some
-transformations were applied to these files in other to clean useless data
+transformations were applied to these files in order to clean useless data
 points or make the data easier to process.
@@
 reasons. First, some analysis are not useful so them being in the data only
-makes the data more complex. By removing such analysis we get a simpler data
+make it more complex. By removing such analysis we get a simpler data
 model. Second, even if some analysis are in fact useful, they are still easily
@@
 reproduced when needed for specific experiments.
-
-Some analysis were made and are described in section~\ref{set:stats}
 
@@
 
+There are a total of 331 entries in the Profile dataset.
+
 \subsection{Moodle Logs}
@@
 features of the Moodle logs, and table~\ref{tab:moodle_logs_enum} shows the
-values of the enumerated fields. There are a total of 25207 entries in the
+values of the enumerated fields. There are a total of 25206 entries in the
 Logs.
@@
     \caption
-        [Moode Logs features]
-        {Moode Logs features.}
+        [Moodle Logs features]
+        {Moodle Logs features.}
 
@@
 
-From the original data we make an initial pre processing in order to have a
-consistent repository of datasets ready for further analysis. The pre-processed
+From the original data we make an initial preprocessing in order to have a
+consistent repository of datasets ready for further analysis. The preprocessed
 datasets have consistent names and fields.
@@
 section~\ref{sec:profile}. The fields are shown in
-table~\ref{tab:profile_fields}.
+table~\ref{tab:profile_features}.
 

### 2017-04-01 ThesisText/statistics.tex

@@
 
-To begin exploring the courses some counts are made. We count the number of
+To begin exploring the courses, some counts are made. We count the number of
 overall courses, the number of departments, and number of degrees.
@@
 part of 9 different departments and make up 11 degrees. While the number of
-courses is much greater then some experiments mentioned in the background, it
+courses is much greater than some experiments mentioned in the background, it
 is important to understand how many students does each of these courses have.
@@
 
-Table~\ref{tab:moodle_users_stats} shows some stats regarding Moodle users.
+Table~\ref{tab:moodle_users_stats} shows some counts regarding Moodle users.
 Some of the explored projects listed in the background, like~\cite{ind_001},
-have a number of students around 800 to 900 which is greater then the number of
+have a number of students around 800 to 900 which is greater than the number of
 students in this dataset, but they also have a lower number of courses, for
@@
     \caption
-        [Professor features]
-        {Professor features.}
+        [Moodle users counts]
+        {Moodle users counts.}
 
@@
 Figure~\ref{fig:stat_004_all} shows the graph for these courses. The number of
-read activities and number of total activities is much greater then the number
+read activities and number of total activities is much greater than the number
 of create, update, and delete activities. Because of this, we plot a graph
@@
 The number of total activities per week drops significantly. This is because
-some courses are longer then others. For example, some courses may have their
+some courses are longer than others. For example, some courses may have their
 last evaluation in the first weeks of the exam season, while others may still
@@
     \caption
-        [Statistics on the Results]
-        {Statistics on the Results.}
+        [Counts on the Results]
+        {Counts on the Results.}
 

### 2017-04-02 ThesisText/introduction.tex
Corrects typos

@@
 
-In the last few years E-Learning as been rising in popularity as a way to
+In the last few years E-Learning has been rising in popularity as a way to
 deliver education and training in a remote setting. Institutions such as
@@
 order to reach more people, and companies have been developing remote training
-solutions to reach people in across different physical locations
+solutions to reach people across different physical locations.
 
@@
 the development of better and more complex systems, but have also made possible
-for those systems to reach more people and obtain grater adoption from users
+for those systems to reach more people and obtain greater adoption from users
 and institutions.
@@
 courses, activities from users, and so on. In some specific cases, like Moodle
-in universities, the system may even keep track of the students grades.
+in universities, the system may even keep track of the student's grades.
 
@@
 regarding the users, courses, access logs, etc. Analysing this data, beyond
-performing some statistical analysis, is not a trivial tasks. Some Data Mining
-techniques may be apply to in order to make less trivial assumptions.
+performing some statistical analysis, is not a trivial task. Some Data Mining
+techniques may be applied in order to make less trivial assumptions.
 
@@
 Some statistical analysis are made to the preprocessed data repository before
-any~\gls{dm} techniques are applied. This is done in order to to understand the
+any~\gls{dm} techniques are applied. This is done in order to understand the
 contents of the data and know which~\gls{dm} techniques should be applied and

### 2017-04-02 ThesisText/statistics.tex
Corrects typos

@@
 Before applying any Data Mining methods, it is important to understand how the
-data is constituted. Knowing things like, how many courses are there, how many
+data is constituted. Knowing things like, how many courses there are, how many
 students, how many students per course, etc, is important in order to train
@@
 courses is much greater than some experiments mentioned in the background, it
-is important to understand how many students does each of these courses have.
+is important to understand how many students each of these courses have.
 
@@
 It was to be expected that every degree would have a similar number of courses,
-or at least that every course have curses for both semesters. It is unknown why
+or at least that every degree has courses for both semesters. It is unknown why
 there are degrees present that have such characteristics, whether it is just
@@
 Observing the behaviour of students in most courses in any University, it is
-common to find that the number of students decreases over time has the course
+common to find that the number of students decreases over time as the course
 progresses. This is due to students giving up on that course for that
@@
     \caption
-        [Total number of active courses per week (bars) vs total active courses (Left)]
-        {Total number of active courses per week (bars) vs total active courses (Left).}
+        [Total number of active courses per week (Bars) vs total active courses (Left)]
+        {Total number of active courses per week (Bars) vs total active courses (Left).}
 

### 2017-06-05 presentation/intro.tex
Started presentation

\section{Introduction}

\begin{frame}
\frametitle{Introduction}
TODO
\end{frame}

### 2017-07-05 presentation/final.tex

\section{Conclusions}

% =============================================================================
\begin{frame}
\frametitle{Conclusions}

\center{\textbf{Recap}}

\vspace{0.5cm}

\begin{itemize}
    \item Objects and their features were extracted from the data
    \item Statistics were calculated
    \item Models were trained using that data
\end{itemize}

\center{\textbf{Results}}

\vspace{0.5cm}

\begin{itemize}
    \item It is possible to make certain prediction about student success based on Moodle Data
    \item Some clusters exits in the data
\end{itemize}

\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Conclusions}

\center{\textbf{Future Work}}

\begin{itemize}
    \item Further analyses to clustering results
    \item Train other models such as Association Rules and Neural Networks
    \item Find why some instances of training have very low accuracy values
\end{itemize}

\vspace{0.5cm}

\center{\textbf{Applications}}

\begin{itemize}
    \item Make systems in which the success of students is predicted before he finishes a course
    \item Such system may signal professors and students about the possibility of failure of some people in courses
\end{itemize}

\end{frame}

% =============================================================================
\begin{frame}
\frametitle{Conclusions}

\center{\textbf{Thank you for Listening}}

\end{frame}

### 2017-07-05 presentation/intro.tex

@@
 
+% =============================================================================
 \begin{frame}
 \frametitle{Introduction}
-TODO
+
+\center{\textbf{Project Consists On}}
+
+\vspace{0.5cm}
+
+\begin{itemize}
+    \item Exploring data from the Moodle of the University of Évora
+    \item Applying Educational Data Mining (EDM) methods to that data
+\end{itemize}
+
+\center{\textbf{Presented Over the Following Steps}}
+
+\vspace{0.5cm}
+
+\begin{itemize}
+    \item Looking into what E-Learning is and how it is implemented
+    \item Exploring Data Mining and Educational Data Mining
+    \item Analysing the data from Évora's Moodle
+    \item Training models to find interesting patterns in data
+\end{itemize}
+
 \end{frame}
+
+% =============================================================================
+\begin{frame}
+\frametitle{Background}
+
+\center{\textbf{E-Learning, LMSs, and Moodle}}
+
+\begin{itemize}
+    \item E-Learning is a way of delivering education and training remotely
+    \item Due to advancements in technologies, E-Learning has been becoming more popular
+    \item Learning Management Systems (LMS), like Moodle, have been successfully used to implement E-Learning
+    \item These systems generate lots of data which are not trivially analysed
+\end{itemize}
+
+\end{frame}
+
+% =============================================================================
+\begin{frame}
+\frametitle{Background}
+
+\center{\textbf{Data Mining}}
+
+\begin{itemize}
+    \item DM aims at finding patterns in data in order to make non trivial assumptions
+    \item This is done by training models using that data
+\end{itemize}
+
+\center{\textbf{Models}}
+
+\begin{itemize}
+    \item Supervised Learning
+    \begin{itemize}
+        \item Decision Trees
+        \item Bayesian Networks
+    \end{itemize}
+
+    \item Unsupervised Learning
+    \begin{itemize}
+        \item K-Means
+        \item Affinity Propagation
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
+% =============================================================================
+\begin{frame}
+\frametitle{Background}
+
+\center{\textbf{Educational Data Mining}}
+
+\begin{itemize}
+    \item Educational Data Mining consists in applying Data Mining methods to educational data
+    \item Successful experiments have been made in which student grades are predicted from their usage of LMS systems
+    \item The used models are often Decision Trees, Bayesian Networks, Neural Networks, among others
+\end{itemize}
+
+\end{frame}

### 2017-07-05 presentation/statistics.tex

\section{Statistics on the Data}

\begin{frame}
\frametitle{Statistics on the Data - Courses}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Courses in Numbers
    \end{itemize}

    \begin{tabular}{| l | r |}
        \hline
        Number of Courses     & 77 \\ \hline
        Number of Departments & 9  \\ \hline
        Number of Degrees     & 11 \\ \hline
    \end{tabular}

    \vspace{1cm}

    \begin{itemize}
        \item 28 out of the 77 courses are from Computer Engineering
        \item Most degrees have courses in both semesters
    \end{itemize}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Number of courses group by credits
    \end{itemize}

    \begin{tabular}{| l | r |}
        \hline
        Credits & Number of Courses \\ \hline
        1       & 3                 \\ \hline
        2       & 3                 \\ \hline
        3       & 3                 \\ \hline
        5       & 6                 \\ \hline
        6       & 46                \\ \hline
        7.5     & 12                \\ \hline
        12      & 2                 \\ \hline
        15      & 2                 \\ \hline
    \end{tabular}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Courses}

\centering
\includegraphics[width=9cm]{../src/stat_007_results/plot}

\begin{itemize}
    \item {\color{gray} Number of Students per Course} and {\color{blue} Number of Approved Students per Course}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Moodle Users Counts
    \end{itemize}

    \begin{tabular}{| l | r |}
        \hline
        Number of Students   & 145 \\ \hline
        Number of Professors & 87  \\ \hline
        Total                & 232 \\ \hline
    \end{tabular}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \begin{itemize}
        \item Some similar studies have numbers of students between 800 and 900
        \item But they also have lower numbers of courses such as 7
    \end{itemize}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}

\center{\textbf{CRUD of Most Active Courses}}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_all}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_0_MAT7514_1_all}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_1_eDocOn-T1_all}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_1_GES11656_all}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}

\center{\textbf{CRUD of Median Active Courses}}

\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_0_CTE-UMCED-2_cud}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_0_MAT7514_1_cud}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_most_1_eDocOn-T1_cud}

    \includegraphics[width=4.7cm]{../src/stat_004_results/fig_middle_1_GES11656_cud}
\end{column}%
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}
\centering

\center{\textbf{Total Activities per Week vs Mean Activities per Week}}

\includegraphics[width=9cm]{../src/stat_005_results/plot}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Moodle Users}
\centering

\center{\textbf{Total Number of Active Courses per Week (Bars) vs Total Active Courses (Line)}}

\includegraphics[width=9cm]{../src/stat_006_results/plot}
\end{frame}

\begin{frame}
\frametitle{Statistics on the Data - Results}

{

\center

\begin{tabular}{| l | r |}
    \hline
    Number of results           & 716 \\ \hline
    Number of non-empty results & 659  \\ \hline
    Number of approved students & 334 \\ \hline
\end{tabular}

}

\end{frame}

### 2017-07-05 ThesisText/background.tex

@@
 create and publish content, perform evaluation of students, etc. Students or
-Learners are the users which actually perform the training and follow courses
-in the platforms.
+Learners are the users who actually perform the training and follow courses in
+the platforms.
 
@@
 The contents of a course may be resources such as PDF files, links, media such
-as images, videos, etc. These resources should be guides for students achieve
-their objectives in learning, or should be resources needed for the student to
-perform certain tasks. A course may also have activities. Activities may be,
-for example, projects that a student needs to complete before a given deadline.
-Activities may also be quizzes made to students over Moodle. These quizzes may
-be used for evaluation.
+as images, videos, etc. These resources should be guides for students to
+achieve their objectives in learning, or should be resources needed for the
+student to perform certain tasks. A course may also have activities. Activities
+may be, for example, projects that a student needs to complete before a given
+deadline.  Activities may also be quizzes made to students over Moodle. These
+quizzes may be used for evaluation.
 
@@
 One difficulty of clustering is to find the number of clusters that optimally
-divide the data. Some algorithms take a number of clusters as an arguments,
+divide the data. Some algorithms take a number of clusters as an argument,
 while others are able to find an optimal value.
 
-By it's own nature, it may not be possible to find clusters in data, that is,
+By its own nature, it may not be possible to find clusters in data, that is,
 there may not be any great division of objects given their features in a given
@@
 
-K-Means is an clustering algorithm that find $ K $ clusters in data. The $ K $
+K-Means is a clustering algorithm that find $ K $ clusters in data. The $ K $
 variable is given to the algorithm as an argument, hence the name K-Means. The
@@
 The model used in this project and described in this section always go through
-two processes. Training and testing. To do this two sets of data are needed,
+two processes, training and testing. To do this, two sets of data are needed,
 one for each process. Data for the two processes is usually taken from the same
@@
 The main objective of~\gls{edm} is to extract meaningful information from
-educational data like web logs. It is not trivial to do so without advance Data
-Mining techniques. \gls{edm}, therefore, aims at training models based on
+educational data like web logs. It is not trivial to do so without advanced
+Data Mining techniques. \gls{edm}, therefore, aims at training models based on
 education data for a variety of tasks.
 
-One such example of an~\gls{edm} tasks is to predict student grades based
+One such example of an~\gls{edm} task is to predict student's grades based
 features from the students themselves, the courses they are undertaking, and
@@
 example, a certain number of activities in a given course may be beneficial to
-it's success because it keeps students engaged, but any number of activities
+its success because it keeps students engaged, but any number of activities
 more might not make a difference.~\cite{ind_001, ind_002, ind_005}
@@
 
-In~\cite{ind_001}, some models are trained in order to predict student grades.
-The tasks described in the paper are classification tasks, in which a student
-is assigned to a given class corresponding to their success in courses based on
-their usage of a~\gls{lms} platform. The classes available are Excellent, Good,
-Pass, and Fail.
+In~\cite{ind_001}, some models are trained in order to predict student's
+grades.  The tasks described in the paper are classification tasks, in which a
+student is assigned to a given class corresponding to their success in courses
+based on their usage of a~\gls{lms} platform. The classes available are
+Excellent, Good, Pass, and Fail.
 
@@
 In~\cite{ind_006}, networks are trained from Moodle logs. The data from the
-logs are structured in a slightly different way than the previous studies. The
+logs is structured in a slightly different way than the previous studies. The
 features of this data are shown in table~\ref{tab:ind_006_fields}.

### 2017-07-05 ThesisText/data.tex

@@
 
-The rest of the sheets contain the objects which are going to be used in the
+The rest of the sheets contains the objects which are going to be used in the
 Data Mining tasks. Those objects and their features are fairly unorganized in
@@
 The logs contain the actions of the Moodle users which role can be of a
-professors or a students. The data contains features for Students but no
+professor or a student. The data contains features for students but no
 features for professors. Table~\ref{tab:students_features}, contains all the
@@
 From table~\ref{tab:students_features} it is observable that a student may be
-identified by it's internal Moodle Student ID, or by it's student number as
+identified by its internal Moodle Student ID, or by its student number as
 given by SIIUE. Some students in this data repository do not have a
@@
 The fields \textit{Results} and \textit{FinalResult} are always empty together
-and in that cases there is never any grade.
+and in those cases there is never any grade.
 
@@
 Each Moodle user has a username, an actual name, and a role. That information
-is stored in \texttt{MoodleUsers}. This dataset contains users which are both
+is stored in \texttt{MoodleUsers}. This dataset contains users who are both
 professors and students. They are identified by their role.

### 2017-07-05 ThesisText/experiments.tex

@@
 
-\subsection{Predicting Students Grades Based on Moodle Usage}
+\subsection{Predicting Student's Grades Based on Moodle Usage}
 \label{sec:exp_001}
@@
 
-A first experiment on this dataset aims at predicting the students grades based
+A first experiment on this dataset aims at predicting the student's grades based
 on their Moodle usage data and in their other characteristics present on this
@@
     \caption
-        [Predicting students grades based on Moodle usage results]
-        {Predicting students grades based on Moodle usage results. Two models
+        [Predicting student's grades based on Moodle usage results]
+        {Predicting student's grades based on Moodle usage results. Two models
         are trained 10 times on three different variations of the same dataset.
@@
 
-\subsection{Clustering Students Grades Based on Moodle Usage}
+\subsection{Clustering Student's Grades Based on Moodle Usage}
 \label{sec:exp_002}
@@
     \caption
-        [Clustering Students Grades Based on Moodle Usage]
-        {Clustering Students Grades Based on Moodle Usage. The table displays
+        [Clustering Student's Grades Based on Moodle Usage]
+        {Clustering Student's Grades Based on Moodle Usage. The table displays
         the results for the silhouette score for clustering. The values for
@@
 
-Table~\ref{tab:exp_002a_res} has a listing of objects that are placed in each
-of the two clusters. It is observable that the major difference is in the
-number of~\gls{crud} activities. Different values for activities remain the
-same for different courses and enrollment numbers, but greater values are
-placed in one cluster while lower values are placed in another.
+Table in attachment~\ref{attach:exp_002} has a listing of objects that are
+placed in each of the two clusters. It is observable that the major difference
+is in the number of~\gls{crud} activities. Different values for activities
+remain the same for different courses and enrollment numbers, but greater
+values are placed in one cluster while lower values are placed in another.
 
@@
 Some classification tasks use the \texttt{ApprovedCourses} feature as a class.
-In this case, feature may be left in it's original form or it may be replaced
-by classes. There are three ways to do this replacement.
+In this case, feature may be left in its original form or it may be replaced by
+classes. There are three ways to do this replacement.
 
@@
 
-The results shown were obtain with the K-Means algorithm. Affinity Propagation
-was also executed, but the results obtained were not good, having scores within
-0.0 and 0.5.
+The results shown were obtained with the K-Means algorithm. Affinity
+Propagation was also executed, but the results obtained were not good, having
+scores within 0.0 and 0.5.
 
@@
 
-In this dataset we relate the profiling of a student in a courses and his
+In this dataset we relate the profiling of a student in a course and his
 results on that course. To begin constructing this dataset we use the Profiling
@@
 
-\subsection{Predicting Students Grades From Student/Course Profiling}
+\subsection{Predicting Student's Grades From Student/Course Profiling}
 \label{sec:exp_005}
@@
 
-Using the described dataset, we train models to predict student grades based on
-profiling data. Table~\ref{tab:exp_005_res} shows the results of six tasks.
-Three for each Model. Each tasks is done in a variant of the dataset.
+Using the described dataset, we train models to predict student's grades based
+on profiling data. Table~\ref{tab:exp_005_res} shows the results of six tasks,
+three for each Model. Each task is done in a variant of the dataset.
 
@@
     \caption
-        [Predicting students grades from student/course profiling results]
-        {Predicting students grades from student/course profiling results. The
+        [Predicting student's grades from student/course profiling results]
+        {Predicting student's grades from student/course profiling results. The
         results show the maximum, minimum, and average for each trained model
@@
 
-\subsection{Clustering Students Grades From Student/Course Profiling}
+\subsection{Clustering Student's Grades From Student/Course Profiling}
 \label{sec:exp_006}
@@
     \caption
-        [Clustering Students Grades From Student/Course Profiling]
-        {Clustering Students Grades From Student/Course Profiling. The values
+        [Clustering Student's Grades From Student/Course Profiling]
+        {Clustering Student's Grades From Student/Course Profiling. The values
         for every variant of the dataset are approximately the same, so only

### 2017-07-05 ThesisText/final.tex
Corrected typos

@@
 Some statistics over that data were then shown. The experiments were broken
-down by dataset. Each dataset was build from the pre processed original data.
+down by dataset. Each dataset was build from the preprocessed original data.
 
@@
 \begin{enumerate}
-    \item Predict student grades from their usage of Moodle;
+    \item Predict student's grades from their usage of Moodle;
         (Section~\ref{sec:exp_001})
@@
         (Section~\ref{sec:exp_003})
-    \item Predict student grades from his usage of features in Moodle.
+    \item Predict student's grades from his usage of features in Moodle.
         (Section~\ref{sec:exp_005})
@@
 success of students. From the trained models, future applications may be
-develop which look into students and their usage of Moodle and finally predict
-their success even before a student as finished a course. With such
+developed which look into students and their usage of Moodle and finally
+predict their success even before a student as finished a course. With such
 applications we can detect student behaviour that might lead to his failure of
@@
 From the Moodle data described in this project, further classification and
-clustering analysis may also be done. Mode Datasets may be built and analyse from
-the original data and more algorithms may be used. For example, a dataset may
-be built that joins Moodle usage data (as described in
+clustering analysis may also be done. More datasets may be built and analysed
+from the original data and more algorithms may be used. For example, a dataset
+may be built that joins Moodle usage data (as described in
 sections~\ref{sec:exp_001_002} and~\ref{sec:exp_005_006}) and profiling data

### 2017-07-16 ThesisText/background.tex
Made corrections

@@
 
-K-Means is a clustering algorithm that find $ K $ clusters in data. The $ K $
+K-Means is a clustering algorithm that finds $ K $ clusters in data. The $ K $
 variable is given to the algorithm as an argument, hence the name K-Means. The
 algorithm works by creating $ K $ random points based on the features of the
-data. The algorithm then iterates updating the location of those points until
-they converge.~\cite{book_ml,book_dm_practical}
+data.
+
+Each point has has many coordinates as there are features, and the domains of
+the coordinates correspond to the domains of the features. Each object in the
+data gets associated with the closets point to it. A set of objects associated
+with the same point is called a cluster.
+
+The algorithm then iterates updating the location of those points by
+calculating the center of of the cluster until convergence is
+reached.~\cite{book_ml,book_dm_practical}
 
@@
 take a number of clusters as an argument. This algorithm finds the optimal
-number of clusters.~\cite{book_ml,book_dm_practical}
+number of clusters.
+
+To find clusters, the algorithm keeps two matrices, a responsibility matrix and
+an availability matrix. Each matrix relates each point in the data to every
+other point. Each step of the algorithm iterates by updating the information on
+these matrices until there is convergence.
+
+~\cite{book_ml,book_dm_practical}
 

### 2017-07-16 ThesisText/data.tex
Made corrections

@@
 in Moodle, and their results. The data comes from Moodle databases and from the
-SIIUE system. Both systems are integrated, making Moodle plus SIIUE the
-University's own~\gls{lms}.
+SIIUE system. SIIUE is the integrated information system of the university. It
+is the system which contains information on courses, students, their grades,
+among many other things. Both systems are integrated, making Moodle plus SIIUE
+the University's own~\gls{lms}.
 
@@
     \item \texttt{export\_activity}
+    \item \texttt{CalculosAlunos.csv}
 \end{itemize}
@@
 Data Mining tasks. Those objects and their features are fairly unorganized in
-the original data, so an analysis on of the available features had to be done.
+the original data, so an analysis of the available features had to be done.
 Section~\ref{sec:objs_feat} states the objects found in the data and details
@@
 After a first analysis of the data, the objects are identified. These objects
-are courses, students, results, etc. Data mining tasks are applied over these
-objects which are described by their features.
+are courses, Moodle users, results, profile, and Moodle logs. Data mining tasks
+are applied over these objects which are described by their features.
 
@@
 5. If it is a student, it has a value of 3. A few users with a value of 4 are
-present. These users have the role of editor, but the Moodle logs only have 11
+present. These users have the role of editor, but the Moodle logs only have 10
 entries for a user of this role, so they are simply discarded.
@@
         \textit{Name}           & String    \\
+        \textit{Role}           & Enumerated \\
     \end{tabular}

### 2017-07-16 ThesisText/experiments.tex
Made corrections

@@
         [Students, Courses, Logs, and Results]
-        {Students, Courses, Logs, and Results. This dataset joins information
-        from four processed datasets in order to relate a student, a course,
-        usage data from that student on that course, and the final grade of it}
+        {Students, Courses, Logs, and Results.}
 
@@
         [Predicting student's grades based on Moodle usage results]
-        {Predicting student's grades based on Moodle usage results. Two models
-        are trained 10 times on three different variations of the same dataset.
-        The maximum, minimum, and average accuracy is displayed for those ten
-        executions.}
+        {Predicting student's grades based on Moodle usage results.}
 
@@
         [Clustering Student's Grades Based on Moodle Usage]
-        {Clustering Student's Grades Based on Moodle Usage. The table displays
-        the results for the silhouette score for clustering. The values for
-        every variant of the dataset are approximately the same, so only one
-        table is shown.}
+        {Clustering Student's Grades Based on Moodle Usage.}
 

### 2017-07-16 ThesisText/statistics.tex
Made corrections

@@
         {Number of Students per Course and Number of Approved Students per
-        Course. The grey bars show the number of students, while the blue bars
-        show the number of approved students. The graph}
+        Course.}
 
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_0_HIS10755_all}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_0_PED9078_all}
         \caption[]{}
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_1_MAT7515_all}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_1_HIS10755_all}
         \caption[]{}
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_0_HIS10755_cud}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_0_PED9078_cud}
         \caption[]{}
@@
         \centering
-        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_1_MAT7515_cud}
+        \includegraphics[width=\linewidth]{../src/stat_004_results/fig_least_1_HIS10755_cud}
         \caption[]{}
@@
     \caption
-        [Total number of active courses per week (Bars) vs total active courses (Left)]
-        {Total number of active courses per week (Bars) vs total active courses (Left).}
+        [Total number of active courses per week vs total active courses]
+        {Total number of active courses per week (Bars) vs total active courses (Line).}
 
@@
 Table~\ref{tab:results_stats} shows some statistics on the results data. We see
-that from the available 716 entries, 334 of the, refer to instances of a
-student getting approved on a course.
+that from the available 716 entries, 334 refer to instances of a student
+getting approved on a course.
 
